### 2020-07-18 Rmd/msc_thesis_figures.tex
commit report

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Msc Thesis Figures},
  pdfauthor={Valentin Lucet},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\title{Msc Thesis Figures}
\author{Valentin Lucet}
\date{17/07/2020}

\begin{document}
\maketitle

\hypertarget{data-prep}{%
\subsubsection{Data prep}\label{data-prep}}

\hypertarget{cleaning-and-formatting}{%
\paragraph{Cleaning and formatting}\label{cleaning-and-formatting}}

We start by preparing the data from the outputs of scripts \texttt{7.1}
and \texttt{8.1}.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Results dir + mun shapefile}
\NormalTok{sce_dir_vec <-}\StringTok{ }\KeywordTok{list.files}\NormalTok{(}\StringTok{"libraries/stsim/monteregie-conncons-scripted.ssim.output"}\NormalTok{, }
                          \DataTypeTok{full.names =}\NormalTok{ T)}
\NormalTok{mun <-}\StringTok{ }\KeywordTok{st_read}\NormalTok{(}\StringTok{"data/mun/munic_SHP_clean.shp"}\NormalTok{, }\DataTypeTok{quiet =} \OtherTok{TRUE}\NormalTok{)}

\CommentTok{# results key}
\NormalTok{results_clean <-}\StringTok{ }\KeywordTok{read_rds}\NormalTok{(}\StringTok{"data/temp/stsim_run_results.RDS"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(scenarioId, name) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{name =}  \KeywordTok{gsub}\NormalTok{(.}\OperatorTok{$}\NormalTok{name, }\DataTypeTok{pattern =} \StringTok{" }\CharTok{\textbackslash{}\textbackslash{}}\StringTok{(.+}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{)"}\NormalTok{, }\DataTypeTok{replacement =} \StringTok{""}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sce =}  \KeywordTok{paste0}\NormalTok{(}\StringTok{"sce_"}\NormalTok{, .}\OperatorTok{$}\NormalTok{scenarioId)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{chapter =} \KeywordTok{c}\NormalTok{(}\StringTok{"none"}\NormalTok{, }\KeywordTok{rep}\NormalTok{(}\StringTok{"chap_1"}\NormalTok{, }\DecValTok{6}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\StringTok{"chap_2"}\NormalTok{, }\DecValTok{9}\NormalTok{))) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{splitted =} \KeywordTok{str_split}\NormalTok{(name, }\StringTok{" }\CharTok{\textbackslash{}\textbackslash{}}\StringTok{| "}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{climate =} \KeywordTok{unlist}\NormalTok{(}\KeywordTok{map}\NormalTok{(splitted, }\OperatorTok{~}\KeywordTok{unlist}\NormalTok{(.x[}\DecValTok{2}\NormalTok{])))) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{run =} \KeywordTok{unlist}\NormalTok{(}\KeywordTok{map}\NormalTok{(splitted, }\OperatorTok{~}\KeywordTok{unlist}\NormalTok{(.x[}\DecValTok{1}\NormalTok{])))) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{splitted) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{replace_na}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DataTypeTok{climate =} \StringTok{"none"}\NormalTok{))}

\CommentTok{# Final extracted datasets}
\NormalTok{df_final <-}\StringTok{ }\KeywordTok{readRDS}\NormalTok{(}\StringTok{"outputs/final/final_df_current_density.RDS"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{timestep =}\NormalTok{ (timestep}\OperatorTok{*}\DecValTok{10}\NormalTok{)}\OperatorTok{+}\DecValTok{1990}\NormalTok{, }\DataTypeTok{source =} \StringTok{"model"}\NormalTok{)}
\NormalTok{df_final_origin <-}\StringTok{ }\KeywordTok{readRDS}\NormalTok{(}\StringTok{"outputs/final/final_df_origin_current_density.RDS"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{timestep =}\NormalTok{ timestep}\OperatorTok{*}\DecValTok{10}\OperatorTok{+}\DecValTok{1990}\NormalTok{, }\DataTypeTok{source =} \StringTok{"model"}\NormalTok{)}

\CommentTok{# Sum all municipalities}
\NormalTok{df_summarised <-}\StringTok{ }\NormalTok{df_final }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(sce, timestep, species, iteration) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{summarise}\NormalTok{(}\DataTypeTok{sum_cur =} \KeywordTok{sum}\NormalTok{(current)) }\OperatorTok{%>%}\StringTok{ }\NormalTok{ungroup }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{source =} \StringTok{"model"}\NormalTok{)}
\NormalTok{df_origin_summarised <-}\StringTok{ }\NormalTok{df_final_origin }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(timestep, species) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{summarise}\NormalTok{(}\DataTypeTok{sum_cur =} \KeywordTok{sum}\NormalTok{(mean)) }\OperatorTok{%>%}\StringTok{ }\NormalTok{ungroup }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sce =} \StringTok{"sce_0"}\NormalTok{, }\DataTypeTok{source =} \StringTok{"observation"}\NormalTok{)}

\CommentTok{# Joined dataset}
\NormalTok{joined <-}\StringTok{ }\KeywordTok{full_join}\NormalTok{(df_summarised, df_origin_summarised, }
                    \DataTypeTok{by=}\KeywordTok{c}\NormalTok{(}\StringTok{"sce"}\NormalTok{, }\StringTok{"source"}\NormalTok{, }\StringTok{"species"}\NormalTok{, }\StringTok{"timestep"}\NormalTok{, }\StringTok{"sum_cur"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{left_join}\NormalTok{(results_clean, }\DataTypeTok{by =} \StringTok{"sce"}\NormalTok{)  }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{replace_na}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DataTypeTok{climate =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{run =} \StringTok{"historic run"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\CommentTok{# Make factors}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sce =} \KeywordTok{as.factor}\NormalTok{(sce), }\DataTypeTok{run =} \KeywordTok{as.factor}\NormalTok{(run), }\DataTypeTok{sum_cur =} \DecValTok{10}\OperatorTok{*}\NormalTok{(sum_cur),}
         \DataTypeTok{climate =} \KeywordTok{factor}\NormalTok{(climate, }\DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{"none"}\NormalTok{, }\StringTok{"historic"}\NormalTok{,}
                                              \StringTok{"baseline"}\NormalTok{, }\StringTok{"RCP 8.5"}\NormalTok{)),}
         \DataTypeTok{run =} \KeywordTok{factor}\NormalTok{(run, }\DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{"historic run"}\NormalTok{, }\StringTok{"BAU run"}\NormalTok{, }
                                      \StringTok{"BAU run + corrs protection"}\NormalTok{, }\StringTok{"BAU run + ref"}\NormalTok{, }
                                      \StringTok{"BAU run + corrs protection + ref"}\NormalTok{,}
                                      \StringTok{"BAU run + corrs protection + ref TARGETED"}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\hypertarget{set-the-ggplot-theme}{%
\paragraph{Set the ggplot theme}\label{set-the-ggplot-theme}}

We see the ggplot theme to \texttt{theme\_minimal()} and change the size
of some plot elements.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{theme_set}\NormalTok{(}\KeywordTok{theme_minimal}\NormalTok{())}
\KeywordTok{theme_update}\NormalTok{(}\DataTypeTok{legend.position =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.85}\NormalTok{, }\FloatTok{0.22}\NormalTok{),}
             \CommentTok{#legend.justification = c(1, -0.2), # for inside plot}
             \CommentTok{#legend.position = "none",}
             \DataTypeTok{legend.box =} \StringTok{"vertical"}\NormalTok{,}
             \DataTypeTok{legend.background =} \KeywordTok{element_blank}\NormalTok{(),}
             \DataTypeTok{legend.box.background =} \KeywordTok{element_rect}\NormalTok{(}\DataTypeTok{colour =} \StringTok{"black"}\NormalTok{),}
             \DataTypeTok{panel.border =} \KeywordTok{element_rect}\NormalTok{(}\DataTypeTok{fill =} \OtherTok{NA}\NormalTok{),}
             \DataTypeTok{legend.title =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{15}\NormalTok{),}
             \DataTypeTok{legend.text =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{12}\NormalTok{),}
             \DataTypeTok{plot.title =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size=}\DecValTok{22}\NormalTok{),}
             \DataTypeTok{plot.subtitle =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size=}\DecValTok{22}\NormalTok{),}
             \DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle=}\DecValTok{65}\NormalTok{, }\DataTypeTok{vjust=}\FloatTok{0.6}\NormalTok{, }\DataTypeTok{size =}\DecValTok{15}\NormalTok{),}
             \DataTypeTok{axis.text.y =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =}\DecValTok{15}\NormalTok{),}
             \DataTypeTok{strip.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{15}\NormalTok{),}
             \DataTypeTok{strip.text.y =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{15}\NormalTok{),}
             \DataTypeTok{axis.title=}\KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size=}\DecValTok{18}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{figures}{%
\subsubsection{Figures}\label{figures}}

\hypertarget{mean-current-flow-change-through-time---historic}{%
\paragraph{Mean current flow change through time -
historic}\label{mean-current-flow-change-through-time---historic}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fig_}\DecValTok{1}\NormalTok{_historic <-}\StringTok{ }\NormalTok{joined }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(climate }\OperatorTok{==}\StringTok{ "none"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(scenarioId, species, iteration) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_modify}\NormalTok{(}\OperatorTok{~}\KeywordTok{mutate}\NormalTok{(.x , }\DataTypeTok{baseline =} \KeywordTok{subset}\NormalTok{(.x, timestep }\OperatorTok{==}\StringTok{ }\DecValTok{1990}\NormalTok{)}\OperatorTok{$}\NormalTok{sum_cur, }
                       \DataTypeTok{the_diff =}\NormalTok{ ((sum_cur}\OperatorTok{-}\NormalTok{baseline)}\OperatorTok{/}\NormalTok{baseline)}\OperatorTok{*}\DecValTok{100}\NormalTok{ )) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{timestep, }\DataTypeTok{y=}\NormalTok{the_diff, }\DataTypeTok{col=}\NormalTok{source)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values=}\KeywordTok{c}\NormalTok{(}\StringTok{'#d8b365'}\NormalTok{,}\StringTok{'#5ab4ac'}\NormalTok{), }
                     \DataTypeTok{labels =} \KeywordTok{c}\NormalTok{(}\StringTok{"Model"}\NormalTok{, }\StringTok{"Observation"}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_smooth}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{group =}\NormalTok{ sce), }\DataTypeTok{method =} \StringTok{"glm"}\NormalTok{, }\DataTypeTok{se=}\OtherTok{FALSE}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_linetype_manual}\NormalTok{(}\DataTypeTok{values =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{3}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{5}\NormalTok{))}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{group =}\NormalTok{ sce)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{species, }\DataTypeTok{scales =} \StringTok{"free"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{y =} \StringTok{"Current flow (% Change)"}\NormalTok{,}
       \DataTypeTok{x =} \StringTok{"Year"}\NormalTok{,}
       \DataTypeTok{col =} \StringTok{"Source"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\OtherTok{NULL}
\NormalTok{fig_}\DecValTok{1}\NormalTok{_historic}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\includegraphics[width=1.5\linewidth]{msc_thesis_figures_files/figure-latex/fig_1_historic-1} \caption{Connectivity change for species through time, 1990-2010}\label{fig:fig_1_historic}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggsave}\NormalTok{(fig_}\DecValTok{1}\NormalTok{_historic,}
       \DataTypeTok{filename =} \StringTok{"outputs/figures/connectivity_decrease_x5species_historic.png"}\NormalTok{,}
       \DataTypeTok{width =} \DecValTok{17}\NormalTok{, }\DataTypeTok{height =} \DecValTok{12}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{mean-current-flow-change-through-time---chap-1}{%
\subparagraph{Mean current flow change through time - chap
1}\label{mean-current-flow-change-through-time---chap-1}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fig_}\DecValTok{1}\NormalTok{_static <-}\StringTok{ }\NormalTok{joined }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(climate }\OperatorTok{!=}\StringTok{ "none"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(scenarioId, species, iteration) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_modify}\NormalTok{(}\OperatorTok{~}\KeywordTok{mutate}\NormalTok{(.x , }\DataTypeTok{baseline =} \KeywordTok{subset}\NormalTok{(.x, timestep }\OperatorTok{==}\StringTok{ }\DecValTok{2010}\NormalTok{)}\OperatorTok{$}\NormalTok{sum_cur, }
                       \DataTypeTok{the_diff =}\NormalTok{ ((sum_cur}\OperatorTok{-}\NormalTok{baseline)}\OperatorTok{/}\NormalTok{baseline)}\OperatorTok{*}\DecValTok{100}\NormalTok{ )) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{timestep, }\DataTypeTok{y=}\NormalTok{the_diff, }\DataTypeTok{col=}\NormalTok{climate)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values=}\KeywordTok{c}\NormalTok{(}\StringTok{'#ffe300'}\NormalTok{, }\StringTok{'#ff7a14'}\NormalTok{, }\StringTok{"#b31400"}\NormalTok{), }
                     \DataTypeTok{labels =} \KeywordTok{c}\NormalTok{(}\StringTok{"Historic"}\NormalTok{, }\StringTok{"Baseline"}\NormalTok{, }\StringTok{"RCP 8.5"}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_smooth}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{group =}\NormalTok{ sce, }\DataTypeTok{linetype=}\NormalTok{run), }\DataTypeTok{alpha=}\FloatTok{0.2}\NormalTok{, }\DataTypeTok{method=}\StringTok{"loess"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_linetype_discrete}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{species, }\DataTypeTok{scales =} \StringTok{"fixed"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_hline}\NormalTok{(}\DataTypeTok{yintercept =} \DecValTok{0}\NormalTok{, }\DataTypeTok{linetype =} \StringTok{"dashed"}\NormalTok{, }\DataTypeTok{alpha =} \FloatTok{0.5}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{xintercept =} \DecValTok{2020}\NormalTok{, }\DataTypeTok{linetype =} \StringTok{"dashed"}\NormalTok{, }\DataTypeTok{color =} \StringTok{"darkGreen"}\NormalTok{, }\DataTypeTok{alpha =} \FloatTok{0.5}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{y =} \StringTok{"Current flow (% Change)"}\NormalTok{,}
       \DataTypeTok{x =} \StringTok{"Year"}\NormalTok{,}
       \DataTypeTok{col =} \StringTok{"Climate"}\NormalTok{,}
       \DataTypeTok{linetype =} \StringTok{"Run"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\OtherTok{NULL}
\NormalTok{fig_}\DecValTok{1}\NormalTok{_static}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using formula 'y ~ x'
\end{verbatim}

\begin{figure}
\includegraphics[width=1.5\linewidth]{msc_thesis_figures_files/figure-latex/figure_1_model-1} \caption{Connectivity change for species through time, 2010-2100}\label{fig:figure_1_model}
\end{figure}

\end{document}