### 2020-07-18 Rmd/msc_thesis_figures.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Msc Thesis Figures},
  pdfauthor={Valentin Lucet},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\title{Msc Thesis Figures}
\author{Valentin Lucet}
\date{17/07/2020}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{5}
\tableofcontents
}
\hypertarget{data-prep}{%
\subsubsection{Data prep}\label{data-prep}}

\hypertarget{cleaning-and-formatting}{%
\paragraph{Cleaning and formatting}\label{cleaning-and-formatting}}

We start by preparing the data from the outputs of scripts \texttt{7.1}
and \texttt{8.1}. The results table shown as output is the key to read
the figures.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Results dir + mun shapefile}
\NormalTok{mun <-}\StringTok{ }\KeywordTok{st_read}\NormalTok{(}\StringTok{"data/mun/munic_SHP_clean.shp"}\NormalTok{, }\DataTypeTok{quiet =} \OtherTok{TRUE}\NormalTok{)}

\CommentTok{# results key}
\NormalTok{sce_code_vec <-}\StringTok{ }
\StringTok{  }\KeywordTok{as.vector}\NormalTok{(}\KeywordTok{t}\NormalTok{(}\KeywordTok{outer}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"BAU-"}\NormalTok{, }\StringTok{"R-"}\NormalTok{, }\StringTok{"CorPr-"}\NormalTok{, }\StringTok{"R-CorPr-"}\NormalTok{, }\StringTok{"R(T)-CorPr-"}\NormalTok{), }
                    \KeywordTok{c}\NormalTok{(}\StringTok{"Hist"}\NormalTok{, }\StringTok{"Base"}\NormalTok{, }\StringTok{"RCP8"}\NormalTok{), paste0)))}
\NormalTok{sce_code_vec_run <-}\StringTok{ }
\StringTok{  }\KeywordTok{rep}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"BAU"}\NormalTok{, }\StringTok{"R"}\NormalTok{, }\StringTok{"CorPr"}\NormalTok{, }\StringTok{"R-CorPr"}\NormalTok{, }\StringTok{"R(T)-CorPr"}\NormalTok{), }\DataTypeTok{each=}\DecValTok{3}\NormalTok{)}

\NormalTok{results_clean <-}\StringTok{ }\KeywordTok{read_rds}\NormalTok{(}\StringTok{"data/temp/stsim_run_results.RDS"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(scenarioId, name) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{name =}  \KeywordTok{gsub}\NormalTok{(.}\OperatorTok{$}\NormalTok{name, }\DataTypeTok{pattern =} \StringTok{" }\CharTok{\textbackslash{}\textbackslash{}}\StringTok{(.+}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{)"}\NormalTok{, }\DataTypeTok{replacement =} \StringTok{""}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sce =}  \KeywordTok{paste0}\NormalTok{(}\StringTok{"sce_"}\NormalTok{, .}\OperatorTok{$}\NormalTok{scenarioId)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{chapter =} \KeywordTok{c}\NormalTok{(}\StringTok{"none"}\NormalTok{, }\KeywordTok{rep}\NormalTok{(}\StringTok{"both"}\NormalTok{, }\DecValTok{3}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\StringTok{"chap_1"}\NormalTok{, }\DecValTok{3}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\StringTok{"chap_2"}\NormalTok{, }\DecValTok{9}\NormalTok{))) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{splitted =} \KeywordTok{str_split}\NormalTok{(name, }\StringTok{" }\CharTok{\textbackslash{}\textbackslash{}}\StringTok{| "}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{climate =} \KeywordTok{unlist}\NormalTok{(}\KeywordTok{map}\NormalTok{(splitted, }\OperatorTok{~}\KeywordTok{unlist}\NormalTok{(.x[}\DecValTok{2}\NormalTok{])))) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{run =} \KeywordTok{unlist}\NormalTok{(}\KeywordTok{map}\NormalTok{(splitted, }\OperatorTok{~}\KeywordTok{unlist}\NormalTok{(.x[}\DecValTok{1}\NormalTok{])))) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{splitted) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{replace_na}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DataTypeTok{climate =} \StringTok{"none"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{code =} \KeywordTok{c}\NormalTok{(}\StringTok{"Control"}\NormalTok{, sce_code_vec)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{code_run =} \KeywordTok{c}\NormalTok{(}\StringTok{"control"}\NormalTok{, sce_code_vec_run))}

\CommentTok{# Final extracted datasets}
\NormalTok{df_final <-}\StringTok{ }\KeywordTok{readRDS}\NormalTok{(}\StringTok{"outputs/final/final_df_current_density.RDS"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{timestep =}\NormalTok{ (timestep}\OperatorTok{*}\DecValTok{10}\NormalTok{)}\OperatorTok{+}\DecValTok{1990}\NormalTok{, }\DataTypeTok{source =} \StringTok{"model"}\NormalTok{)}
\NormalTok{df_final_origin <-}\StringTok{ }\KeywordTok{readRDS}\NormalTok{(}\StringTok{"outputs/final/final_df_origin_current_density.RDS"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{timestep =}\NormalTok{ timestep}\OperatorTok{*}\DecValTok{10}\OperatorTok{+}\DecValTok{1990}\NormalTok{, }\DataTypeTok{source =} \StringTok{"model"}\NormalTok{)}

\CommentTok{# Stratum key}
\NormalTok{stratum_key <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"config/stsim/SecondaryStratum.csv"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{ID =} \KeywordTok{as.factor}\NormalTok{(ID)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{zone=}\NormalTok{ID, }\DataTypeTok{MUS_NM_MUN=}\NormalTok{Name) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{left_join}\NormalTok{(df_final, }\DataTypeTok{by=}\StringTok{"zone"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(MUS_NM_MUN}\OperatorTok{!=}\StringTok{"Not Monteregie"}\NormalTok{)}

\CommentTok{# Sum all municipalities}
\NormalTok{df_summarised <-}\StringTok{ }\NormalTok{df_final }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(sce, timestep, species, iteration) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{summarise}\NormalTok{(}\DataTypeTok{sum_cur =} \KeywordTok{sum}\NormalTok{(current)) }\OperatorTok{%>%}\StringTok{ }\NormalTok{ungroup }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{source =} \StringTok{"model"}\NormalTok{)}
\NormalTok{df_origin_summarised <-}\StringTok{ }\NormalTok{df_final_origin }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(timestep, species) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{summarise}\NormalTok{(}\DataTypeTok{sum_cur =} \KeywordTok{sum}\NormalTok{(mean)) }\OperatorTok{%>%}\StringTok{ }\NormalTok{ungroup }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sce =} \StringTok{"sce_0"}\NormalTok{, }\DataTypeTok{source =} \StringTok{"observation"}\NormalTok{)}

\CommentTok{# Joined dataset}
\NormalTok{joined <-}\StringTok{ }\KeywordTok{full_join}\NormalTok{(df_summarised, df_origin_summarised, }
                    \DataTypeTok{by=}\KeywordTok{c}\NormalTok{(}\StringTok{"sce"}\NormalTok{, }\StringTok{"source"}\NormalTok{, }\StringTok{"species"}\NormalTok{, }\StringTok{"timestep"}\NormalTok{, }\StringTok{"sum_cur"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{left_join}\NormalTok{(results_clean, }\DataTypeTok{by =} \StringTok{"sce"}\NormalTok{)  }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{replace_na}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DataTypeTok{climate =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{run =} \StringTok{"historic run"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\CommentTok{# Make factors}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sce =} \KeywordTok{as.factor}\NormalTok{(sce), }\DataTypeTok{run =} \KeywordTok{as.factor}\NormalTok{(run), }\DataTypeTok{sum_cur =} \DecValTok{10}\OperatorTok{*}\NormalTok{(sum_cur),}
         \DataTypeTok{climate =} \KeywordTok{factor}\NormalTok{(climate, }\DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{"none"}\NormalTok{, }\StringTok{"historic"}\NormalTok{,}
                                              \StringTok{"baseline"}\NormalTok{, }\StringTok{"RCP 8.5"}\NormalTok{)),}
         \DataTypeTok{run =} \KeywordTok{factor}\NormalTok{(run, }\DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{"historic run"}\NormalTok{, }\StringTok{"BAU run"}\NormalTok{, }
                                      \StringTok{"BAU run + corrs protection"}\NormalTok{, }\StringTok{"BAU run + ref"}\NormalTok{, }
                                      \StringTok{"BAU run + corrs protection + ref"}\NormalTok{,}
                                      \StringTok{"BAU run + corrs protection + ref TARGETED"}\NormalTok{)))}

\CommentTok{# Roc curves data}
\NormalTok{urb <-}\StringTok{ }\KeywordTok{readRDS}\NormalTok{(}\StringTok{"data/temp/fit_rs_outcome_rf_urb_2.RDS"}\NormalTok{)}
\NormalTok{agex <-}\StringTok{ }\KeywordTok{readRDS}\NormalTok{(}\StringTok{"data/temp/fit_rs_outcome_rf_agex_2.RDS"}\NormalTok{)}

\CommentTok{# Histogram data }
\NormalTok{histograms <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"outputs/final/final_values_output.csv"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{left_join}\NormalTok{(results_clean, }\DataTypeTok{by=}\StringTok{"sce"}\NormalTok{)}

\CommentTok{# SURF Data}
\NormalTok{surf <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"surf/surf_output.csv"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{timestep =}\NormalTok{ timestep}\OperatorTok{*}\DecValTok{10}\OperatorTok{+}\DecValTok{1990}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{left_join}\NormalTok{(results_clean, }\DataTypeTok{by=}\KeywordTok{c}\NormalTok{(}\StringTok{"scenario"}\NormalTok{=}\StringTok{"scenarioId"}\NormalTok{))}

\NormalTok{results_clean}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##    scenarioId                                                 name    sce
## 1          37                                         historic run sce_37
## 2          38                                   BAU run | historic sce_38
## 3          39                                   BAU run | baseline sce_39
## 4          40                                    BAU run | RCP 8.5 sce_40
## 5          41                             BAU run + ref | historic sce_41
## 6          42                             BAU run + ref | baseline sce_42
## 7          43                              BAU run + ref | RCP 8.5 sce_43
## 8          44                BAU run + corrs protection | historic sce_44
## 9          45                BAU run + corrs protection | baseline sce_45
## 10         46                 BAU run + corrs protection | RCP 8.5 sce_46
## 11         47          BAU run + corrs protection + ref | historic sce_47
## 12         48          BAU run + corrs protection + ref | baseline sce_48
## 13         49           BAU run + corrs protection + ref | RCP 8.5 sce_49
## 14         50 BAU run + corrs protection + ref TARGETED | historic sce_50
## 15         51 BAU run + corrs protection + ref TARGETED | baseline sce_51
## 16         52  BAU run + corrs protection + ref TARGETED | RCP 8.5 sce_52
##    chapter  climate                                       run            code
## 1     none     none                              historic run         Control
## 2     both historic                                   BAU run        BAU-Hist
## 3     both baseline                                   BAU run        BAU-Base
## 4     both  RCP 8.5                                   BAU run        BAU-RCP8
## 5   chap_1 historic                             BAU run + ref          R-Hist
## 6   chap_1 baseline                             BAU run + ref          R-Base
## 7   chap_1  RCP 8.5                             BAU run + ref          R-RCP8
## 8   chap_2 historic                BAU run + corrs protection      CorPr-Hist
## 9   chap_2 baseline                BAU run + corrs protection      CorPr-Base
## 10  chap_2  RCP 8.5                BAU run + corrs protection      CorPr-RCP8
## 11  chap_2 historic          BAU run + corrs protection + ref    R-CorPr-Hist
## 12  chap_2 baseline          BAU run + corrs protection + ref    R-CorPr-Base
## 13  chap_2  RCP 8.5          BAU run + corrs protection + ref    R-CorPr-RCP8
## 14  chap_2 historic BAU run + corrs protection + ref TARGETED R(T)-CorPr-Hist
## 15  chap_2 baseline BAU run + corrs protection + ref TARGETED R(T)-CorPr-Base
## 16  chap_2  RCP 8.5 BAU run + corrs protection + ref TARGETED R(T)-CorPr-RCP8
##      code_run
## 1     control
## 2         BAU
## 3         BAU
## 4         BAU
## 5           R
## 6           R
## 7           R
## 8       CorPr
## 9       CorPr
## 10      CorPr
## 11    R-CorPr
## 12    R-CorPr
## 13    R-CorPr
## 14 R(T)-CorPr
## 15 R(T)-CorPr
## 16 R(T)-CorPr
\end{verbatim}

Preparing for spatial plotting (time consuming).

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Pivoted dataset for mapping}
\CommentTok{# df_final_fordiff_pivoted <- df_final %>% }
\CommentTok{#   left_join(results_clean, by = "sce")  %>% }
\CommentTok{#   group_by(zone, sce, species, timestep, source, chapter) %>% }
\CommentTok{#   summarise(current = mean(current)) %>% ungroup %>% }
\CommentTok{#   filter(timestep %in% c(2010, last(unique(stratum_key$timestep)))) %>% }
\CommentTok{#   pivot_wider(names_from=timestep, values_from=current) %>% drop_na() %>% }
\CommentTok{#   rename(before=last_col(offset = 1), after=last_col()) %>% }
\CommentTok{#   left_join(stratum_key, by=c("zone", "sce", "species", "source")) %>% }
\CommentTok{#   filter(MUS_NM_MUN!="Not Monteregie") %>% }
\CommentTok{#   mutate(change=((after-before)/before)*100) %>% }
\CommentTok{#   left_join(mun, by="MUS_NM_MUN") %>% }
\CommentTok{#   st_as_sf()}
\CommentTok{# }
\CommentTok{# df_final_fordiff_pivoted <- }
\CommentTok{#   st_transform(df_final_fordiff_pivoted, crs=4326)}
\end{Highlighting}
\end{Shaded}

\hypertarget{set-the-ggplot-theme}{%
\paragraph{Set the ggplot theme}\label{set-the-ggplot-theme}}

We see the ggplot theme to \texttt{theme\_minimal()} and change the size
of some plot elements.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{theme_set}\NormalTok{(}\KeywordTok{theme_minimal}\NormalTok{())}
\KeywordTok{theme_update}\NormalTok{(}\DataTypeTok{legend.position =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.85}\NormalTok{, }\FloatTok{0.20}\NormalTok{),}
             \CommentTok{#legend.justification = c(1, -0.2), # for inside plot}
             \CommentTok{#legend.position = "none",}
             \DataTypeTok{legend.box =} \StringTok{"vertical"}\NormalTok{,}
             \DataTypeTok{legend.background =} \KeywordTok{element_blank}\NormalTok{(),}
             \DataTypeTok{legend.box.background =} \KeywordTok{element_rect}\NormalTok{(}\DataTypeTok{colour =} \StringTok{"black"}\NormalTok{),}
             \DataTypeTok{panel.border =} \KeywordTok{element_rect}\NormalTok{(}\DataTypeTok{fill =} \OtherTok{NA}\NormalTok{),}
             \DataTypeTok{legend.title =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{10}\NormalTok{),}
             \DataTypeTok{legend.text =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{8}\NormalTok{),}
             \DataTypeTok{plot.title =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size=}\DecValTok{20}\NormalTok{),}
             \DataTypeTok{plot.subtitle =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size=}\DecValTok{20}\NormalTok{),}
             \DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle=}\DecValTok{65}\NormalTok{, }\DataTypeTok{vjust=}\FloatTok{0.6}\NormalTok{, }\DataTypeTok{size =}\DecValTok{13}\NormalTok{),}
             \DataTypeTok{axis.text.y =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =}\DecValTok{10}\NormalTok{),}
             \DataTypeTok{strip.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{10}\NormalTok{),}
             \DataTypeTok{strip.text.y =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{10}\NormalTok{),}
             \DataTypeTok{axis.title=}\KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size=}\DecValTok{15}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{figures}{%
\subsubsection{Figures}\label{figures}}

\hypertarget{mean-current-flow-change}{%
\subsubsection{Mean current flow
change}\label{mean-current-flow-change}}

\hypertarget{historic}{%
\paragraph{Historic}\label{historic}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fig_}\DecValTok{1}\NormalTok{_historic <-}\StringTok{ }\NormalTok{joined }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(climate }\OperatorTok{==}\StringTok{ "none"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(scenarioId, species, iteration) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_modify}\NormalTok{(}\OperatorTok{~}\KeywordTok{mutate}\NormalTok{(.x , }\DataTypeTok{baseline =} \KeywordTok{subset}\NormalTok{(.x, timestep }\OperatorTok{==}\StringTok{ }\DecValTok{1990}\NormalTok{)}\OperatorTok{$}\NormalTok{sum_cur, }
                       \DataTypeTok{the_diff =}\NormalTok{ ((sum_cur}\OperatorTok{-}\NormalTok{baseline)}\OperatorTok{/}\NormalTok{baseline)}\OperatorTok{*}\DecValTok{100}\NormalTok{ )) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{timestep, }\DataTypeTok{y=}\NormalTok{the_diff, }\DataTypeTok{col=}\NormalTok{source)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values=}\KeywordTok{c}\NormalTok{(}\StringTok{'#d8b365'}\NormalTok{,}\StringTok{'#5ab4ac'}\NormalTok{), }
                     \DataTypeTok{labels =} \KeywordTok{c}\NormalTok{(}\StringTok{"Model"}\NormalTok{, }\StringTok{"Observation"}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_smooth}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{group =}\NormalTok{ sce), }\DataTypeTok{method =} \StringTok{"glm"}\NormalTok{, }\DataTypeTok{se=}\OtherTok{FALSE}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_linetype_manual}\NormalTok{(}\DataTypeTok{values =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{3}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{5}\NormalTok{))}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{group =}\NormalTok{ sce)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{species, }\DataTypeTok{scales =} \StringTok{"free"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{y =} \StringTok{"Current flow (% Change)"}\NormalTok{,}
       \DataTypeTok{x =} \StringTok{"Year"}\NormalTok{,}
       \DataTypeTok{col =} \StringTok{"Source"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\OtherTok{NULL}
\NormalTok{fig_}\DecValTok{1}\NormalTok{_historic}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\includegraphics[width=1\linewidth]{msc_thesis_figures_files/figure-latex/fig_1_historic-1} \caption{Connectivity change for species through time, 1990-2010}\label{fig:fig_1_historic}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggsave}\NormalTok{(fig_}\DecValTok{1}\NormalTok{_historic,}
       \DataTypeTok{filename =} \StringTok{"outputs/figures/connectivity_decrease_x5species_historic.png"}\NormalTok{,}
       \DataTypeTok{width =} \DecValTok{15}\NormalTok{, }\DataTypeTok{height =} \DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{chap-1}{%
\paragraph{Chap 1}\label{chap-1}}

\hypertarget{line-graph}{%
\subparagraph{Line graph}\label{line-graph}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fig_}\DecValTok{1}\NormalTok{_static_chap_}\DecValTok{1}\NormalTok{ <-}\StringTok{ }\NormalTok{joined }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(climate }\OperatorTok{!=}\StringTok{ "none"}\NormalTok{, chapter }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"none"}\NormalTok{, }\StringTok{"both"}\NormalTok{, }\StringTok{"chap_1"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(scenarioId, species, iteration) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_modify}\NormalTok{(}\OperatorTok{~}\KeywordTok{mutate}\NormalTok{(.x , }\DataTypeTok{baseline =} \KeywordTok{subset}\NormalTok{(.x, timestep }\OperatorTok{==}\StringTok{ }\DecValTok{2010}\NormalTok{)}\OperatorTok{$}\NormalTok{sum_cur, }
                       \DataTypeTok{the_diff =}\NormalTok{ ((sum_cur}\OperatorTok{-}\NormalTok{baseline)}\OperatorTok{/}\NormalTok{baseline)}\OperatorTok{*}\DecValTok{100}\NormalTok{ )) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{timestep, }\DataTypeTok{y=}\NormalTok{the_diff, }\DataTypeTok{col=}\NormalTok{climate)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values=}\KeywordTok{c}\NormalTok{(}\StringTok{'#ffe300'}\NormalTok{, }\StringTok{'#ff7a14'}\NormalTok{, }\StringTok{"#b31400"}\NormalTok{), }
                     \DataTypeTok{labels =} \KeywordTok{c}\NormalTok{(}\StringTok{"Historic"}\NormalTok{, }\StringTok{"Baseline"}\NormalTok{, }\StringTok{"RCP 8.5"}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_smooth}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{group =}\NormalTok{ sce, }\DataTypeTok{linetype=}\NormalTok{code_run), }\DataTypeTok{alpha=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{method=}\StringTok{"loess"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_linetype_manual}\NormalTok{(}\DataTypeTok{values=}\KeywordTok{c}\NormalTok{(}\StringTok{"solid"}\NormalTok{, }\StringTok{"twodash"}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{species, }\DataTypeTok{scales =} \StringTok{"fixed"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_hline}\NormalTok{(}\DataTypeTok{yintercept =} \DecValTok{0}\NormalTok{, }\DataTypeTok{linetype =} \StringTok{"dashed"}\NormalTok{, }\DataTypeTok{alpha =} \FloatTok{0.5}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{xintercept =} \DecValTok{2020}\NormalTok{, }\DataTypeTok{linetype =} \StringTok{"dashed"}\NormalTok{, }\DataTypeTok{color =} \StringTok{"darkGreen"}\NormalTok{, }\DataTypeTok{alpha =} \FloatTok{0.5}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{y =} \StringTok{"Current flow (% Change)"}\NormalTok{,}
       \DataTypeTok{x =} \StringTok{"Year"}\NormalTok{,}
       \DataTypeTok{col =} \StringTok{"Climate"}\NormalTok{,}
       \DataTypeTok{linetype =} \StringTok{"Run"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{col =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{ncol =} \DecValTok{2}\NormalTok{, }\DataTypeTok{nrow=}\DecValTok{2}\NormalTok{), }
         \DataTypeTok{linetype =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{nrow=}\DecValTok{1}\NormalTok{, }\DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{colour =} \StringTok{'black'}\NormalTok{))) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.85}\NormalTok{, }\FloatTok{0.15}\NormalTok{))}\OperatorTok{+}
\StringTok{  }\OtherTok{NULL}
\NormalTok{fig_}\DecValTok{1}\NormalTok{_static_chap_}\DecValTok{1}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\includegraphics[width=1\linewidth]{msc_thesis_figures_files/figure-latex/figure_1_model_1-1} \caption{Connectivity change for species through time, 2010-2100}\label{fig:figure_1_model_1}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggsave}\NormalTok{(fig_}\DecValTok{1}\NormalTok{_static_chap_}\DecValTok{1}\NormalTok{, }
       \DataTypeTok{filename =} \StringTok{"outputs/figures/connectivity_decrease_x5species_chap1.png"}\NormalTok{, }
       \DataTypeTok{width =} \DecValTok{15}\NormalTok{, }\DataTypeTok{height =} \DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{radar-graph}{%
\subparagraph{Radar Graph}\label{radar-graph}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{radar_data_chap1 <-}\StringTok{ }\NormalTok{joined }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(climate }\OperatorTok{!=}\StringTok{ "none"}\NormalTok{, chapter }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"none"}\NormalTok{, }\StringTok{"both"}\NormalTok{, }\StringTok{"chap_1"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(timestep, species, code) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{summarise}\NormalTok{(}\DataTypeTok{sum_cur =} \KeywordTok{mean}\NormalTok{(sum_cur)) }\OperatorTok{%>%}\StringTok{ }\NormalTok{ungroup }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{pivot_wider}\NormalTok{(}\DataTypeTok{names_from =}\NormalTok{ timestep, }\DataTypeTok{values_from =}\NormalTok{ sum_cur) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{diff =}\NormalTok{ ((}\StringTok{`}\DataTypeTok{2100}\StringTok{`}\OperatorTok{-}\StringTok{`}\DataTypeTok{2010}\StringTok{`}\NormalTok{)}\OperatorTok{/}\StringTok{`}\DataTypeTok{2010}\StringTok{`}\NormalTok{)}\OperatorTok{*}\DecValTok{100}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(}\StringTok{`}\DataTypeTok{2010}\StringTok{`}\OperatorTok{:}\StringTok{`}\DataTypeTok{2100}\StringTok{`}\NormalTok{)) }\OperatorTok{%>%}\StringTok{  }
\StringTok{  }\KeywordTok{pivot_wider}\NormalTok{(}\DataTypeTok{names_from =}\NormalTok{ code, }\DataTypeTok{values_from =}\NormalTok{ diff) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{group =}\NormalTok{ species)}

\NormalTok{radar_chap1 <-}\StringTok{  }
\StringTok{  }\KeywordTok{ggradar}\NormalTok{(radar_data_chap1, }\DataTypeTok{centre.y =} \DecValTok{-20}\NormalTok{, }\DataTypeTok{legend.position =} \StringTok{"right"}\NormalTok{,}
          \DataTypeTok{grid.min =} \DecValTok{-20}\NormalTok{, }\DataTypeTok{grid.max =} \DecValTok{3}\NormalTok{, }\DataTypeTok{grid.mid =} \DecValTok{0}\NormalTok{, }
          \DataTypeTok{values.radar =} \KeywordTok{c}\NormalTok{(}\StringTok{"-20 %"}\NormalTok{, }\StringTok{"0 %"}\NormalTok{, }\StringTok{""}\NormalTok{), }
          \DataTypeTok{gridline.label.offset =}\DecValTok{5}\NormalTok{, }\DataTypeTok{legend.text.size=} \DecValTok{12}\NormalTok{,}
          \DataTypeTok{group.line.width =}\FloatTok{0.8}\NormalTok{,}
          \DataTypeTok{background.circle.colour  =} \StringTok{"#ffffff"}\NormalTok{, }\DataTypeTok{group.point.size   =} \DecValTok{2}\NormalTok{, }
          \DataTypeTok{axis.label.size   =}\DecValTok{3}\NormalTok{)}
\NormalTok{radar_chap1}
\end{Highlighting}
\end{Shaded}

\includegraphics[width=1\linewidth]{msc_thesis_figures_files/figure-latex/figure_1_model_1_radar-1}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggsave}\NormalTok{(}\StringTok{"outputs/figures/radar_ggradar_chap1.png"}\NormalTok{, radar_chap1, }\DataTypeTok{width =} \DecValTok{15}\NormalTok{, }\DataTypeTok{height =} \DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{chap-2}{%
\paragraph{Chap 2}\label{chap-2}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fig_}\DecValTok{1}\NormalTok{_static_chap_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\NormalTok{joined }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(climate }\OperatorTok{!=}\StringTok{ "none"}\NormalTok{, chapter }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"none"}\NormalTok{, }\StringTok{"both"}\NormalTok{, }\StringTok{"chap_2"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(scenarioId, species, iteration) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_modify}\NormalTok{(}\OperatorTok{~}\KeywordTok{mutate}\NormalTok{(.x , }\DataTypeTok{baseline =} \KeywordTok{subset}\NormalTok{(.x, timestep }\OperatorTok{==}\StringTok{ }\DecValTok{2010}\NormalTok{)}\OperatorTok{$}\NormalTok{sum_cur, }
                       \DataTypeTok{the_diff =}\NormalTok{ ((sum_cur}\OperatorTok{-}\NormalTok{baseline)}\OperatorTok{/}\NormalTok{baseline)}\OperatorTok{*}\DecValTok{100}\NormalTok{ )) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{timestep, }\DataTypeTok{y=}\NormalTok{the_diff, }\DataTypeTok{col=}\NormalTok{climate)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values=}\KeywordTok{c}\NormalTok{(}\StringTok{'#ffe300'}\NormalTok{, }\StringTok{'#ff7a14'}\NormalTok{, }\StringTok{"#b31400"}\NormalTok{), }
                     \DataTypeTok{labels =} \KeywordTok{c}\NormalTok{(}\StringTok{"Historic"}\NormalTok{, }\StringTok{"Baseline"}\NormalTok{, }\StringTok{"RCP 8.5"}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_smooth}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{group =}\NormalTok{ sce, }\DataTypeTok{linetype=}\NormalTok{code_run), }\DataTypeTok{alpha=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{method=}\StringTok{"loess"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_linetype_manual}\NormalTok{(}\DataTypeTok{values=}\KeywordTok{c}\NormalTok{(}\StringTok{"solid"}\NormalTok{, }\StringTok{"dashed"}\NormalTok{, }\StringTok{"longdash"}\NormalTok{, }\StringTok{"dotted"}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{species, }\DataTypeTok{scales =} \StringTok{"fixed"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_hline}\NormalTok{(}\DataTypeTok{yintercept =} \DecValTok{0}\NormalTok{, }\DataTypeTok{linetype =} \StringTok{"dashed"}\NormalTok{, }\DataTypeTok{alpha =} \FloatTok{0.5}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{xintercept =} \DecValTok{2020}\NormalTok{, }\DataTypeTok{linetype =} \StringTok{"dashed"}\NormalTok{, }\DataTypeTok{color =} \StringTok{"darkGreen"}\NormalTok{, }\DataTypeTok{alpha =} \FloatTok{0.5}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{y =} \StringTok{"Current flow (% Change)"}\NormalTok{,}
       \DataTypeTok{x =} \StringTok{"Year"}\NormalTok{,}
       \DataTypeTok{col =} \StringTok{"Climate"}\NormalTok{,}
       \DataTypeTok{linetype =} \StringTok{"Run"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{col =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{ncol =} \DecValTok{2}\NormalTok{, }\DataTypeTok{nrow=}\DecValTok{2}\NormalTok{), }
         \DataTypeTok{linetype =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{nrow =} \DecValTok{2}\NormalTok{, }\DataTypeTok{ncol=}\DecValTok{2}\NormalTok{, }
                                 \DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{colour =} \StringTok{'black'}\NormalTok{))) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.85}\NormalTok{, }\FloatTok{0.15}\NormalTok{))}\OperatorTok{+}
\StringTok{  }\OtherTok{NULL}
\NormalTok{fig_}\DecValTok{1}\NormalTok{_static_chap_}\DecValTok{2}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\includegraphics[width=1\linewidth]{msc_thesis_figures_files/figure-latex/figure_1_model_2-1} \caption{Connectivity change for species through time, 2010-2100}\label{fig:figure_1_model_2}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggsave}\NormalTok{(fig_}\DecValTok{1}\NormalTok{_static_chap_}\DecValTok{2}\NormalTok{, }
       \DataTypeTok{filename =} \StringTok{"outputs/figures/connectivity_decrease_x5species_chap2.png"}\NormalTok{, }
       \DataTypeTok{width =} \DecValTok{15}\NormalTok{, }\DataTypeTok{height =} \DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{radar-graph-1}{%
\subparagraph{Radar Graph}\label{radar-graph-1}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{radar_data_chap2 <-}\StringTok{ }\NormalTok{joined }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(climate }\OperatorTok{!=}\StringTok{ "none"}\NormalTok{, chapter }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"none"}\NormalTok{, }\StringTok{"both"}\NormalTok{,}\StringTok{"chap_2"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(timestep, species, code) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{summarise}\NormalTok{(}\DataTypeTok{sum_cur =} \KeywordTok{mean}\NormalTok{(sum_cur)) }\OperatorTok{%>%}\StringTok{ }\NormalTok{ungroup }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{pivot_wider}\NormalTok{(}\DataTypeTok{names_from =}\NormalTok{ timestep, }\DataTypeTok{values_from =}\NormalTok{ sum_cur) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{diff =}\NormalTok{ ((}\StringTok{`}\DataTypeTok{2100}\StringTok{`}\OperatorTok{-}\StringTok{`}\DataTypeTok{2010}\StringTok{`}\NormalTok{)}\OperatorTok{/}\StringTok{`}\DataTypeTok{2010}\StringTok{`}\NormalTok{)}\OperatorTok{*}\DecValTok{100}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(}\StringTok{`}\DataTypeTok{2010}\StringTok{`}\OperatorTok{:}\StringTok{`}\DataTypeTok{2100}\StringTok{`}\NormalTok{)) }\OperatorTok{%>%}\StringTok{  }
\StringTok{  }\KeywordTok{pivot_wider}\NormalTok{(}\DataTypeTok{names_from =}\NormalTok{ code, }\DataTypeTok{values_from =}\NormalTok{ diff) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{group =}\NormalTok{ species)}

\NormalTok{radar_chap2 <-}\StringTok{  }
\StringTok{  }\KeywordTok{ggradar}\NormalTok{(radar_data_chap2, }\DataTypeTok{centre.y =} \DecValTok{-20}\NormalTok{, }\DataTypeTok{legend.position =} \StringTok{"right"}\NormalTok{,}
          \DataTypeTok{grid.min =} \DecValTok{-20}\NormalTok{, }\DataTypeTok{grid.max =} \DecValTok{3}\NormalTok{, }\DataTypeTok{grid.mid =} \DecValTok{0}\NormalTok{, }
          \DataTypeTok{values.radar =} \KeywordTok{c}\NormalTok{(}\StringTok{"-20 %"}\NormalTok{, }\StringTok{"0 %"}\NormalTok{, }\StringTok{""}\NormalTok{), }
          \DataTypeTok{gridline.label.offset =}\DecValTok{5}\NormalTok{, }\DataTypeTok{legend.text.size=} \DecValTok{12}\NormalTok{,}
          \DataTypeTok{group.line.width =}\FloatTok{0.8}\NormalTok{,}
          \DataTypeTok{background.circle.colour  =} \StringTok{"#ffffff"}\NormalTok{, }\DataTypeTok{group.point.size   =} \DecValTok{2}\NormalTok{, }
          \DataTypeTok{axis.label.size   =}\DecValTok{3}\NormalTok{)}
\NormalTok{radar_chap2}
\end{Highlighting}
\end{Shaded}

\includegraphics[width=1\linewidth]{msc_thesis_figures_files/figure-latex/figure_1_model_2_radar-1}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggsave}\NormalTok{(}\StringTok{"outputs/figures/radar_ggradar_chap2.png"}\NormalTok{, radar_chap2, }\DataTypeTok{width =} \DecValTok{15}\NormalTok{, }\DataTypeTok{height =} \DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{both}{%
\paragraph{Both}\label{both}}

\hypertarget{radar-graph-2}{%
\subparagraph{Radar Graph}\label{radar-graph-2}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{radar_data_all <-}\StringTok{ }\NormalTok{joined }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(climate }\OperatorTok{!=}\StringTok{ "none"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(timestep, species, code) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{summarise}\NormalTok{(}\DataTypeTok{sum_cur =} \KeywordTok{mean}\NormalTok{(sum_cur)) }\OperatorTok{%>%}\StringTok{ }\NormalTok{ungroup }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{pivot_wider}\NormalTok{(}\DataTypeTok{names_from =}\NormalTok{ timestep, }\DataTypeTok{values_from =}\NormalTok{ sum_cur) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{diff =}\NormalTok{ ((}\StringTok{`}\DataTypeTok{2100}\StringTok{`}\OperatorTok{-}\StringTok{`}\DataTypeTok{2010}\StringTok{`}\NormalTok{)}\OperatorTok{/}\StringTok{`}\DataTypeTok{2010}\StringTok{`}\NormalTok{)}\OperatorTok{*}\DecValTok{100}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(}\StringTok{`}\DataTypeTok{2010}\StringTok{`}\OperatorTok{:}\StringTok{`}\DataTypeTok{2100}\StringTok{`}\NormalTok{)) }\OperatorTok{%>%}\StringTok{  }
\StringTok{  }\KeywordTok{pivot_wider}\NormalTok{(}\DataTypeTok{names_from =}\NormalTok{ code, }\DataTypeTok{values_from =}\NormalTok{ diff) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{group =}\NormalTok{ species)}

\NormalTok{radar_all <-}\StringTok{  }
\StringTok{  }\KeywordTok{ggradar}\NormalTok{(radar_data_all, }\DataTypeTok{centre.y =} \DecValTok{-20}\NormalTok{, }\DataTypeTok{legend.position =} \StringTok{"right"}\NormalTok{,}
          \DataTypeTok{grid.min =} \DecValTok{-20}\NormalTok{, }\DataTypeTok{grid.max =} \DecValTok{3}\NormalTok{, }\DataTypeTok{grid.mid =} \DecValTok{0}\NormalTok{, }
          \DataTypeTok{values.radar =} \KeywordTok{c}\NormalTok{(}\StringTok{"-20 %"}\NormalTok{, }\StringTok{"0 %"}\NormalTok{, }\StringTok{""}\NormalTok{), }
          \DataTypeTok{gridline.label.offset =}\DecValTok{5}\NormalTok{, }\DataTypeTok{legend.text.size=} \DecValTok{12}\NormalTok{,}
          \DataTypeTok{group.line.width =}\FloatTok{0.8}\NormalTok{,}
          \DataTypeTok{background.circle.colour  =} \StringTok{"#ffffff"}\NormalTok{, }\DataTypeTok{group.point.size   =} \DecValTok{2}\NormalTok{, }
          \DataTypeTok{axis.label.size   =}\DecValTok{3}\NormalTok{)}
\NormalTok{radar_all}
\end{Highlighting}
\end{Shaded}

\includegraphics[width=1\linewidth]{msc_thesis_figures_files/figure-latex/figure_1_model_both_radar-1}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggsave}\NormalTok{(}\StringTok{"outputs/figures/radar_ggradar_both.png"}\NormalTok{, radar_all, }\DataTypeTok{width =} \DecValTok{15}\NormalTok{, }\DataTypeTok{height =} \DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{map-of-change}{%
\subsubsection{Map of change}\label{map-of-change}}

Mapping takes a lot of time, not knitted by default

\hypertarget{chap}{%
\paragraph{Chap}\label{chap}}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# change_1 <- df_final_fordiff_pivoted %>% }
\CommentTok{#   filter(chapter %in% c("none", "both", "chap_1")) %>%}
\CommentTok{#   filter(sce %in% c('sce_38', 'sce_39')) %>% }
\CommentTok{#   ggplot(aes(fill=change)) +}
\CommentTok{#   geom_sf(show.legend=T, lwd = 0)  + }
\CommentTok{#   facet_wrap(~sce) + }
\CommentTok{#   ggtitle("Connectivity change in %", }
\CommentTok{#           subtitle = "2010-2100") +}
\CommentTok{#   scale_fill_binned(low='#d13e11', high='#fff7bc', }
\CommentTok{#                     breaks=c(-50, -40, -30, -20, -10, 0, 10))}
\CommentTok{# change_1}
\CommentTok{#ggsave("outputs/figures/connectivit_change_mun_chap1.png", change_1)}
\end{Highlighting}
\end{Shaded}

\hypertarget{chap-2-1}{%
\paragraph{Chap 2}\label{chap-2-1}}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# change_2 <- df_final_fordiff_pivoted %>% }
\CommentTok{#   filter(chapter %in% c("none", "both", "chap_2")) %>%}
\CommentTok{#   ggplot(aes(fill=change)) +}
\CommentTok{#   geom_sf(show.legend=T, lwd = 0)  + }
\CommentTok{#   facet_wrap(~sce) + }
\CommentTok{#   ggtitle("Connectivity change in %", }
\CommentTok{#           subtitle = "2010-2100") +}
\CommentTok{#   scale_fill_binned(low='#d13e11', high='#fff7bc', }
\CommentTok{#                     breaks=c(-50, -40, -30, -20, -10, 0, 10))}
\CommentTok{# change_2}
\CommentTok{#ggsave("outputs/figures/connectivit_change_mun_chap2.png", change_2)}
\end{Highlighting}
\end{Shaded}

\hypertarget{both-1}{%
\paragraph{Both}\label{both-1}}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# change_both <- df_final_fordiff_pivoted %>% }
\CommentTok{#   ggplot(aes(fill=change)) +}
\CommentTok{#   geom_sf(show.legend=T, lwd = 0)  + }
\CommentTok{#   facet_wrap(~sce) + }
\CommentTok{#   ggtitle("Connectivity change in %", }
\CommentTok{#           subtitle = "2010-2100") +}
\CommentTok{#   scale_fill_binned(low='#d13e11', high='#fff7bc', }
\CommentTok{#                     breaks=c(-50, -40, -30, -20, -10, 0, 10))}
\CommentTok{# change_both}
\CommentTok{#ggsave("outputs/figures/connectivit_change_mun_both.png", change_both)}
\end{Highlighting}
\end{Shaded}

\hypertarget{roc-curves}{%
\subsubsection{ROC Curves}\label{roc-curves}}

ROC curves from fitting both models, plotting with \texttt{patchwork}
package.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{p1 <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(urb, }\DataTypeTok{.id =} \StringTok{"fold"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{Fold =} \KeywordTok{factor}\NormalTok{(fold, }\DataTypeTok{levels =} \KeywordTok{as.character}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{10}\NormalTok{))) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(Fold) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{roc_curve}\NormalTok{(.pred, }\DataTypeTok{truth =}\NormalTok{ outcome_fact) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{autoplot}\NormalTok{(}\DataTypeTok{add=}\NormalTok{T) }\OperatorTok{+}
\StringTok{  }\KeywordTok{ggtitle}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"Urbanisation"}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{annotate}\NormalTok{(}\DataTypeTok{x =} \FloatTok{0.75}\NormalTok{, }\DataTypeTok{y =} \FloatTok{0.25}\NormalTok{, }\DataTypeTok{geom=}\StringTok{"label"}\NormalTok{, }
           \DataTypeTok{label =} \KeywordTok{as.character}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"Av AUC = 0.938 +/- 0.002"}\NormalTok{))) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"none"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\OtherTok{NULL}

\NormalTok{p2 <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(agex, }\DataTypeTok{.id =} \StringTok{"fold"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{Fold =} \KeywordTok{factor}\NormalTok{(fold, }\DataTypeTok{levels =} \KeywordTok{as.character}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{10}\NormalTok{))) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(Fold) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{roc_curve}\NormalTok{(.pred, }\DataTypeTok{truth =}\NormalTok{ outcome_fact) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{autoplot}\NormalTok{(}\DataTypeTok{add=}\NormalTok{T) }\OperatorTok{+}
\StringTok{  }\KeywordTok{ggtitle}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"Agricultural Expansion"}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{annotate}\NormalTok{(}\DataTypeTok{x =} \FloatTok{0.75}\NormalTok{, }\DataTypeTok{y =} \FloatTok{0.25}\NormalTok{, }\DataTypeTok{geom=}\StringTok{"label"}\NormalTok{, }
           \DataTypeTok{label =} \KeywordTok{as.character}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"Av AUC = 0.929 +/- 0.002"}\NormalTok{))) }\OperatorTok{+}
\StringTok{  }\OtherTok{NULL}

\NormalTok{full =}\StringTok{ }\NormalTok{p1 }\OperatorTok{+}\StringTok{ }\NormalTok{p2}

\NormalTok{full}
\end{Highlighting}
\end{Shaded}

\includegraphics[width=1\linewidth]{msc_thesis_figures_files/figure-latex/roc_curves-1}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggsave}\NormalTok{(}\StringTok{"outputs/figures/double_roc_resample.png"}\NormalTok{, full)}
\end{Highlighting}
\end{Shaded}

\hypertarget{histograms}{%
\subsubsection{Histograms}\label{histograms}}

\hypertarget{chap-1-1}{%
\paragraph{Chap 1}\label{chap-1-1}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{hist_plot_}\DecValTok{1}\NormalTok{ <-}\StringTok{ }\NormalTok{histograms }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{ts =} \KeywordTok{as.factor}\NormalTok{(ts)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(chapter }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{'both'}\NormalTok{, }\StringTok{'chap_1'}\NormalTok{)) }\OperatorTok{%>%}\StringTok{  }
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{bins, }\DataTypeTok{y=}\NormalTok{n, }\DataTypeTok{group=}\NormalTok{ts, }\DataTypeTok{colour=}\NormalTok{ts)) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{geom_density}\NormalTok{(}\DataTypeTok{stat=}\StringTok{'identity'}\NormalTok{, }\DataTypeTok{show.legend=}\NormalTok{F, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\KeywordTok{factor}\NormalTok{(ts)), }\DataTypeTok{alpha=}\FloatTok{0.3}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_grid}\NormalTok{(code}\OperatorTok{~}\NormalTok{species) }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x =} \StringTok{"Flow intensity distribution (log)"}\NormalTok{, }
       \DataTypeTok{y =} \StringTok{""}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{strip.text.y =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle=}\DecValTok{360}\NormalTok{, }\DataTypeTok{size=}\DecValTok{10}\NormalTok{, }\DataTypeTok{hjust =} \DecValTok{0}\NormalTok{), }
        \DataTypeTok{axis.text.x=}\KeywordTok{element_blank}\NormalTok{(),}
        \DataTypeTok{axis.ticks.x=}\KeywordTok{element_blank}\NormalTok{(), }
        \DataTypeTok{axis.text.y.left =} \KeywordTok{element_blank}\NormalTok{(),}
        \DataTypeTok{axis.ticks.y.left =} \KeywordTok{element_blank}\NormalTok{()) }\OperatorTok{+}
\StringTok{  }\OtherTok{NULL}
\NormalTok{hist_plot_}\DecValTok{1}
\end{Highlighting}
\end{Shaded}

\includegraphics[width=1\linewidth,height=1.5\textheight]{msc_thesis_figures_files/figure-latex/hist_plot_1-1}

\hypertarget{chap-2-2}{%
\paragraph{Chap 2}\label{chap-2-2}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{hist_plot_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\NormalTok{histograms }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{ts =} \KeywordTok{as.factor}\NormalTok{(ts)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(chapter }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{'both'}\NormalTok{, }\StringTok{'chap_2'}\NormalTok{)) }\OperatorTok{%>%}\StringTok{  }
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{bins, }\DataTypeTok{y=}\NormalTok{n, }\DataTypeTok{group=}\NormalTok{ts, }\DataTypeTok{colour=}\NormalTok{ts)) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{geom_density}\NormalTok{(}\DataTypeTok{stat=}\StringTok{'identity'}\NormalTok{, }\DataTypeTok{show.legend=}\NormalTok{F, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\KeywordTok{factor}\NormalTok{(ts)), }\DataTypeTok{alpha=}\FloatTok{0.3}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_grid}\NormalTok{(code}\OperatorTok{~}\NormalTok{species) }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x =} \StringTok{"Flow intensity distribution (log)"}\NormalTok{, }
       \DataTypeTok{y =} \StringTok{""}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{strip.text.y =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle=}\DecValTok{360}\NormalTok{, }\DataTypeTok{size=}\DecValTok{10}\NormalTok{, }\DataTypeTok{hjust =} \DecValTok{0}\NormalTok{), }
        \DataTypeTok{axis.text.x=}\KeywordTok{element_blank}\NormalTok{(),}
        \DataTypeTok{axis.ticks.x=}\KeywordTok{element_blank}\NormalTok{(), }
        \DataTypeTok{axis.text.y.left =} \KeywordTok{element_blank}\NormalTok{(),}
        \DataTypeTok{axis.ticks.y.left =} \KeywordTok{element_blank}\NormalTok{()) }\OperatorTok{+}
\StringTok{  }\OtherTok{NULL}
\NormalTok{hist_plot_}\DecValTok{2}
\end{Highlighting}
\end{Shaded}

\includegraphics[width=1\linewidth]{msc_thesis_figures_files/figure-latex/hist_plot_2-1}

\hypertarget{surf-analysis}{%
\subsubsection{SURF Analysis}\label{surf-analysis}}

\hypertarget{chap1}{%
\paragraph{Chap1}\label{chap1}}

\hypertarget{linear-graph}{%
\subparagraph{Linear Graph}\label{linear-graph}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{surf_}\DecValTok{1}\NormalTok{ <-}\StringTok{ }\NormalTok{surf }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{scenario =} \KeywordTok{as.factor}\NormalTok{(scenario)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(climate }\OperatorTok{!=}\StringTok{ "none"}\NormalTok{, chapter }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"none"}\NormalTok{, }\StringTok{"both"}\NormalTok{, }\StringTok{"chap_1"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(climate }\OperatorTok{!=}\StringTok{ "none"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(scenario, species, iter) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_modify}\NormalTok{(}\OperatorTok{~}\KeywordTok{mutate}\NormalTok{(.x , }\DataTypeTok{baseline =} \KeywordTok{subset}\NormalTok{(.x, timestep }\OperatorTok{==}\StringTok{ }\DecValTok{2010}\NormalTok{)}\OperatorTok{$}\NormalTok{kp_nb, }
                       \DataTypeTok{the_diff =}\NormalTok{ ((kp_nb}\OperatorTok{-}\NormalTok{baseline)}\OperatorTok{/}\NormalTok{baseline)}\OperatorTok{*}\DecValTok{100}\NormalTok{ )) }\OperatorTok{%>%}\StringTok{ }\NormalTok{ungroup }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ timestep, }\DataTypeTok{y =}\NormalTok{ the_diff, }\DataTypeTok{color=}\NormalTok{climate)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_smooth}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{group =}\NormalTok{ scenario, }\DataTypeTok{linetype=}\NormalTok{code_run), }\DataTypeTok{alpha=}\FloatTok{0.2}\NormalTok{, }\DataTypeTok{method=}\StringTok{"loess"}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\CommentTok{#geom_point(aes(group=scenario), alpha= 0.05, size = 0.5)+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{species, }\DataTypeTok{scales =} \StringTok{"free"}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values=}\KeywordTok{c}\NormalTok{(}\StringTok{'#ffe300'}\NormalTok{, }\StringTok{'#ff7a14'}\NormalTok{, }\StringTok{"#b31400"}\NormalTok{), }
                     \DataTypeTok{labels =} \KeywordTok{c}\NormalTok{(}\StringTok{"Historic"}\NormalTok{, }\StringTok{"Baseline"}\NormalTok{, }\StringTok{"RCP 8.5"}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{y =} \StringTok{"Change in nb of Keypoints detected (%)"}\NormalTok{,}
       \DataTypeTok{x =} \StringTok{"Year"}\NormalTok{,}
       \DataTypeTok{col =} \StringTok{"Climate"}\NormalTok{,}
       \DataTypeTok{linetype =} \StringTok{"Run"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{col =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{ncol =} \DecValTok{2}\NormalTok{, }\DataTypeTok{nrow=}\DecValTok{2}\NormalTok{), }
         \DataTypeTok{linetype =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{nrow=}\DecValTok{1}\NormalTok{, }\DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{colour =} \StringTok{'black'}\NormalTok{))) }\OperatorTok{+}
\StringTok{  }\OtherTok{NULL}
\NormalTok{surf_}\DecValTok{1}
\end{Highlighting}
\end{Shaded}

\includegraphics{msc_thesis_figures_files/figure-latex/surf_1-1.pdf}

\hypertarget{radar-graph-3}{%
\subparagraph{Radar Graph}\label{radar-graph-3}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{surf_radar_data_}\DecValTok{1}\NormalTok{ <-}\StringTok{ }\NormalTok{surf }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(climate }\OperatorTok{!=}\StringTok{ "none"}\NormalTok{, chapter }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"none"}\NormalTok{, }\StringTok{"both"}\NormalTok{,}\StringTok{"chap_1"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(sce }\OperatorTok{!=}\StringTok{ "sce_0"}\NormalTok{, name }\OperatorTok{!=}\StringTok{ "historic run"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(timestep, species, code) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{summarise}\NormalTok{(}\DataTypeTok{kp_nb =} \KeywordTok{mean}\NormalTok{(kp_nb)) }\OperatorTok{%>%}\StringTok{ }\NormalTok{ungroup }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{pivot_wider}\NormalTok{(}\DataTypeTok{names_from =}\NormalTok{ timestep, }\DataTypeTok{values_from =}\NormalTok{ kp_nb) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{diff =}\NormalTok{ ((}\StringTok{`}\DataTypeTok{2100}\StringTok{`}\OperatorTok{-}\StringTok{`}\DataTypeTok{2010}\StringTok{`}\NormalTok{)}\OperatorTok{/}\StringTok{`}\DataTypeTok{2010}\StringTok{`}\NormalTok{)}\OperatorTok{*}\DecValTok{100}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(}\StringTok{`}\DataTypeTok{2010}\StringTok{`}\OperatorTok{:}\StringTok{`}\DataTypeTok{2100}\StringTok{`}\NormalTok{)) }\OperatorTok{%>%}\StringTok{  }
\StringTok{  }\KeywordTok{pivot_wider}\NormalTok{(}\DataTypeTok{names_from =}\NormalTok{ code, }\DataTypeTok{values_from =}\NormalTok{ diff) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{group =}\NormalTok{ species)}

\NormalTok{surf_radar_}\DecValTok{1}\NormalTok{ <-}\StringTok{ }
\StringTok{  }\KeywordTok{ggradar}\NormalTok{(surf_radar_data_}\DecValTok{1}\NormalTok{, }\DataTypeTok{centre.y =} \DecValTok{-60}\NormalTok{, }\DataTypeTok{legend.position =} \StringTok{"right"}\NormalTok{,}
          \DataTypeTok{grid.min =} \DecValTok{-60}\NormalTok{, }\DataTypeTok{grid.max =} \DecValTok{5}\NormalTok{, }\DataTypeTok{grid.mid =} \DecValTok{0}\NormalTok{, }
          \DataTypeTok{values.radar =} \KeywordTok{c}\NormalTok{(}\StringTok{"-60 %"}\NormalTok{, }\StringTok{"0 %"}\NormalTok{, }\StringTok{""}\NormalTok{), }
          \DataTypeTok{gridline.label.offset =}\DecValTok{22}\NormalTok{, }\DataTypeTok{legend.text.size=} \DecValTok{12}\NormalTok{,}
          \DataTypeTok{group.line.width =}\FloatTok{0.8}\NormalTok{,}
          \DataTypeTok{background.circle.colour  =} \StringTok{"#ffffff"}\NormalTok{, }\DataTypeTok{group.point.size   =} \DecValTok{2}\NormalTok{, }
          \DataTypeTok{axis.label.size   =}\DecValTok{3}\NormalTok{)}
\NormalTok{surf_radar_}\DecValTok{1}
\end{Highlighting}
\end{Shaded}

\includegraphics{msc_thesis_figures_files/figure-latex/surf_radar_1-1.pdf}

\hypertarget{chap-2-3}{%
\paragraph{Chap 2}\label{chap-2-3}}

\hypertarget{linear-graph-1}{%
\subparagraph{Linear Graph}\label{linear-graph-1}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{surf_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\NormalTok{surf }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{scenario =} \KeywordTok{as.factor}\NormalTok{(scenario)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(climate }\OperatorTok{!=}\StringTok{ "none"}\NormalTok{, chapter }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"none"}\NormalTok{, }\StringTok{"both"}\NormalTok{, }\StringTok{"chap_2"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(climate }\OperatorTok{!=}\StringTok{ "none"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(scenario, species, iter) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_modify}\NormalTok{(}\OperatorTok{~}\KeywordTok{mutate}\NormalTok{(.x , }\DataTypeTok{baseline =} \KeywordTok{subset}\NormalTok{(.x, timestep }\OperatorTok{==}\StringTok{ }\DecValTok{2010}\NormalTok{)}\OperatorTok{$}\NormalTok{kp_nb, }
                       \DataTypeTok{the_diff =}\NormalTok{ ((kp_nb}\OperatorTok{-}\NormalTok{baseline)}\OperatorTok{/}\NormalTok{baseline)}\OperatorTok{*}\DecValTok{100}\NormalTok{ )) }\OperatorTok{%>%}\StringTok{ }\NormalTok{ungroup }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ timestep, }\DataTypeTok{y =}\NormalTok{ the_diff, }\DataTypeTok{color=}\NormalTok{climate)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_smooth}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{group =}\NormalTok{ scenario, }\DataTypeTok{linetype=}\NormalTok{code_run), }\DataTypeTok{alpha=}\FloatTok{0.2}\NormalTok{, }\DataTypeTok{method=}\StringTok{"loess"}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\CommentTok{#geom_point(aes(group=scenario), alpha= 0.05, size = 0.5)+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{species, }\DataTypeTok{scales =} \StringTok{"free"}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values=}\KeywordTok{c}\NormalTok{(}\StringTok{'#ffe300'}\NormalTok{, }\StringTok{'#ff7a14'}\NormalTok{, }\StringTok{"#b31400"}\NormalTok{), }
                     \DataTypeTok{labels =} \KeywordTok{c}\NormalTok{(}\StringTok{"Historic"}\NormalTok{, }\StringTok{"Baseline"}\NormalTok{, }\StringTok{"RCP 8.5"}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{y =} \StringTok{"Change in nb of Keypoints detected (%)"}\NormalTok{,}
       \DataTypeTok{x =} \StringTok{"Year"}\NormalTok{,}
       \DataTypeTok{col =} \StringTok{"Climate"}\NormalTok{,}
       \DataTypeTok{linetype =} \StringTok{"Run"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{col =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{ncol =} \DecValTok{2}\NormalTok{, }\DataTypeTok{nrow=}\DecValTok{2}\NormalTok{), }
         \DataTypeTok{linetype =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{nrow=}\DecValTok{2}\NormalTok{, }\DataTypeTok{ncol=}\DecValTok{2}\NormalTok{, }
                                 \DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{colour =} \StringTok{'black'}\NormalTok{))) }\OperatorTok{+}
\StringTok{  }\OtherTok{NULL}
\NormalTok{surf_}\DecValTok{2}
\end{Highlighting}
\end{Shaded}

\includegraphics{msc_thesis_figures_files/figure-latex/surf_2-1.pdf}

\hypertarget{radar-graph-4}{%
\subparagraph{Radar Graph}\label{radar-graph-4}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{surf_radar_data_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\NormalTok{surf }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(climate }\OperatorTok{!=}\StringTok{ "none"}\NormalTok{, chapter }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"none"}\NormalTok{, }\StringTok{"both"}\NormalTok{,}\StringTok{"chap_2"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(sce }\OperatorTok{!=}\StringTok{ "sce_0"}\NormalTok{, name }\OperatorTok{!=}\StringTok{ "historic run"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(timestep, species, code) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{summarise}\NormalTok{(}\DataTypeTok{kp_nb =} \KeywordTok{mean}\NormalTok{(kp_nb)) }\OperatorTok{%>%}\StringTok{ }\NormalTok{ungroup }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{pivot_wider}\NormalTok{(}\DataTypeTok{names_from =}\NormalTok{ timestep, }\DataTypeTok{values_from =}\NormalTok{ kp_nb) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{diff =}\NormalTok{ ((}\StringTok{`}\DataTypeTok{2100}\StringTok{`}\OperatorTok{-}\StringTok{`}\DataTypeTok{2010}\StringTok{`}\NormalTok{)}\OperatorTok{/}\StringTok{`}\DataTypeTok{2010}\StringTok{`}\NormalTok{)}\OperatorTok{*}\DecValTok{100}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(}\StringTok{`}\DataTypeTok{2010}\StringTok{`}\OperatorTok{:}\StringTok{`}\DataTypeTok{2100}\StringTok{`}\NormalTok{)) }\OperatorTok{%>%}\StringTok{  }
\StringTok{  }\KeywordTok{pivot_wider}\NormalTok{(}\DataTypeTok{names_from =}\NormalTok{ code, }\DataTypeTok{values_from =}\NormalTok{ diff) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{group =}\NormalTok{ species)}

\NormalTok{surf_radar_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }
\StringTok{  }\KeywordTok{ggradar}\NormalTok{(surf_radar_data_}\DecValTok{2}\NormalTok{, }\DataTypeTok{centre.y =} \DecValTok{-60}\NormalTok{, }\DataTypeTok{legend.position =} \StringTok{"right"}\NormalTok{,}
          \DataTypeTok{grid.min =} \DecValTok{-60}\NormalTok{, }\DataTypeTok{grid.max =} \DecValTok{5}\NormalTok{, }\DataTypeTok{grid.mid =} \DecValTok{0}\NormalTok{, }
          \DataTypeTok{values.radar =} \KeywordTok{c}\NormalTok{(}\StringTok{"-60 %"}\NormalTok{, }\StringTok{"0 %"}\NormalTok{, }\StringTok{""}\NormalTok{), }
          \DataTypeTok{gridline.label.offset =}\DecValTok{22}\NormalTok{, }\DataTypeTok{legend.text.size=} \DecValTok{12}\NormalTok{,}
          \DataTypeTok{group.line.width =}\FloatTok{0.8}\NormalTok{,}
          \DataTypeTok{background.circle.colour  =} \StringTok{"#ffffff"}\NormalTok{, }\DataTypeTok{group.point.size   =} \DecValTok{2}\NormalTok{, }
          \DataTypeTok{axis.label.size   =}\DecValTok{3}\NormalTok{)}
\NormalTok{surf_radar_}\DecValTok{2}
\end{Highlighting}
\end{Shaded}

\includegraphics{msc_thesis_figures_files/figure-latex/surf_radar_2-1.pdf}

\end{document}