### 2020-08-01 thesis/appendix_2.tex

% !TEX root = ./thesis.tex

\chapter*{\textbf{Appendix 2: Workshop Materials \\ \hspace{1em}}}
\addcontentsline{toc}{chapter}{Chapter II Workshop Materials}

\setcounter{chapter}{4}
\setcounter{table}{0}
\setcounter{figure}{0}

%TODO Add description of each elements in this

\includepdf[pages=-,pagecommand={},width=1.3\textwidth]{thesis/appendix/Cahier_du_Participant.pdf}
\includepdf[pages=-,pagecommand={},width=1.3\textwidth]{thesis/appendix/Document_du_facilitateur.pdf}
\includepdf[pages=-,pagecommand={},width=1.3\textwidth]{thesis/appendix/Ethics_approval_Lucet.pdf}