### 2020-08-01 ReadMe.rmd
update sreadme

---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### M.Sc. Thesis 2018-2020

![Docker image build](https://github.com/VLucet/landchange-connectivity-monteregie/workflows/Build%20/%20Push%20Docker%20image/badge.svg)

[GitHub page](https://vlucet.github.io/landchange-connectivity-monteregie/)

[Thesis](https://github.com/VLucet/landchange-connectivity-monteregie/blob/master/thesis/build/thesis.pdf)