### 2020-08-28 thesis/appendix_3.tex
add apendix 3

% !TEX root = ./thesis.tex

\chapter*{\textbf{Appendix 3: LANDIS model details \\ \hspace{1em}}}
\addcontentsline{toc}{chapter}{Chapter II Workshop Materials}

\setcounter{chapter}{5}
\setcounter{table}{0}
\setcounter{figure}{0}

We used the simulation framework LANDIS-II v7 \citep{scheller_forest_2004} to project forest type changes in the future. We used the output to parameterize STSIM forest type transitions. It was necessary to combine these outputs in that way because LANDIS-II does not support land use change natively.
In LANDIS-II, stand level forest processes occur in each cell of the grid that represents the landscape. Landscape-level processes can affect multiple cells. 
We set a resolution of 90 m and a 10-year time step. Forest composition in each cell was derived from the forest inventory plots of the Quebec government and the Canadian National Forest Inventory. Each of these cells was then assigned to a “landtype” with homogeneous soil and climate conditions (see details in \cite{boulanger_climate_2019}).\\

%\subsubsection*{Climate scenarios}

Climate scenarios were derived from the Canadian Earth System Model version 2 (CanESM2). Climate projections from CanESM2 were accessed from the World Climate Research Program (WCRP) Climate Model Intercomparison Project Phase 5 (CMIP5) archive for RCP 4.5 and RCP 8.5 (details in \cite{boulanger_climate_2019}).
LANDIS-II requires specific input parameters which are generally derived from more ecologically detailed, fine-scale forest stand models. We used PICUS version 1.5. (see details in \cite{lexer_modified_2001, tremblay_harvesting_2018}). \\

%\subsubsection*{Forest Harvesting}

Forest harvesting was simulated using the Biomass Harvest extension (v4.3; \cite{gustafson_spatial_2000}). The harvesting scenario - ecosystem based forest management, EBFM - emulates different intensities of clearcutting and partial harvesting, which vary according to vegetations as defined by Quebec’s Hierarchical System for Territorial Ecological Classification \citep{bergeron_quebec_1992}. It should closely mimic the historical disturbance regimes. Harvest was performed by ‘‘management areas’’: forest management units (public lands) or ecodistricts (private lands, see details in \cite{boulanger_climate_2019}).