### 2020-08-29 thesis/thesis.tex

\documentclass[12pt,Bold,TexShade]{thesis/mcgilletdclass}
\usepackage{graphicx}
\usepackage{times}
\usepackage{subfig}
\usepackage{longtable}
\usepackage{float}
\usepackage{setspace}
\usepackage{multirow} % allows multiple rows in table
\usepackage{adjustbox}
\usepackage{enumitem}
\usepackage{rotating}
\usepackage{lipsum,lineno} % for placeholder text
\usepackage{array} % for sizing of tables
\usepackage{lscape} % for landscape tables
\usepackage{pdfpages} % to add pdf appendix
\usepackage{siunitx} % for km2 nottation

% Bib style
% \bibliographystyle{unsrt}
%\setcitestyle{authoryear,open={((},close={))}}
  \usepackage[
    style=authoryear,
    maxnames = 2,
    backend=biber,
    natbib=true %, 
    %refsection=chapter
  ]{biblatex}

\addbibresource{thesis/mcgilletd.bib}
\addbibresource{thesis/software.bib}

% remove section numbering without using section* (keeps in TOC)
\setcounter{secnumdepth}{4}

% % for PDF/A
% \usepackage{hyperref}
% \hypersetup{
%     colorlinks=true,
%     linkcolor=black,
%     filecolor=black,
%     urlcolor=blue,
%     citecolor=black
% }
% \usepackage[a-1b]{pdfx}

% Reduce number of words breaking over lines
\hyphenpenalty=1000

% Cross-referencing
\usepackage{xr}
\externaldocument{"thesis/chap_1"}
\externaldocument{"thesis/chap_2"}
\externaldocument{"thesis/appendix_1"}

\usepackage{hyperref} % for hyperlink accross the document

% May or may not be necessary to have separate bibliographies per chapter
% McGill requirements ask for it, but others did not do so
% https://engineering.purdue.edu/~mark/puthesis/faq/chapter-bibliographies/

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Have you configured your TeX system for proper  %%
%% page alignment? See the McGillETD documentation %%
%% for two methods that can be used to control     %%
%% page alignment. One method is demonstrated      %%
%% below. See documentation and the ufalign.tex    %%
%% file for instructions on how to adjust these    %%
%% parameters.                                     %%
\addtolength{\hoffset}{0pt}                        %%
\addtolength{\voffset}{0pt}                        %%
%%                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%       Define student-specific info

\SetTitle{\huge{Integrating land use and land cover change simulations and connectivity modelling: a case study in the Montérégie region in southern Québec}}%

\SetAuthor{\large{\textit{Valentin Lucet}}}%
\SetDegreeType{Master of Science\vspace{0cm}}%
\SetDepartment{Department of Biology\vspace{-2.1cm}}%
\SetUniversity{McGill University}%
\SetUniversityAddr{Montréal, Québec}%
\SetThesisDate{August 2020}%
\SetRequirements{A thesis submitted to McGill University in partial fulfillment \\of the requirements of the degree of Master of Science}%

\SetCopyright{\textcopyright Valentin Lucet, 2020 \\ All rights reserved}%

%% Input any special commands below
\listfiles

% Has to be loaded LAST

\begin{document}

\maketitle

\frontmatter

\begin{romanPagenumber}{2}%
\TOCHeading{\MakeUppercase{Table of Contents}}%
\LOTHeading{{List of Tables}}%
\LOFHeading{{List of Figures}}%
\tableofcontents %

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%         English Abstract                        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\SetAbstractEnName{{Abstract}}%
\SetAbstractEnText{Connectivity conservation science, whose goal is to preserve the continuity of habitat throughout a given landscape, proceeds by identifying priority areas given the current configuration of the landscape. However, current connectivity conservation planning methods often do not take into account risks associated with future land use and climate change, and often do not confront the results of the prioritization with the priorities perceived by stakeholders. In this thesis, we attempted to remedy those two issues in the context of an ongoing effort of connectivity conservation planning for the region of Montérégie in Southern Québec, Canada. In the first chapter, we built on past work of connectivity modelling using circuit theory in the region and complemented it with land use and climate change modelling that uses a combination of statistical modelling and MCMC-based simulations. Models trained on past land use data were used to project future land use changes and estimate future changes in potential functional connectivity for 5 different umbrella species. In a second chapter, we integrate these results with the perceived conservation priorities in the region. We conducted a day-long workshop with stakeholders involved in the ongoing connectivity conservation planning effort in the region and collected information on landscape features considered “core” and “linkages” priority areas. We showed the importance of considering land use changes to produce a resilient network of protected areas and highlight the need for a multi-stakeholder approach in the definition of conservation scenarios and the identification of regional priorities.}

\AbstractEn%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%         French Abstract                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\SetAbstractFrName{{Abr\'{e}g\'{e}}}%
\SetAbstractFrText{La science de la conservation de la connectivité, dont le but est de préserver la continuité de l'habitat dans un paysage donné, procède par l'identification des zones prioritaires compte tenu de la configuration actuelle du paysage. Toutefois, les méthodes actuelles de planification de la conservation de la connectivité ne tiennent souvent pas compte des risques associés aux changements futurs d'utilisation des terres, et vont rarement jusqu'à confronter les résultats de la priorisation avec les priorités perçues par les parties prenantes. Dans cette thèse, nous avons tenté de remédier à ces deux problèmes dans le cadre d'un effort continu de planification de la conservation de la connectivité pour la région de la Montérégie, dans le sud du Québec, au Canada. Dans le premier chapitre, nous nous sommes appuyés sur les travaux antérieurs de modélisation de la connectivité basés sur la théorie des circuits dans la région et nous les avons complétés par une modélisation de l'utilisation des terres et du changement climatique qui utilise une combinaison de modélisation statistique et de simulations à base de chaînes Markov Monte Carlo. Le modèle construit à partir des données d'utilisation des terres passées ont été utilisés pour projeter les changements futurs d'utilisation des terres et estimer les changements futurs de la connectivité fonctionnelle potentielle pour 5 espèces de parapluie différentes. Dans un deuxième chapitre, nous comparons ces résultats avec les priorités de conservation perçues dans la région. Nous avons organisé un atelier d'une journée avec les parties prenantes impliquées dans l'effort continu de planification de la conservation de la connectivité dans la région et nous avons recueilli des informations sur les caractéristiques du paysage considérées comme des zones prioritaires "centrales" et "de liaison". Nous avons montré l'importance de prendre en compte les changements d'utilisation des terres pour produire un réseau de zones protégées résilient et avons souligné la nécessité d'une approche multipartite dans la définition des scénarios de conservation et l'identification des priorités régionales.}%
\AbstractFr%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%         Acknowledgements                   %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\SetAcknowledgeName{{Acknowledgements}}%

% Among other acknowledgements, the student is required to declare the extent to which assistance (paid or unpaid) has been given by members of staff, fellow students, research assistants, technicians, or others in the collection of materials and data, the design and construction of apparatus, the performance of experiments, the analysis of data, and the preparation of the thesis (including editorial help).
% In addition, it is appropriate to recognize the supervision and advice given by the thesis supervisor(s) and advisors.

\SetAcknowledgeText{
For providing vital guidance and parts of the code, I thank Dr. Bronwyn Rayfield. For providing help with computing resources parts of the code, I thank Dr. Guillaume Larocque. For providing advice on STSIM implementation and code, I thank Dr. Colin Daniel and Dr. Leonardo Frid. I thank Dr. Guillaume Larocque and Dr. Yann Boulanger for their help in providing the LANDIS outputs and editorial help for the associated appendix. \\
\indent I thank all students, staff and organizations involved in the design and realisation of the workshop. I thank Aurélie Sierra for her invaluable guidance and counsel. I thank all students and staff facilitators: Madeleine Gauthier, Mikayla Salmon-Beitel and Violette Bertrand. I thank Geneviève Poirier and Mariève Lafontaine for their participation in the design, planning and organization of the workshop. I also thank Renée Gagnon for her precious help securing a location for the workshop. \\
\indent For providing essential funding for this project, I thank the Fonds de recherche du Québec – Nature et technologies (FRQNT). For providing logistical support I thank Dr. Helen Elina and Mr. Philippe Auzel. \\
\indent For providing key information on land use data, I thank Martin Joly, Marie-Josée Côté and Benoit Jobin. \\
\indent I thank my supervisor, Prof. Andrew Gonzalez, for his support and for believing in my abilities every step of the way. I also thank members of my committee, Prof. Brian Leung and Prof. Elena Bennett for their timely and wise council at every committee meeting. I thank all my fellow grad students and especially Alex Arkilanian for all the help revising proposals, discussion ideas, and providing key support. \\
\indent I thank my mother Laura, my father Philippe, and Talia for their unwavering support. Thank you to Blanka and Peter for sheltering me and putting up with my writing process during the thick of the pandemic. \\
}%
\Acknowledge%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%       THESIS FORMAT AND STYLE       %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\SetLinkName{{Thesis Format}}%
\SetLinkText{This thesis is written in a manuscript-based format.
\\
\\
\textbf{Chapter 1:} Lucet, V. Gonzalez, A. (2020), Integrating Land Use Change Modelling with Connectivity Modelling. %\textit{Journal of Ecographic Informatics} 10 (1), 878887.
\\
\\
\textbf{Chapter 2:} Lucet, V. Gonzalez, A. (2020), Integrating Stakeholders Perceptions of Connectivity Conservation Priorities into a Spatially Explicit Land Use and Connectivity Change Model. %\textit{Journal of Ecographic Informatics} 10 (1), 878887.
\\
\\
}
\Link%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%        Contribution of Authors                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% A doctoral thesis must clearly state the elements of the thesis that are considered original scholarship and distinct contributions to knowledge.
% Contributions of the student to each chapter must be explicitly stated.
% Contributions of any co-authors to each chapter must be explicitly stated.
\SetContribAuthorsName{{Contribution of Authors}}
\SetContribAuthorsText{\noindent
I am the first author for all chapters and the appendix in this
thesis.
\\
\\
\textbf{Chapter 1:} 
I wrote the manuscript with input from my supervisor.
\\
\\
\textbf{Chapter 2:}
I wrote the manuscript with input from my supervisor.
\\
\\
}%
\ContribAuthors %
\listoffigures %
\listoftables %

\end{romanPagenumber}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%        General Introduction                     %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Clearly state the rationale and objectives of the research.
\newcommand*{\SetIntroName}[1]{\renewcommand*{\ETDIntroName}{#1}}%
\newcommand*{\ETDIntroName}{Intro}%

\newcommand{\SetIntroText}[1]{\renewcommand*{\ETDIntroText}{#1}}%
\newcommand*{\ETDIntroText}{Intro text goes here!}%

\newcommand{\Intro}{%
    \begin{simpleenv}{}{}{}{}%
    \pagestyle{plain}%
  \setboolean{SetDSpace}{false}%
    \GoSingle%
    \begin{QZ@Cent}%
          \bfseries{\ETDIntroName}
      \end{QZ@Cent}%
    \vspace*{0.5in}
      \par%
      \GoDouble%
      \setlength{\parskip}{-2em}
      %\chapter*{\ETDIntroName}%
      \phantomsection\section*{}
      \addcontentsline{toc}{extrachapter}{\ETDIntroName}%
      \ETDIntroText%
    \end{simpleenv}%
    \setboolean{SetDSpace}{true}}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Introduction
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\SetIntroName{{General Introduction}}
\SetIntroText{

Space is a finite resource: there is a limit on how much land is available for different human activities. How communities manage and govern space is a reflection of the choices and trade offs made by different people and organizations at different spatial and temporal scales, making the study of land systems a social, historical, and ecological discipline. The political and governance processes by which those trade-offs are resolved is of vital importance for biodiversity and ecosystem function. These processes ultimately determine and regulate land use and land cover change (LUC): if and how the resources held on the land are exploited, transformed or conserved. \\

LUC is a major cause of habitat loss and fragmentation, and one of the most important threats to biodiversity and ecosystem function worldwide \citep{newbold_global_2015}. It is estimated that LUC was responsible for more than one-third of animal extinctions between 1600 and 1992 \citep{centre_global_1992}. Moreover, a recent report from the IPBES (Intergovernmental Science-Policy Platform on Biodiversity and Ecosystem Services) shows that about about 9\% of all described terrestrial species might suffer from habitat loss to the point that it threatens their long-term persistence \citep{ipbes_summary_2019}. Rarely, however, is habitat loss a complete process; once continuous tracts of habitat are often left as a mosaic of habitat fragments of different sizes separated by land cover with differing degrees of value to the resident fauna and flora.  Habitat fragmentation drives additional loss of diversity and ecosystem services because it reduces the connectivity of the landscape \citep{haddad_habitat_2015}. \\

Ecological connectivity is the extent to which the landscape supports the movements of organisms \citep{gonzalez_spatial_2017}, and is paramount for the resilience of both populations and ecosystem services \citep{mitchell_monteregie_2015} in heterogeneous and fragmented landscapes. Land use changes such as urban sprawl can cause deforestation, fragmenting habitats, and slowly eroding ecological connectivity. Many urban landscapes are experiencing uncontrolled urban sprawl and have suffered losses in connectivity and ecosystem services as a consequence. Examples include cities like Barcelona \citep{marulli_gis_2005}, New York City \citep{mcphearson_urban_2014}, and also Montréal \citep{dupras_urban_2015}.  The drivers behind those land use changes are complex and understanding them is an obstacle to conservation planning  \citep{worboys_connectivity_2010}, including connectivity conservation planning. \\

Connectivity conservation planning (CCP), refers to the enterprise that engages multiple actors such as academics, NGOs, governmental bodies at different scales and in a common goal to conserve and restore the ecological connectivity of the landscape. CCP methods usually involve modelling the connectivity of the landscape of interest and using a prioritization method to determine conservation priorities, and modify land use practices to conserve or improve connectivity. This thesis  contributes to an ongoing CCP in Montérégie, in southern Québec. In the following, we introduce connectivity modelling. We then introduce the study region and describe previous CCP and research. We point out ways in which CCP methods can be improved and how this thesis contributes to doing so. We then introduce the two other key concepts of the thesis: land use change modelling and participatory methods. \\ \vspace{2em}

\subsection*{\textit{Connectivity modelling \\ \vspace{1em}}}

Connectivity is a pillar of spatial ecology, the study of how ecological dynamics and interactions are structured in space. Connectivity is well known to influence various aspects of populations and community ecology, including dispersal, demographic rescue, and the spread of diseases. In conservation, connectivity is seen as a landscape feature that can be protected and restored to counteract the effects of  land use and climate change, and ensure the persistence of species in a landscape. \\

Connectivity as a concept has been used loosely in ecology and conservation. In their textbook “Spatial Ecology and Conservation Modeling”, \cite{fletcher_spatial_2018} identify up to 11 different variations on the concept of connectivity. For the focus of this thesis, it is important to make two important distinctions with regard to connectivity First, a distinction is often made between structural and functional connectivity. Structural connectivity emphasizes landscape structure and configuration, while functional connectivity attempts to capture species’s diverse responses to landscape’s configuration, and pays more attention to species-specific ways of moving across the landscape. Within functional connectivity another distinction is to be made between potential and realized connectivity \citep{calabrese_comparison_2004}. Potential connectivity estimates how much a landscape could facilitate or impede animal movements, based on indirect information, such as species-specific models and expert knowledge. Realized connectivity, on the other hand, relies on observational data to measure animal movement directly to quantify connectivity. This thesis focuses on modelling potential functional connectivity. \\

Connectivity modelling methods depend on the level of organization (individual, population, species or community) and the spatial scale (patch or  landscape) that connectivity is considered. To cover them all is beyond the scope of this introduction, but it is important to replace our choice of methodology in the larger body of foundational work on spatial ecological modelling. This thesis focuses on species level modelling at the landscape scale, and this focus has to be justified. Important efforts in the field of spatial ecological modelling relied on metapopulation theory, island biogeography (IBT), and population genetics. These bodies of theory spurred different approaches to modelling. For example, an IBT approach is focused on predicting diversity (number of species) in island patches \citep{macarthur_theory_1967} and has largely influenced other bodies of work in spat \citep{warren_how_2020}. On the other hand, metapopulation theory is focused on predicting long-term patch occupancy \citep{hanski_habitat_1999, hanski_metapopulation_2000}. Finally, population genetics focuses on predicting gene variation and flow within populations \citep{lewontin_population_1985, charlesworth_population_2017}. Unlike these methods, our goal is to estimate potential animal movement at the species level, over a large spatial and temporal extent. We therefore rely on a more appropriate approach which combines habitat suitability analysis and circuit theory modelling.\\

\cite{fletcher_spatial_2018} emphasizes three common approaches to connectivity modelling: structural landscape modelling, resistance-based modelling and graph-based measures. Structural models are simple representations of physical landscape barriers. Resistance based approaches can go beyond structural approaches by abstracting the landscape into a resistance surface representing the level of hindrance to animal movement into the landscape. The parameterization of resistance surfaces depends on the life history and ecology of a species, making it an appealing method for functional connectivity modeling.  This is the method used in this thesis: we derive resistance surfaces for multiple species as input to a habitat suitability analysis. Graph based approaches depend on graph theory and are also common and often seen as complementary to resistance-based approaches. \\

Resistance surfaces can be used directly to calculate the shortest paths between patches with the lowest cumulative resistance (the so-called “least-cost paths”), and derive species-specific networks based on resistance thresholding \citep{lechner_characterising_2017}. An alternative method to treat resistance surfaces is to use circuit theory to estimate resistance distances by simulating random walks. This allows prediction of flow for entire landscapes. According to \cite{fletcher_spatial_2018}, this approach can be useful because it accounts for the redundancy in paths \cite{mcrae_isolation_2006, mcrae_mcrae_2008}. It is also an appealing approach as it allows the identification of “pinch-points” (bottlenecks to animal movement). \\

This thesis makes use of circuit theory to estimate change to potential functional connectivity for 5 umbrella species, under a certain number of land use change scenarios. These 5 species were selected as their connectivity needs have been shown to be representative of the needs of a larger pool of 14 species (see \cite{albert_applying_2017, meurant_selecting_2018}). See table \ref{tab:species} for a detailed description of the 5 species. Next, we introduce the study region on which we carry these analyses. \\ \vspace{2em}

\subsection*{\textit{Study Region and previous efforts \\ \vspace{1em}}}

Montérégie is one of the 17 administrative regions of the Canadian province of Québec. It is situated southeast of the city of Montréal, Québec, on the south bank of the Saint Lawrence, and is therefore located in the ecoregion of the Saint Lawrence Lowlands. Its surface area is \SI{11852}{\km\squared} and is home to about $1.5$ million inhabitants. Agriculture is the predominant land use in the region, and about half of this land area is covered by agricultural land \citep{isq_quebec_2020}. Statistique Québec indicates that between 1995 and 2008, 8\% of natural habitat and 4\% of agricultural lands has been artificialised. These changes are spread over the 15 MRC (“Municipalite Regionales de Comte”) making up Montérégie. MRCs are essentially sub-administrative regions at the level of which land use orientations are decided through two important documents: the PDZA (Plan de Développement de la Zone Agricole) and the SAD (Schéma d'Aménagement et de Développement). These MRCs are made up of 177 municipalities. In addition to MRCs, two First Nation reserves are present in Montérégie. \\

The eastern part of Montérégie is also under the jurisdiction of the Greater Montréal Area (GMA). The ecological connectivity of the GMA and its benefits as a provider of ecosystem services has recently been assessed in a report to the Québec ministry of the environment \citep{rayfield_priorisation_2018}. This study focused on identifying regions of highest ecological connectivity, and therefore of highest priority for the conservation of biodiversity and ecosystem services. Other work by Rayfield et al. (2019, unpublished) has extended the analysis of connectivity to the whole of the Saint Lawrence Lowlands. Prior to this research, the project of “Montérégie Connection”, spearheaded by an interdisciplinary team of researchers from McGill University, studied the provision of ecosystem services in the region and produced important knowledge for conservation planning. \\

Deriving insight from these contributions, multiple stakeholders have become mobilized to set connectivity conservation goals for the region. A number of Non-Governmental NGOs have designed corridor networks for their jurisdiction, with varying degrees of inputs from previous scientific work; these include Nature Action Québec (NAQ), Centre de la Nature du Mont Saint Hilaire (CDN), theCentre d'interprétation du milieu écologique du Haut-Richelieu (CIME), and Appalachian Corridors (CA). These associations are now beginning to work together to design science-driven connectivity conservation plans for the whole region, and are in need of science-based decision support tools. They are also concerned with the social acceptability of conservation, and with the resilience of their plans given expected future land use and climate change. \\ \vspace{2em}

\subsection*{\textit{Limitations of current CCP methods \\ \vspace{1em}}}

The maps produced by previous analyses represent a snapshot of the current state of connectivity in the region. But researchers are developing methods for including future land use and climate change impacts in their analyses. For example, \cite{albert_applying_2017}, used a simple land use and projection from regional climate change models to identify a network of protected habitat robust to climate change for a subset of Montérégie. This was an important step, because ignoring land use change could lead to ill-adapted conservation plans. However, connectivity conservation models rarely include both land use and climate change in their calculations \citep{costanza_landscape_2019}, which and so it remains a limitation in many science-based approaches to  CCP. \\

Similarly, current CCP methods often do not confront the results of the prioritization with the priorities perceived by stakeholders. There is a need for planning tools to allow for this interaction to happen. Those tools need to be able to guide stakeholders through an iterative process of conservation and land use scenario building. Methods are also available to include stakeholder inputs into CCP. Those “participatory research methods” are designed to conduct the research process “with those people whose life-world and meaningful actions are under study” \citep{bergold_participatory_2012}. Those methods often rely on collecting data through a community-driven process during workshops (for an example of a recent local participatory effort in Montérégie see \cite{martins_results_2020}). \\

In this thesis, we attempt to remedy those two limitations. In the first chapter, we combined a connectivity model using circuit theory with a land use and climate change model that uses a combination of statistical modelling and MCMC-based simulations. In the second chapter, we ran a stakeholder workshop to derive conservation priorities and use those results as inputs to derive simple, community driven, alternative conservation scenarios. We compare and contrast the results of these scenarios, and discuss the importance of considering stakeholder inputs to produce a resilient network of protected areas. \\

Next, we introduce the two other key conceptual themes of the thesis: land use change modelling and participatory methods. \\ \vspace{10em}

\subsection*{\textit{Land use change modelling \\ \vspace{1em}}}

A classic spatially defined land use change model is what can be called a “state change model”, where the landscape is divided in a grid and each grid element (pixel) is assigned a state that changes in discrete time \citep{daniel_state-and-transition_2016}. The difficulty of land use change modelling resides in setting the transition probabilities that define the probability of a pixel transitioning to another state (or to remain as is), given its current state and given a set of conditions both intrinsic and extrinsic to that pixel. \\

A simple land use change model will assume that the rules of state changes obey markovian laws: i.e. that the next state will always only depend on the current timestep and not previous timesteps. This is a simplification: in real landscapes, the history of the pixel (beyond its present state) can play a large role in the rules governing state change. Probably due to their simplicity, markovian models such as cellular automata have a long history in land use change models \citep{agarwal_review_2002}. They have been shown to deliver accurate results, notably in urban spread modelling \citep{soares-filho_dinamicastochastic_2002, jokar_arsanjani_integration_2013, iacono_markov_2015}. They are yet another application of markov chains, and are fairly easy to set up and to run. Markov chain models represent a phenomenological approach to land use change, because the model applies the rules of transition without encoding the mechanisms underlying the  change. For example, the model defines that the state “forest” has a probability of 0.23 to transition into urban if it contains at least 2 urban pixels in a 4 pixel radius (this is an adjacency rule), but it is oblivious to the drivers of land use change. \\

In comparison to the phenomenological approach of cellular automata, a mechanistic approach to land use change modelling would consider that land change happens because of multi-scale decision making, and attempts to model those processes directly. Agent-Based Models (ABMs) are probably the best example of a mechanistic approach to land use change modelling. ABMs simulate the behavior of individuals in a network of decision processes whose effects are designed to sum up to the observed change \citep{parker_agent-based_2002, filatova_spatial_2013}. Those models are promising for the future of land use change modelling, but tend to be applied at smaller spatial scales, and tend to rely on expert knowledge or on a rather involved data collection procedure (such as interviews with actors, \cite{taylor_agent-based_2016}). This makes the modelling process more difficult (albeit rewarding) as it requires a deeper understanding of the specific drivers of change in the region. \\

In the absence of a solid body of methodological comparisons between statistical frameworks in land use change modelling, it is difficult to decide on the best framework for raster-based (i.e. based on grids) models. The Random Forest - Cellular Automata model presented in chapter 1 is a  hybrid approach in which a statistical approach is taken (statistical modelling through random forest) and combined with a more mechanistic approach (a cellular automaton with transition targets and spatial dependency rules, CA). The RF-CA model is similar to other popular approaches like the CLUE family of models (Conversion of Land Use and its Effects \cite{verburg_modeling_2002, verburg_combining_2009}), which are some of the most commonly used models in land system science. However, in CLUE, the allocation is determined by the local and regional “demand” for each land use. This demand is a latent variable for the complex and multi-scalar social-environmental processes that determine land use change, and which more mechanistic models attempt to simulate \citep{verburg_combining_2009}. \\

We choose here to present the two steps of probability modelling and allocation as separate because they are methodologically distinct in our model. But it is worthwhile to note that recent efforts in land use change modelling have managed to integrate those steps. Those recent methods can integrate suitability and allocation because the model is capable of encoding both processes at the same time. We can mention important work on modelling land use change with machine learning techniques such as Neural Networks (NNs) \citep{tayyebi_simulating_2013} or exploiting the potential of non frequentist statistics in the case of Belief Networks (BBNs, see \cite{celio_modeling_2014} for a great example). BBNs, like ABMs are promising methods but tend to be applied at smaller spatial scales. \\

This thesis therefore makes use of a hybrid modeling approach. We combine this model with scenarios who were derived from a participatory method, which we introduce next. We end with simple predictions as to the results of our modelling effort. \\ \vspace{2em}

\subsection*{\textit{Participatory methods in Connectivity Conservation \\ \vspace{1em}}}

CCP researchers are increasingly recognizing landscapes are social ecological systems. A social-ecological system (SES, \cite{ostrom_general_2009}) can be conceived  as the set of human and non-human actors, the set of natural habitats they inhabit and resources and use, and the set of interactions that are maintained between all the components of the system. SESs thus form complex and integrated aggregates of interactions \citep{hinkel_enhancing_2014}. Those interactions also impact governance, the process by which actors in power establish rules and laws \citep{bissonnette_comparing_2018}. Conservation and land use change are inherently social practices, with repercussions of both social and ecological nature. Connectivity conservation science is therefore called to take an increasingly social-ecological approach to modelling and planning. \\

A wide range of socio-ecological methods have been used in general conservation planning. In this thesis, we focus on participatory mapping and scenario-based approaches. Participatory mapping involves leveraging the knowledge that stakeholders hold about the geography of their landscape, in order to achieve a community-driven objective  \citep{plieninger_assessing_2013}. Scenarios are a flexible approach to forecasting the range of possible futures in landscapes \citep{peterson_scenario_2003}. In this thesis, we derive simple land use change and conservation scenarios from the results of a participatory mapping workshop. We integrate those scenarios with connectivity modelling in combination with two “warming” scenarios (Baseline and RCP 8.5) and one “control” scenario (Historic). \\

Depending on the scenario, the model generates different forest change dynamics. The Historic scenario assumes no change in forest composition and stands as our control. While the baseline scenario assumes no further warming and status quo forest change dynamics, the RCP 8.5 scenario is one of the most severe climate change scenarios, and the trends observed in the baseline scenario are exacerbated \cite{IPCC}. We also derive two simple land use change scenarios: Business as usual (BAU) and Reforestation. Those two scenarios are meant to represent two extremes in land use change, and it is likely that the most probable trajectories for our landscape will be within those two boundaries. In combination with the climate scenarios, this brings our total scenario number to 6. \\

What predictions can be made as to how connectivity will be affected under our model? There has been a significant research effort to study and predict forest dynamics under climate change in Québec. It has been shown that the distributions of many of the most abundant trees occurring in the region are out of equilibrium with climate at the margin of their range \citep{talluto_extinction_2017}. Some papers points to existing and expected lags in the response of forest to climate change \citep{savage_elevational_2015}, and others have commented on how disturbances and forest management practices are likely to fasten and facilitate range shifts \citep{leithead_northward_2010, boulanger_climate_2019, vieira_paying_2020}. A major prediction that can be made is that although boreal type (coniferous) forest is likely to respond more slowly to climate range than deciduous forest, both coniferous and mixed population forests in southern Québec are likely to be replaced by deciduous stands. \\

We can therefore derive different predictions as to how connectivity will be affected for each species under our model. Three main predictions can be made. First, under the BAU scenario, we can expect that our land use change model will simulate an decrease in connectivity change for our focal species. We can also expect connectivity to maintain itself under the Reforestation scenario. Finally. We can expect climate change to impact more strongly species that prefer coniferous  and mixed forest types. \\

This thesis has two chapters.  In the first chapter, we combined a connectivity model, using circuit theory, with a land use and climate change model that uses a combination of statistical modelling and MCMC-based simulations. In the second chapter, we ran a stakeholder workshop to derive conservation priorities and use those results as inputs to derive simple, community driven, alternative conservation scenarios. We also compare and contrast the results of these scenarios and discuss the importance of considering stakeholder inputs to produce a resilient network of protected areas. \\
}%
\Intro

%  Chapter I
\include{thesis/chap_1}

\SetLinkName{Linking Statement}
\SetLinkText{In Chapter I, we integrated a land use model and a climate change model with a habitat suitability and connectivity model. We then explored outcomes under different scenarios for land use and climate change. We showed how different species might react to changes in land use and climate by analyzing the resulting flow surfaces. We provided  a reproducible yet perfectible methodology that takes into account important drivers like land use and climate change. We applied these methods to the case of the Montérégie in Québec. We demonstrate the power of simple scenarios to identify the boundaries of the prediction space for potential functional connectivity. However, we recognize the need for scenarios that take into account the needs and perceptions of stakeholders of the region. In Chapter 2, we take a first step toward deriving the basis for such community-driven scenarios by presenting the results of a community workshop organized between multiple stakeholders in Montérégie. We use those results to derive simple land use change scenarios and use those scenarios to constrain the integrated model from Chapter 1. These scenarios can serve as a basis with which to iterate a community-driven process of conservation scenario building. We derive the same analysis for the species described in Chapter 1 and compare and contrast the results between scenarios. This integrated approach has the potential to support ongoing connectivity conservation planning in the south of Québec.}
\Link

%  Chapter 2
\include{thesis/chap_2}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%        General Discussion                     %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Clearly state the rationale and objectives of the research.
\newcommand*{\SetDiscName}[1]{\renewcommand*{\ETDDiscName}{#1}}%
\newcommand*{\ETDDiscName}{Disc}%

\newcommand{\SetDiscText}[1]{\renewcommand*{\ETDDiscText}{#1}}%
\newcommand*{\ETDDiscText}{Disc text goes here!}%

\newcommand{\Disc}{%
    \begin{simpleenv}{}{}{}{}%
    \pagestyle{plain}%
  \setboolean{SetDSpace}{false}%
    \GoSingle%
    \begin{QZ@Cent}%
          \bfseries{\ETDDiscName}
      \end{QZ@Cent}%
    \vspace*{0.5in}
      \par%
      \GoDouble%
      \setlength{\parskip}{-2em}
      %\chapter*{\ETDDiscName}%
      \phantomsection\section*{}
      \addcontentsline{toc}{extrachapter}{\ETDDiscName}%
      \ETDDiscText%
    \end{simpleenv}%
    \setboolean{SetDSpace}{true}}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Discussion
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\SetDiscName{General Discussion}%

\SetDiscText{

The two chapters of this thesis alleviate two limitations to Connectivity Conservation Planning methods identified in the general introduction: the lack of methodological integration between connectivity, land use change and climate change models; and the need for stakeholder inputs in the planning process. \\

In chapter 1, we integrated circuit-theory based connectivity models with a simple land use and scenarios from a climate change model based on habitat suitability for 5 umbrella species. We also presented patterns of land use change in our study region for the 1990-2010 period, and demonstrated that our land use change model captures these general patterns across the landscape. We analysed changes in connectivity under simple land use and climate change scenarios, and showed that different species have different responses, depending on the forest type they depend on, and on whether this forest type is under threat from land use change, climate change or both. The American marten showed the most complex response because it is dependent on coniferous forest, which is the most at risk under both drivers. We also used a pattern-detection method (SURF) to identify pinch-points and salient connectivity flow features. Although the number of features detected is species-specific, we showed a mostly consistent pattern of change over time: reforestation seems to simplify the flows of movement on the landscape. More research to understand the links between flow complexity and SURF measures is needed. \\

Then, in chapter 2, we capitalized on the progress in chapter 1 by including the results of a stakeholder workshop during which we mapped areas of consensus among stakeholders that were deemed important for connectivity in the region. During this participatory process, we also identified key assets (such as the drive and momentum of conservation NGOs in the region) and obstacles (such as the predominance of an agricultural matrix) to connectivity conservation in the region. We derived simple conservation scenarios based on priorities identified during the participatory mapping exercise. We simulated connectivity changes for our five species under these scenarios and made clear the complexity of restoring connectivity to cater to species with different needs while accounting for multiple stakeholder interests. \\

I now address some general limitations that concern both chapters. We argued that using markov-chain monte carlo (MCMC) based modelling for land use change allowed us to take into account the uncertainty of land use change simulations. The principle is to run the model multiple times (iterations), resulting in slightly different simulations every time. This allowed us to see that our model was quite constrained and that over 10 iterations we produced narrow confidence intervals when analysing the resulting connectivity; this was true for both chapters. However, there remains a large amount of uncertainty unaccounted for in the model. STSIM required a large number of parameters (for example, neighboring parameters, see \ref{tab:neigh_rules}) that were chosen and fixed to a certain value. To correctly account for uncertainty in those parameters, we could decide on a distribution for those parameters and draw from this distribution at each iteration. This is possible in STSIM, which allows the user to specify distributions. However this brings its own set of drawbacks. First, in order to get a representative sample from this distribution, more iterations would need to be computed, which would increase computing time. Second, the shape of the distribution for parameters like neighboring rules is not easy to determine a priori. This last point could potentially be alleviated by sampling the distribution from historical land use and land cover change data ; however, the question of how to conduct this sampling appropriately, for instance to account for spatial autocorrelation, would require more research. \\

Our approach to modeling functional connectivity was focused on potential functional connectivity (PFC). In both chapters, we use circuit theory to model PFC. A limitation of PFC modelling is that what the model produces is an informed prediction about potential movement, which is still in need of statistical validation with data derived from movement events observed on the landscape. Therefore, the quality of the data used for parameterization of the resistance surface is extremely important for the accuracy of the results. In addition, circuit theory has been criticized as a PFC approach. \cite{fletcher_spatial_2018} mention a set of issues: first, animals might not move in a way that can realistically be modelled as random walks. It is computationally more expensive than least-cost path analysis - although the recent Julia version of Circuitscape is faster and more efficient than previous iterations of the software \citep{anantharaman_circuitscape_2020}. They also argue that results can be difficult to interpret across large landscapes, given the fact that flow tends to be predicted as being diffuse. Finally, they mention an issue that “has been neglected in the ecological literature but has been repeatedly acknowledged in the network literature.” According to \cite{luxburg_hitting_2014}, on large graphs, the resistance distance  between patches converges on the local properties at a cell level, making the use of random walks less helpful. \\

However, connectivity research can be hopeful in its ability to move beyond those limitations. Resistance-based models have been validated with genetic data \citep{marrotte_landscape_2014} and occurrence data (\cite{gray_landscape_2016}) and movement data \citep{lapoint_animal_2013}. Although genetic data alone cannot be expected to predict demographic connectivity without information on local demographic rates \citep{lowe_what_2010}, it can provide important information on connectivity when combined with other data sources. Connectivity modelling therefore needs to move toward new, integrative, and data-rich approaches. One example is combining multi-scale resistance surfaces from multiple data types, as \cite{zeller_multi-level_2017} did for pumas (\textit{Puma concolor}) in southern California. The role of novel forms of modelling, such as Individual-BasedModels (IBMs, a form of Agents-Based Models) is another potential direction for connectivity research: they can add realism to resistance based and circuit theory models and bring attention to mechanisms overlooked by those models, like mortality \citep{day_individual-based_2020}. In all cases, these approaches are data-intensive and seem to point to a need for more data on animal movements. \\

Finally, in both chapters we pointed out the limitations in the way we defined land use change and conservation scenarios. Although the goal was to derive simple and extreme cases to show the range of variations in species’ responses, this method will remain unsatisfactory for stakeholders, until it is inscribed in an interactive process of community-driven scenario building (see discussion at the  end of Chapter 2). Both chapters outline the design principles for a tool that could be used to facilitate this iterative process, by providing information on the consequence of “what if” scenarios on potential functional connectivity. It is important to ask what framework would best allow us to operationalize this tool and allow researchers to provide guidance to stakeholders during a decision making process. In the following, we discuss the importance of the Spatialized SES (SpaSES, \cite{williamson_spatially_2018}), and Ecosystem Services Bright Spots \citep{frei_bright_2018} frameworks to guide the design of connectivity conservation planning tools. \\

Having recognized that conservation tools often fail to consider social factors affecting their implementation, \cite{williamson_spatially_2018} developed the SpaSES framework as a recontextualization of Ostrom’s framework applied to conservation tools. In the original SES framework, SES variables are numerous, and acquiring data on them for decision making at large spatial scales is often impossible. Therefore, \citeauthor{williamson_spatially_2018} bundled them in three categories, forming a three dimensional space across which probability of conservation action can be modelled. First, Ecological Value (EV) describes how important an habitat is for conservation. Then, Social Willingness (SW) and Institutional Capacity (IC) account for social factors. Social willingness variables essentially take stock of the level of social acceptability to conservation, based on theories of individual and collective behavior. Institutional Capacity variables measure governance systems, institutional fit, and factors that influence the capacity of institutions of different types (government, action groups) to reach conservation goals. \citeauthor{williamson_spatially_2018} models probability of connectivity as a function of those three latent variables (EV, SW, IC), for which they suggest a number of proxies that will vary depending on study region and conservation objectives. \\

For our study region, an appropriate measure of EV could be the results of a prioritization of habitat patches based on habitat suitability and connectivity at different scales under land use and climate change, such as in \cite{albert_applying_2017}. But \citeauthor{williamson_spatially_2018} notes that this type of prioritization effects “generally assume that stakeholders are interested in an optimal solution to resource management concerns and pursue those solutions rationally, ” an assumption they argue is not confirmed by SES research. One way to tackle this rational optimization fallacy is to use multi-criteria weightings systems to avoid optimizing with only one outcome in mind. This justifies the need for taking other dimensions of conservation action (SW and IC) into account. Participatory methods such as the one we employed could be an appropriate way of developing measures for these two axes, as stakeholders most likely know what variables would be better representative of each axis. An iterative process of scenario building could be initiated by efforts to define (1) multiple scenarios based on stakeholder’s perceptions of planning priorities, (2) a measure of SW for conservation, partially based on alignment with the different scenarios developed and (3) a measure of stakeholders’s confidence in IC in the region to conduct different conservation plans, and an idea of what the needs are in IC for such plans to succeed. This framework would lead to better estimates of where conservation is both critical and implementable in the region. \\

Workshops might also be a good tool to build up SW and IC themselves and increase the probability of conservation action, which will in a large part depend on the obstacles identified in any workshop. According to our results, the most obvious of these obstacles in our region is the predominance of agricultural land use. The role of the agricultural matrix in fragmented landscapes became under scrutiny when the SLOSS debate in ecology \citep{diamond_island_1975, simberloff_island_1976}, was for a time renewed in the land sparing versus land sharing debate \citep{fischer_land_2014}. The concept at the base of these dichotomies, is that smaller fragments experience higher extinction, and that connectivity can help by restoring immigration between fragments. \cite{perfecto_natures_2019} points out that focusing on connectivity and immigration rates in fragmented landscapes actually “means focusing on the agroecosystems, which in turn means examining what is happening socio-politically in the agricultural matrix” \citep{perfecto_natures_2019}. Both the “single large/several small” and “sharing/sparing” oppositions are now seen as false dichotomies, rendered obsolete by the SES literature. \cite{bennett_changing_2017} agrees and argues that those dichotomies have tended to look at agroecosystems too simply, mostly within the angle of food production. They propose we go beyond those dichotomies by acknowledging the multiple facets of agroecosystems and considering the complexity of the agro-ecological matrix. \\

One way to embrace the complexity of the matrix is to carefully look at the relationship between the structure of a landscape and its impact on connectivity and associated ecosystem functions, such as carbon storage or pollination \citep{mitchell_linking_2013}. Both Models \citep{mitchell_strong_2015} and empirical studies \citep{mitchell_forest_2014} have shown that landscape complexity can modify biodiversity, connectivity and ecosystem service supply. While earlier studies focused on single services, more research is looking into how connectivity of forest fragments impact a wide diversity of services\citep{mitchell_forest_2014}. For example, crop heterogeneity is linked to higher biodiversity \citep{sirami_increasing_2019}. Agricultural matrices with smaller average field sizes tend to be more biodiverse \citep{fahrig_farmlands_2015}. These phenomena have been attributed to an increase in connectivity due to the presence of more semi-natural habitats. These contributions are important and should be considered for land use planning. They are appealing for the design of connectivity conservation tools, because they point out clear management scenarios that could be tested through models, and can make use of the increasing availability of high resolution remotely sensed data. \\

But, as we quoted from \cite{williamson_spatially_2018} above, this view tends to assume that stakeholders want to pursue rational and optimal designs in their approach to conservation. There is a tendency to take a single-faceted approach to the optimisation problem, focusing on the trade-off between food production and the support for biodiversity. An alternative way is through the lens of ecosystem services  \cite{bennett_changing_2017}. Ecosystem services refer to the diversity of benefits that ecosystems provide to, and the many ways in which they support, human systems. Specifically, the concept of “bright spots'' has potential for guiding the discussion beyond dichotomies and single faceted approaches, by focusing on multifunctionality in agroecosystems. In a landscape, “bright spots” are areas exceeding expectations for goals, such as agro-ecological multifunctionality, biodiversity, or connectivity \citep{frei_bright_2018}. They show promise as a tool  to identify levers of change, and guide stakeholder discussion. Indeed, an approach that focuses on identifying bright spots for the provision of those ecosystem services can be valuable to an interactive process of stakeholder engagement. They offer starting points to work with to design scenarios. They are concrete examples to reason around what bundles of ecosystem services are desirable (elements of SW) for the conservation of ecological connectivity and which types of governance is needed to work toward them (elements of IC). In our study region, this framework is already being tested: \cite{frei_bright_2018} identified bright, dark and average spots within Montérégie for both landscape multifunctionality and avian biodiversity. They showed important spatial properties of bright spots, for instance that they were “neither spatially congruent nor in conflict”.  Connectivity could be fully integrated with this approach is two ways: first, it would be interesting to see if current bright spots are congruent with connectivity, in order to asses the mismatch between connectivity and other bundles of ecosystem services. One example is that of bee diversity which was found to be higher in connected habitats in Montérégie \citep{martins_complementary_2018}. Second, areas that are prioritized as important for connectivity under models such as the ones we developed in chapter 1 could be used in the identification of bright spots. \\

\citeauthor{frei_bright_2018} also pointed out a range of challenges for the identification of bright spots, the development of multicriteria measures being one of them. Connectivity measures are already being developed with  multifunctionality in mind. For instance, \cite{rayfield_priorisation_2018}, assessed whether certain habitat networks in the Montérégie would be more resilient to future change using several multi-scale metrics of connectivity. Another example is \cite{fletcher_towards_2019} who developed an unifying model for connectivity that takes into account multiple types of mobility. Connectivity is therefore a field that fits in the multicriteria framework of bright spots, as it itself moves in this direction.  Bright spots show promise for being integrated in more complex connectivity conservation planning tools. Finally, bright spots are intuitive, being based on statistical deviation from a given distribution and are therefore always relative to the context in which they are implemented. They are suited to the type of participatory framework we implemented in chapter 2: bundles of ecosystem services that are important to stakeholders can be identified and bright spots can be discussed, all in the iterative participatory setting we detailed above. \\

\vspace{3em}

\subsection*{\textit{Conclusion \\ \vspace{2em}}}

The integration of models of multiple drivers of changes in ecological connectivity with stakeholder inputs into connectivity models will contribute to the design of better connectivity conservation planning tools. In this thesis, we contributed some design elements for those tools, such as a methodology to account for uncertainty and an approach to integrating the different  priorities among local stakeholders. Future research will need to engage in an iterative process of regional stakeholder engagement, possibly including a scenario building exercise. This process will be all the more successful if it is made within a coherent framework that addresses the social-ecological complexity of fragmented landscapes, such as SpaSES and Bright spots. These frameworks will guide discussion and facilitate the visualization of positive alternative scenarios to business as usual. \\
}%

\Disc%

% Appendix
%\ETDAppendix{Chapter I Supplementary Material}{
%\input{appendix_1}
%}
\include{thesis/appendix_1}
\include{thesis/appendix_2}
\include{thesis/appendix_3}

%\ETDAppendix{References}{
%\printbibliography
%}
\printbibliography[heading=bibintoc, section=0, title={General Bibliography \vspace{1em}}]

\end{document}