### 2022-08-28 docs/msc_thesis_figures.Rmd
add new code

---
title: "Msc Thesis Figures"
author: "Valentin Lucet"
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'index.html')) })
output:
  html_document:
    df_print: paged
    code_folding: hide
    self_contained: yes
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: yes
      smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: '5'
---

**This regroups most of the figures in the thesis (appendix figures are not included, except for the ROC resample figures, the BAU simulated land use change and the radar graph for change in flow for both chapters).**

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages({
  library(vegan)
  library(dendextend)
  library(ggbiplot)
  library(ggnewscale)
  library(raster)
  library(sf)
  library(tidyverse)
  library(ggplot2)
  library(gganimate)
  library(transformr)
  library(animation)
  library(rasterVis)
  library(ggmap)
  library(gifski)
  library(gtools)
  library(fmsb)
  library(RColorBrewer)
  library(scales)
  library(ggradar)
  library(tidymodels)
  library(patchwork)
  library(ggplotify)
  library(gt)
  library(webshot)
  library(wesanderson)
  library(cowplot)
})

theme_set(theme_minimal())
theme_update(legend.position = c(0.85, 0.20),
             legend.box = "vertical",
             legend.background = element_blank(),
             legend.box.background = element_rect(colour = "black"),
             panel.border = element_rect(fill = NA),
             legend.title = element_text(size = 22),
             legend.text = element_text(size = 22),
             plot.title = element_text(size=22),
             plot.subtitle = element_text(size=22),
             axis.text.x = element_text(angle=65, vjust=0.6, size =22),
             axis.text.y = element_text(size =22),
             strip.text.x = element_text(size = 22),
             strip.text.y = element_text(22),
             axis.title=element_text(size=22))

# Set the REPROG_FIGS_ONLY
REPRO_FIGS_ONLY <- Sys.getenv("REPRO_FIGS_ONLY")
if(REPRO_FIGS_ONLY == ""){
  REPRO_FIGS_ONLY <- FALSE
} else {
  REPRO_FIGS_ONLY <- TRUE
}
```

### Data prep

#### Cleaning and formatting

We start by preparing the data from the outputs of scripts `2.6`, `7.1` and `8.1`. The results table shown as output summarised all simulation efforts for the thesis.

```{r data_prep, message=FALSE, warning=FALSE, paged.print=TRUE}
# Land use data 
trans_df <- read_csv("../outputs/final/land_use_trans_df.csv")
vals_df <- read_csv("../outputs/final/land_use_vals_df.csv")

# Results dir + mun and mrc shapefile
# mun <- st_read("../data/mun/munic_SHP_clean.shp", quiet = TRUE)
# mrc <- st_read("../data/raw/vector/mrc_SHP/mrc_polygone.shp",  quiet = TRUE)
# mrc.mont <- mrc %>% filter(MRS_NM_REG=="Montérégie") 
# mrc.mont.reproj <- st_transform(mrc.mont, crs = st_crs(mun))

# Classes
classes <- read_csv("../config/rcl_tables/land_use/recode_table.csv")
forest_classes <- read_csv("../config/rcl_tables/land_use/recode_table_forest.csv") %>% 
  rename(new_code = "ID", new_class = "Name")

classes_unique <- unique(classes[, c("new_code", "new_class")]) %>% 
  bind_rows(forest_classes)

# results key
sce_code_vec <- 
  as.vector(t(outer(c("BAU-", "BAUR-", "CP-", "CPR-", "CPTR-"), 
                    c("Hist", "Base", "RCP8"), paste0)))
sce_code_vec_run <- 
  rep(c("BAU", "BAUR", "CP", "CPR", "CPTR"), each=3)

results_clean <- read_csv("../outputs/final/stsim_run_results.csv") %>% 
  dplyr::select(scenarioId, name) %>% 
  mutate(name =  gsub(.$name, pattern = " \\(.+\\)", replacement = "")) %>% 
  mutate(sce =  paste0("sce_", .$scenarioId)) %>% 
  mutate(chapter = c("none", rep("both", 3), rep("chap_1", 3), rep("chap_2", 9))) %>% 
  mutate(splitted = str_split(name, " \\| ")) %>% 
  mutate(climate = unlist(map(splitted, ~unlist(.x[2])))) %>% 
  mutate(run = unlist(map(splitted, ~unlist(.x[1])))) %>% 
  dplyr::select(-splitted) %>% 
  replace_na(list(climate = "none")) %>% 
  mutate(code = c("Control", sce_code_vec)) %>% 
  mutate(code_run = c("control", sce_code_vec_run))

# Final extracted datasets
df_final <- read_csv("../outputs/final/final_df_current_density_part1.csv") %>%
  bind_rows(read_csv("../outputs/final/final_df_current_density_part2.csv")) %>% 
  mutate(timestep = (timestep*10)+1990, source = "model", zone = as.character(zone))
df_final_origin <- read_csv("../outputs/final/final_df_origin_current_density.csv") %>%
  mutate(timestep = timestep*10+1990, source = "model")

# Stratum key
stratum_key <- read_csv("../config/stsim/SecondaryStratum.csv") %>%
  mutate(ID = as.factor(ID)) %>%
  rename(zone=ID, MUS_NM_MUN=Name) %>%
  mutate(zone = as.character(zone)) %>% 
  left_join(df_final, by="zone") %>%
  filter(MUS_NM_MUN!="Not Monteregie")

# Sum all municipalities
df_summarised <- df_final %>%
  group_by(sce, timestep, species, iteration) %>% 
  summarise(sum_cur = sum(current)) %>% ungroup %>%
  mutate(source = "model")
df_origin_summarised <- df_final_origin %>%
  group_by(timestep, species) %>% 
  summarise(sum_cur = sum(mean)) %>% ungroup %>% 
  mutate(sce = "sce_0", source = "observation")

# Joined dataset
joined <- full_join(df_summarised, df_origin_summarised, 
                    by=c("sce", "source", "species", "timestep", "sum_cur")) %>% 
  left_join(results_clean, by = "sce")  %>% 
  replace_na(list(climate = "none", run = "historic run")) %>% 
  # Make factors
  mutate(sce = as.factor(sce), run = as.factor(run), sum_cur = 10*(sum_cur),
         climate = factor(climate, levels = c("none", "baseline",
                                              "RCP 8.5")),
         run = factor(run, levels = c("historic run", "BAU run", 
                                      "BAU run + corrs protection", "BAU run + ref", 
                                      "BAU run + corrs protection + ref",
                                      "BAU run + corrs protection + ref TARGETED")))

# Histogram data 
hist_original <- read_csv("../outputs/final/final_values_output_original.csv")
hist_true <- read_csv("../outputs/final/final_values_output_TRUE.csv") %>% 
  mutate(sce = "TRUE")
histograms <- read_csv("../outputs/final/final_values_output.csv") %>% 
  bind_rows(hist_original) %>% 
  bind_rows(hist_true) %>% 
  left_join(results_clean, by="sce")

# SURF Data
surf <- read_csv("../surf/surf_output.csv") %>% 
  mutate(timestep = timestep*10+1990) %>% 
  left_join(results_clean, by=c("scenario"="scenarioId")) %>% 
  replace_na(list(climate = "none", run = "historic run")) %>% 
  # Make factors
  mutate(sce = as.factor(sce), run = as.factor(run),
         climate = factor(climate, levels = c("none", "historic",
                                              "baseline", "RCP 8.5")),
         run = factor(run, levels = c("historic run", "BAU run", 
                                      "BAU run + corrs protection", "BAU run + ref", 
                                      "BAU run + corrs protection + ref",
                                      "BAU run + corrs protection + ref TARGETED")))

# Bar plot data
bar_data <- read_csv("../outputs/final/final_bar_plot_data.csv")

results_clean
```

<!-- Preparing for spatial plotting (time consuming).

```{r data_prep_spatial, eval=FALSE, message=FALSE, warning=FALSE}
# Pivoted dataset for mapping
# df_final_fordiff_pivoted <- df_final %>% 
#   left_join(results_clean, by = "sce")  %>% 
#   group_by(zone, sce, species, timestep, source, chapter) %>% 
#   summarise(current = mean(current)) %>% ungroup %>% 
#   filter(timestep %in% c(2010, last(unique(stratum_key$timestep)))) %>% 
#   pivot_wider(names_from=timestep, values_from=current) %>% drop_na() %>% 
#   rename(before=last_col(offset = 1), after=last_col()) %>% 
#   left_join(stratum_key, by=c("zone", "sce", "species", "source")) %>% 
#   filter(MUS_NM_MUN!="Not Monteregie") %>% 
#   mutate(change=((after-before)/before)*100) %>% 
#   left_join(mun, by="MUS_NM_MUN") %>% 
#   st_as_sf()
# 
# df_final_fordiff_pivoted <- 
#   st_transform(df_final_fordiff_pivoted, crs=4326)
```

-->

### Figures

### Land use trends

We prepare the data for ordination and compute clustering (figures not displayed here).

```{r land_use_trends_prep, message=FALSE,  warning=FALSE, out.width="100%"}

# Trans

trans_mat_1990to2000 <- 
  trans_df %>% 
  dplyr::filter(Trans == "1990to2000", From!=To) %>% 
  mutate(trans_type = paste0(From,"to", To)) %>% 
  dplyr::select(-Trans, -From, -To) %>%
  spread(key=trans_type, value=Freq) %>% 
  rename(Mun=Municipality) %>% 
  arrange(Mun)
rownames(trans_mat_1990to2000) <- trans_mat_1990to2000$Mun
trans_only_mat <- Filter(function(x)!all(is.na(x)), trans_mat_1990to2000) %>% 
  mutate(Mun = paste0(str_sub(Mun,1,3), 1:length(Mun)))
trans_only_mat[is.na(trans_only_mat)] <- 0

trans_only_mat_noZero <- trans_only_mat[, c(TRUE, colSums(select(trans_only_mat, -Mun))>0)]
trans_only_mat_noZero <- 
  trans_only_mat_noZero[, c(1, order(colSums(select(trans_only_mat_noZero, -Mun)), decreasing=T)+1)]
rownames(trans_only_mat_noZero) <- trans_only_mat$Mun

# Values

values_mat_1990to2000 <- 
  vals_df %>% 
  dplyr::filter(Year == 2010) %>%
  dplyr::select(-Year, -Unclassified) %>% 
  rename(Mun=Municipality) %>% 
  arrange(Mun) %>% 
  select(Mun, everything())
rownames(values_mat_1990to2000) <- values_mat_1990to2000$Mun
values_only_mat <- Filter(function(x)!all(is.na(x)), values_mat_1990to2000) %>% 
  mutate(Mun = paste0(str_sub(Mun,1,3), 1:length(Mun)))
values_only_mat[is.na(values_only_mat)] <- 0

values_only_mat_noZero <- values_only_mat[, c(TRUE, colSums(select(values_only_mat, -Mun))>0)]
values_only_mat_noZero <- 
  values_only_mat_noZero[, c(1, order(colSums(select(values_only_mat_noZero, -Mun)), decreasing=T)+1)]
rownames(values_only_mat_noZero) <- values_only_mat$Mun

```

Clustering for land use change transitions.

```{r land_use_trends_clust, message=FALSE,  warning=FALSE, out.width="100%"}
# Trans

trans_dist <- dist(decostand(trans_only_mat_noZero[,2:5], "hel"))
trans_clust <- hclust(trans_dist, method = "ward.D2")

trans_dendro <- as.dendrogram(trans_clust)
trans_memberships <- cutree(trans_clust,4)
# trans_memberships_names <- c("Urb-Def","AgEx-Def",
#                              "AgEx-DeTree","Urb-DeAg")[trans_memberships]
labels_colors(trans_dendro) <- trans_memberships[trans_clust$order]
#plot(trans_dendro)
```

Clustering for land use values.

```{r land_use_trends_clust_2, message=FALSE,  warning=FALSE, out.width="100%"}
# Vals

values_dist <- dist(decostand(values_only_mat_noZero[,2:5], "hel"))
values_clust <- hclust(values_dist, method = "ward.D2")

values_dendro <- as.dendrogram(values_clust)
values_memberships <- cutree(values_clust,5)
# values_memberships_names <- c("Forest - Dominant","Forest - Agriculture",
#                               "Agriculture - Dominant"," Urban - MD", 
#                               "Urban - HD")[values_memberships]
labels_colors(values_dendro) <- values_memberships[values_clust$order]
#plot(values_dendro)

# Figure out % of urban land in smallest group

# grp <- as.numeric(names(sort(table(values_memberships))[1]))
# urb <- values_only_mat_noZero[which(values_memberships == 4),]
```

Ordination of land use data (proportions) for municipalities

```{r land_use_trends_PCA, message=FALSE,  warning=FALSE, out.width="100%"}

# c('#1f78b4','#33a02c','#b2df8a','#a6cee3')

# PCA + map

the_dpi = 300

# Trans

trans_minim.mat_noZero <- decostand(trans_only_mat_noZero[,2:5], "hel")
# "41to51"  "51to21"     "45to51"     "41to21"
names(trans_minim.mat_noZero) <- c("Fo_to_Ag", "Ag_to_Urb", "Tre_to_Ag", "Fo_to_Urb")

Trans_Pca <- prcomp(trans_minim.mat_noZero)

trans_biplot <- ggbiplot(Trans_Pca, scale = 1,
                         obs.scale = 1, var.scale = 1,
                         groups =  as.factor(trans_memberships),
                         ellipse = T, circle = TRUE) +
  scale_color_manual(name = "Profiles",
                     values = c('#1f78b4','#33a02c','#b2df8a','#a6cee3'), 
                     labels = c("Urbain Spread/Deforestation",
                                "Agricultural Expansion/Deforestation",
                                "Agricultural Expansion/Fragmentation", 
                                "Urban Spread/Agricultural Loss")) +
  theme(panel.background = element_rect(fill = "white", colour = "grey50"),
        legend.direction = 'horizontal', legend.position = 'top') +
  ggtitle("PCA of Municipality profiles") +
  theme(legend.box.background = element_blank(),
        panel.border = element_rect(fill = NA),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8))
#trans_biplot

ggsave(plot = trans_biplot, "../thesis/figures/PCA_trans_profiles.png", 
       height = 8, width = 10, dpi = the_dpi)

# trans_biplot
```

![](../thesis/figures/PCA_trans_profiles.png)

Geographical distribution of the 5 profiles identified.

```{r land_use_trends_map, message=FALSE,  warning=FALSE, out.width="100%"}

trans_member.table <- data.frame(MUS_NM_MUN = rownames(trans_mat_1990to2000)) %>% 
  mutate(code = paste0(str_sub(MUS_NM_MUN,1,3), 1:length(MUS_NM_MUN)), 
         membership = as.factor(trans_memberships))

# tbl.names <- table(mun.sub.18.clean$MUS_NM_MUN)
# more.than.one <- tbl.names[tbl.names>1]

mun %>% 
  select(MUS_NM_MUN) %>% 
  group_by(MUS_NM_MUN) %>%
  dplyr::summarise(MUS_NM_MUN_New = first(MUS_NM_MUN)) ->
  mun.sub.18.clean.merged

trans_mun.sub.18.clean.mod <- 
  left_join(mun.sub.18.clean.merged, trans_member.table, by="MUS_NM_MUN")

map_trans <- ggplot() + geom_sf(data = trans_mun.sub.18.clean.mod, 
                                aes(fill = as.factor(membership))) +
  scale_fill_manual(name = "Profiles",
                    values = c('#1f78b4','#33a02c','#b2df8a','#a6cee3'), 
                    labels = c("Urbain Spread/Deforestation",
                               "Agricultural Expansion/Deforestation",
                               "Agricultural Expansion/Fragmentation", 
                               "Urban Spread/Agricultural Loss")) +
  geom_sf(data = mrc.mont.reproj, fill=NA, color=alpha("black",1)) +
  theme_bw() +
  theme(legend.justification=c(0,1), legend.position=c(0.02,0.98))

ggsave(plot = map_trans, "../thesis/figures/transition_prof_map.png", 
       height = 8, width = 12, dpi = the_dpi)

#map_trans
```

![](../thesis/figures/transition_prof_map.png)

Ordination of land use transition data for municipalities.

```{r land_use_trends_PCA_2, message=FALSE,  warning=FALSE, out.width="100%"}

# Vals

# c("#B2DF8A", "#33A02C", "#FF7F00", "#CAB2D6", "#6A3D9A")

values_minim.mat_noZero <- decostand(values_only_mat_noZero[,c(2:5)], "hel")

Values_Pca <- prcomp(values_minim.mat_noZero)

vals_biplot <- ggbiplot(Values_Pca, scale = 1,
                        obs.scale = 1, var.scale = 1,
                        groups = as.factor(values_memberships),
                        ellipse = T, circle = TRUE) +
  scale_color_manual(name = "Profiles",
                     values = c("#33A02C", "#B2DF8A", "#FF7F00", "#CAB2D6", "#6A3D9A"), 
                     labels = c("Forest - Agriculture",
                                "Forest - Dominant",
                                "Agriculture - Dominant",
                                "Urban - Medium Density", 
                                "Urban - High Density"))+
  theme(panel.background = element_rect(fill = "white", colour = "grey50"),
        legend.direction = 'horizontal', legend.position = 'top') +
  ggtitle("PCA of Municipality profiles")  +
  theme(legend.box.background = element_blank(),
        panel.border = element_rect(fill = NA),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10))

ggsave(plot = vals_biplot, "../thesis/figures/PCA_data_profiles.png", 
       height = 8, width = 10, dpi = the_dpi)

#vals_biplot
```

![](../thesis/figures/PCA_data_profiles.png)

Geographical distribution of the 4 change profiles identified

```{r land_use_trends_map_2, message=FALSE,  warning=FALSE, out.width="100%"}

values_member.table <- data.frame(MUS_NM_MUN = rownames(values_mat_1990to2000)) %>% 
  mutate(code = paste0(str_sub(MUS_NM_MUN,1,3), 1:length(MUS_NM_MUN)), 
         membership = as.factor(values_memberships))

values_mun.sub.18.clean.mod <- 
  left_join(mun.sub.18.clean.merged, values_member.table, by="MUS_NM_MUN")

map_vals <- ggplot() + geom_sf(data = values_mun.sub.18.clean.mod, 
                               aes(fill = as.factor(membership))) +
  scale_fill_manual(name = "Profiles",
                    values = c("#33A02C", "#B2DF8A", "#FF7F00", "#CAB2D6", "#6A3D9A"), 
                    labels = c("Forest - Agriculture",
                               "Forest - Dominant",
                               "Agriculture - Dominant",
                               "Urban - Medium Density", 
                               "Urban - High Density")) +
  geom_sf(data = mrc.mont.reproj, fill=NA, color=alpha("black",1)) +
  theme_bw() +
  theme(legend.justification=c(0,1), legend.position=c(0.02,0.98))

ggsave(plot = map_vals, "../thesis/figures/profiles_land_use.png", 
       height = 8, width = 12, dpi = the_dpi)

#map_vals
```

![](../thesis/figures/profiles_land_use.png)

### Mean current flow change

#### Historic

Change in mean flow (in \% of the the 1990 flow) between 1990 and 2010, contrasting observed historic change (1990-2010) and simulated change from historic model run.

```{r fig_1_historic, message=FALSE, out.width="100%", fig.cap="Connectivity change for species through time, 1990-2010"}

the_width = 18 
the_height = 15
the_dpi = 500

fig_1_historic <- joined %>% 
  filter(climate == "none") %>% 
  group_by(scenarioId, species, iteration) %>% 
  group_modify(~mutate(.x , baseline = subset(.x, timestep == 1990)$sum_cur, 
                       the_diff = ((sum_cur-baseline)/baseline)*100 )) %>% 
  ggplot(aes(x=timestep, y=the_diff, col=source)) +
  scale_color_manual(values=c('#d8b365','#5ab4ac'), 
                     labels = c("Model", "Observation")) +
  geom_smooth(aes(group = sce), method = "glm", se=FALSE) +
  scale_linetype_manual(values = c(1,3,2,5))+
  geom_point(aes(group = sce)) +
  facet_wrap(~species, scales = "fixed") +
  labs(y = "Current flow (% Change)",
       x = "Year",
       col = "Source") +
  NULL
#fig_1_historic
ggsave(fig_1_historic,
       filename = "../thesis/figures/connectivity_decrease_x5species_historic.png",
       width = the_width, height = the_height, dpi = the_dpi)
```

![](../thesis/figures/connectivity_decrease_x5species_historic.png)

#### Chap 1

##### Line graph

Change in mean flow (in \% of the the 2010 flow) between 2010 (observed) and 2100 (predicted), contrasting BAU scenario (solid line) with other land use change scenarios - linear graph.

```{r figure_1_model_1, message=FALSE,  warning=FALSE, out.width="100%", fig.cap="Connectivity change for species through time, 2010-2100"}

the_width = 18 
the_height = 15
the_dpi = 500

colors_sce <- 
  data.frame(sce = c("Baseline", "RCP85"), 
             color = c("#8DA0CB", "#800026"), stringsAsFactors = F)

fig_1_static_chap_1 <- joined %>% 
  filter(climate != "none", chapter %in% c("none", "both", "chap_1")) %>% 
  group_by(scenarioId, species, iteration) %>% 
  group_modify(~mutate(.x , baseline = subset(.x, timestep == 2010)$sum_cur, 
                       the_diff = ((sum_cur-baseline)/baseline)*100 )) %>% 
  ggplot(aes(x=timestep, y=the_diff, col=climate)) +
  scale_color_manual(values = colors_sce$color, 
                     labels = colors_sce$sce) +
  geom_smooth(aes(group = sce, linetype=code_run), alpha=0.5, method="loess") +
  #geom_point(aes(group = sce)) +
  scale_linetype_manual(values=c("solid", "twodash")) +
  facet_wrap(~species, scales = "fixed") +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  #geom_vline(xintercept = 2020, linetype = "dashed", color = "darkGreen", alpha = 0.5) +
  labs(y = "Current flow (% Change)",
       x = "Year",
       col = "Climate",
       linetype = "Run") +
  guides(col = guide_legend(ncol = 2, nrow=2), 
         linetype = guide_legend(nrow=1, override.aes = list(colour = 'black'))) +
  theme(legend.position = c(0.85, 0.15))+
  NULL
#fig_1_static_chap_1
ggsave(fig_1_static_chap_1, 
       filename = "../thesis/figures/connectivity_decrease_x5species_chap1.png", 
       width = the_width, height = the_height, dpi = the_dpi)
```

![](../thesis/figures/connectivity_decrease_x5species_chap1.png)

##### Radar Graph

Change in mean flow (in \% of the the 2010 flow) between 2010 (observed) and 2100 (predicted), contrasting BAU scenario with other land use change scenarios - radar graph.

```{r figure_1_model_1_radar, message=FALSE, warning=FALSE, out.width="100%"}

# Default size
gridline.label.offset	= 4
legend.text.size = 20
group.line.width = 0.9
grid.label.size = 10
background.circle.colour	= "#ffffff"
group.point.size	= 5
axis.label.size = 8
group.colours = RColorBrewer::brewer.pal(name = "Paired",n=10)[c(2,4,6,8,10)]

radar_data_chap1 <- joined %>% 
  filter(climate != "none", chapter %in% c("none", "both", "chap_1")) %>% 
  group_by(timestep, species, code) %>% 
  summarise(sum_cur = mean(sum_cur)) %>% ungroup %>% 
  pivot_wider(names_from = timestep, values_from = sum_cur) %>% 
  mutate(diff = ((`2100`-`2010`)/`2010`)*100) %>% 
  select(-c(`2010`:`2100`)) %>%  
  pivot_wider(names_from = code, values_from = diff) %>% 
  rename(group = species)

radar_chap1 <-  
  ggradar(radar_data_chap1, centre.y = -20, legend.position = "right",
          grid.min = -20, grid.max = 3, grid.mid = 0, 
          values.radar = c("-20 %", "0 %", ""), 
          group.colours = group.colours,
          gridline.label.offset	= gridline.label.offset, 
          legend.text.size = legend.text.size,
          group.line.width = group.line.width, 
          grid.label.size = grid.label.size,
          background.circle.colour = background.circle.colour, 
          group.point.size	= group.point.size, 
          axis.label.size = axis.label.size)
#radar_chap1
ggsave("../thesis/figures/radar_ggradar_chap1.png", radar_chap1,
       width = the_width, height = the_height, dpi = the_dpi)
```

![](../thesis/figures/radar_ggradar_chap1.png)

#### Chap 2

Change in mean flow (in \% of the the 2010 flow) between 2010 (observed) and 2100 (predicted), contrasting BAU scenario (solid line) with other conservation scenarios - linear graph.

```{r figure_1_model_2, message=FALSE, warning=FALSE, out.width="100%", fig.cap="Connectivity change for species through time, 2010-2100"}
fig_1_static_chap_2 <- joined %>% 
  filter(climate != "none", chapter %in% c("none", "both", "chap_2")) %>% 
  group_by(scenarioId, species, iteration) %>% 
  group_modify(~mutate(.x , baseline = subset(.x, timestep == 2010)$sum_cur, 
                       the_diff = ((sum_cur-baseline)/baseline)*100 )) %>% 
  ggplot(aes(x=timestep, y=the_diff, col=climate)) +
  scale_color_manual(values = colors_sce$color,
                     labels = colors_sce$sce) +
  geom_smooth(aes(group = sce, linetype=code_run), alpha=0.5, method="loess") +
  scale_linetype_manual(values=c("solid", "dashed", "longdash", "dotted")) +
  facet_wrap(~species, scales = "fixed") +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  geom_vline(xintercept = 2020, linetype = "dashed", color = "darkGreen", alpha = 0.5) +
  labs(y = "Current flow (% Change)",
       x = "Year",
       col = "Climate",
       linetype = "Run") +
  guides(col = guide_legend(ncol = 2, nrow=2), 
         linetype = guide_legend(nrow = 2, ncol=2, 
                                 override.aes = list(colour = 'black'))) +
  theme(legend.position = c(0.85, 0.15))+
  NULL
#fig_1_static_chap_2
ggsave(fig_1_static_chap_2, 
       filename = "../thesis/figures/connectivity_decrease_x5species_chap2.png", 
       width = the_width, height = the_height, dpi = the_dpi)
fig_1_static_chap_2
```

![](../thesis/figures/connectivity_decrease_x5species_chap2.png)

##### Radar Graph

Change in mean flow (in \% of the the 2010 flow) between 2010 (observed) and 2100 (predicted), contrasting BAU scenario (solid line) with other conservation scenarios - radar graph.

```{r figure_1_model_2_radar, message=FALSE, warning=FALSE, out.width="100%"}
radar_data_chap2 <- joined %>% 
  filter(climate != "none", chapter %in% c("none", "both","chap_2")) %>% 
  group_by(timestep, species, code) %>% 
  summarise(sum_cur = mean(sum_cur)) %>% ungroup %>% 
  pivot_wider(names_from = timestep, values_from = sum_cur) %>% 
  mutate(diff = ((`2100`-`2010`)/`2010`)*100) %>% 
  select(-c(`2010`:`2100`)) %>%  
  pivot_wider(names_from = code, values_from = diff) %>% 
  rename(group = species)

radar_chap2 <-  
  ggradar(radar_data_chap2, centre.y = -20, legend.position = "right",
          grid.min = -20, grid.max = 3, grid.mid = 0, 
          values.radar = c("-20 %", "0 %", ""), 
          group.colours = group.colours,
          gridline.label.offset	= gridline.label.offset, 
          legend.text.size = legend.text.size,
          group.line.width = group.line.width, 
          grid.label.size = grid.label.size,
          background.circle.colour = background.circle.colour, 
          group.point.size	= group.point.size, 
          axis.label.size = axis.label.size-2)
#radar_chap2
ggsave("../thesis/figures/radar_ggradar_chap2.png", radar_chap2,
       width = the_width, height = the_height, dpi = the_dpi)
```

![](../thesis/figures/radar_ggradar_chap2.png)

#### Both

##### Radar Graph

Change in mean flow (in \% of the the 2010 flow) between 2010 (observed) and 2100 (predicted), contrasting BAU scenario (solid line) with all other scenarios from borth chapters - radar graph.

```{r figure_1_model_both_radar, message=FALSE, warning=FALSE, out.width="100%"}
radar_data_all <- joined %>% 
  filter(climate != "none") %>% 
  group_by(timestep, species, code) %>% 
  summarise(sum_cur = mean(sum_cur)) %>% ungroup %>% 
  pivot_wider(names_from = timestep, values_from = sum_cur) %>% 
  mutate(diff = ((`2100`-`2010`)/`2010`)*100) %>% 
  select(-c(`2010`:`2100`)) %>%  
  pivot_wider(names_from = code, values_from = diff) %>% 
  rename(group = species)

radar_all <-  
  ggradar(radar_data_all, centre.y = -20, legend.position = "right",
          grid.min = -20, grid.max = 3, grid.mid = 0, 
          values.radar = c("-20 %", "0 %", ""), 
          group.colours = group.colours,
          gridline.label.offset	= gridline.label.offset, 
          legend.text.size = legend.text.size,
          group.line.width = group.line.width, 
          grid.label.size = grid.label.size,
          background.circle.colour = background.circle.colour, 
          group.point.size	= group.point.size, 
          axis.label.size = axis.label.size-2)
#radar_all
ggsave("../thesis/figures/radar_ggradar_both.png", radar_all,
       width = the_width, height = the_height, dpi = the_dpi)
```

![](../thesis/figures/radar_ggradar_both.png)

### ROC Curves

ROC curves from fitting both models (with resampling), plotted with the `patchwork` package.

```{r roc_curves, eval=!(REPRO_FIGS_ONLY), message=FALSE, warning=FALSE, out.width="100%"}
# Roc curves data
urb <- readRDS("../data/temp/fit_rs_outcome_rf_urb_2.RDS")
agex <- readRDS("../data/temp/fit_rs_outcome_rf_agex_2.RDS")

p1 <- bind_rows(urb, .id = "fold") %>% 
  mutate(Fold = factor(fold, levels = as.character(1:10))) %>% 
  group_by(Fold) %>% 
  roc_curve(.pred, truth = outcome_fact) %>% 
  autoplot(add=T) +
  ggtitle(paste("Urbanisation")) +
  annotate(x = 0.75, y = 0.25, geom="label", size = 3,
           label = as.character(paste("Av AUC = 0.938 +/- 0.002"))) +
  theme(legend.position = "none") +
  NULL

p2 <- bind_rows(agex, .id = "fold") %>% 
  mutate(Fold = factor(fold, levels = as.character(1:10))) %>% 
  group_by(Fold) %>% 
  roc_curve(.pred, truth = outcome_fact) %>% 
  autoplot(add=T) +
  ggtitle(paste("Agricultural Expansion")) +
  annotate(x = 0.75, y = 0.25, geom="label", size = 3,
           label = as.character(paste("Av AUC = 0.929 +/- 0.002"))) +
  NULL

full = p1 + p2

#full
ggsave("../thesis/figures/double_roc_resample.png", full)
```

![](../thesis/figures/double_roc_resample.png)

### Histograms

#### Historic

Histograms of flow values for the year 2010 comparing the historic model run (1990-2010) with observations.

```{r original_hists, message=FALSE, warning=FALSE, out.width="100%"}

bi_colors <- c("#67a9cf", "#ef8a62")

hist_plot_origin <- histograms %>%
  filter(sce %in% c("TRUE", "sce_37")) %>% 
  filter(ts == 2) %>% 
  mutate(ts = ts*10+1990) %>% 
  mutate(ts = as.factor(ts)) %>% 
  ggplot(aes(x=bins, y=n, fill=sce, colour =sce)) + 
  geom_density(stat='identity', show.legend=T, alpha=0.3)+
  scale_fill_manual(values = bi_colors, name="Source", labels=c("Model", "Observation")) +
  scale_color_manual(values = bi_colors, name="Source", labels=c("Model", "Observation")) +
  facet_wrap(~species) +
  labs(x = "Flow intensity distribution (log)", 
       y = "") +
  theme(strip.text.y = element_text(angle=360, size=10, hjust = 0), 
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        axis.text.y.left = element_blank(),
        axis.ticks.y.left = element_blank()) +
  NULL
#hist_plot_origin
ggsave("../thesis/figures/original_hists.png", hist_plot_origin,
       width = the_width, height = the_height, dpi = the_dpi)

hist_plot_origin
```

![](../thesis/figures/original_hists.png)

#### Chap 1

Histograms of flow values for all species and scenarios, for 2010 (observed) and 2100 (predicted).

```{r hist_plot_1, message=FALSE, warning=FALSE, out.width="100%"}

bi_colors <- c("#67a9cf", "#ef8a62")

hist_plot_1 <- histograms %>% 
  mutate(ts = as.factor(ts)) %>% 
  filter(chapter %in% c('both', 'chap_1')) %>%  
  filter(climate != "baseline") %>% 
  ggplot(aes(x=bins, y=n, group=ts)) + 
  geom_density(stat='identity', show.legend=T, aes(fill=factor(ts), color=factor(ts)), alpha=0.3)+
  facet_grid(code~species) +
  scale_fill_manual(values = bi_colors, 
                    labels = c("2010", "2100")) +
  scale_color_manual(values = bi_colors, 
                     labels = c("2010", "2100")) +
  labs(x = "Flow intensity distribution (log)", 
       y = "")+
  theme(strip.text.y = element_text(angle=360, size=20, hjust = 0), 
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        axis.text.y.left = element_blank(),
        axis.ticks.y.left = element_blank(), 
        legend.position = c(0.94, 0.94)) +
  labs(color="Timestep", fill="Timestep") +
  NULL
#hist_plot_1
ggsave("../thesis/figures/hist_chap1.png", hist_plot_1, width = 18, height = 15)
hist_plot_1
```

![](../thesis/figures/hist_chap1.png)

#### Chap 2

Histograms of flow values for all species and scenarios, for 2010 (observed) and 2100 (predicted).

```{r hist_plot_2, message=FALSE, warning=FALSE, out.width="100%"}
hist_plot_2 <- histograms %>% 
  mutate(ts = as.factor(ts)) %>% 
  filter(chapter %in% c('chap_2')) %>%  
  filter(climate != "baseline") %>% 
  ggplot(aes(x=bins, y=n, group=ts)) + 
  geom_density(stat='identity', aes(fill=factor(ts),  colour=factor(ts)), alpha=0.3)+
  facet_grid(species~code) +
  scale_fill_manual(values = bi_colors, 
                    labels = c("2010", "2100")) +
  scale_color_manual(values = bi_colors, 
                     labels = c("2010", "2100")) +
  labs(x = "Flow intensity distribution (log)", 
       y = "") +
  theme(strip.text.y = element_text(angle=360, size=10, hjust = 0), 
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        axis.text.y.left = element_blank(),
        axis.ticks.y.left = element_blank(),
        legend.position = c(0.94, 0.94)) +
  labs(color="Timestep", fill="Timestep") +
  NULL
#hist_plot_2
ggsave("../thesis/figures/hist_chap2.png", hist_plot_2, width = 18, height = 20)

hist_plot_2
```

![](../thesis/figures/hist_chap2.png)

#### BOTH
```{r hist_plot_3, message=FALSE, warning=FALSE, out.width="100%"}
wes_2_colors = wes_palette("Zissou1", 2, type = "continuous")
hist_plot_3 <- histograms %>% 
  mutate(ts = as.factor(ts)) %>%

  filter(chapter %in% c('both', 'chap_1', 'chap_2')) %>%  
  filter(climate != "baseline") %>%

  ggplot(aes(x=bins, y=n, group=ts)) + 
  geom_density(stat='identity', aes(colour=factor(ts)), alpha=0.3)+
  facet_grid(code~species) +
  # scale_fill_manual(values = wes_2_colors, 
  #                   labels = c("2010", "2100")) +
  scale_color_manual(values = wes_2_colors, 
                     labels = c("2010", "2100")) +
  labs(x = "Flow intensity distribution (log)", 
       y = "") +
  theme(strip.text.y = element_text(angle=360, size=25, hjust = 0), 
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        axis.text.y.left = element_blank(),
        axis.ticks.y.left = element_blank(),
        legend.position = c(0.94, 0.94)) +
  labs(color="Timestep") +
  NULL
#hist_plot_2
ggsave("../thesis/figures/hist_both_chap.png", hist_plot_3, width = 18, height = 15)

hist_plot_3
```

### SURF Analysis

#### Chap1

##### Linear Graph

Change in the number of features (in \% of the the 2010 flow) identified by the SURF analyssis between 2010 (observed) and 2100 (predicted), contrasting BAU scenario with other land use change scenarios - linear graph.

```{r surf_1, message=FALSE, warning=FALSE, out.width="100%"}
surf_1 <- surf %>% 
  mutate(scenario = as.factor(scenario)) %>% 
  filter(climate != "none", chapter %in% c("none", "both", "chap_1")) %>% 
  filter(climate != "none") %>% 
  group_by(scenario, species, iter) %>% 
  group_modify(~mutate(.x , baseline = subset(.x, timestep == 2010)$kp_nb, 
                       the_diff = ((kp_nb-baseline)/baseline)*100 )) %>% ungroup %>% 
  ggplot(aes(x = timestep, y = the_diff, color=climate)) +
  scale_linetype_manual(values=c("solid", "twodash")) +
  geom_smooth(aes(group = scenario, linetype=code_run), alpha=0.2, method="loess")+
  #geom_point(aes(group=scenario), alpha= 0.05, size = 0.5)+
  facet_wrap(~species, scales = "fixed")+
  # scale_color_manual(values = colors_sce$color,
  #                    labels = colors_sce$sce) +
  labs(y = "Change in nb of Keypoints detected (%)",
       x = "Year",
       col = "Climate",
       linetype = "Run") +
  guides(col = guide_legend(ncol = 2, nrow=2), 
         linetype = guide_legend(nrow=1, override.aes = list(colour = 'black'))) +
  NULL
#surf_1
ggsave("../thesis/figures/surf_chap1.png", surf_1,
       width = the_width, height = the_height, dpi = the_dpi)
```

![](../thesis/figures/surf_chap1.png)

##### Radar Graph

Change in the number of features (in \% of the the 2010 flow) identified by the SURF analyssis between 2010 (observed) and 2100 (predicted), contrasting BAU scenario with other land use change scenarios - radar graph.

```{r surf_radar_1, message=FALSE, warning=FALSE, out.width="100%"}
surf_radar_data_1 <- surf %>% 
  filter(climate != "none", chapter %in% c("none", "both","chap_1")) %>% 
  filter(sce != "sce_0", name != "historic run") %>% 
  group_by(timestep, species, code) %>% 
  summarise(kp_nb = mean(kp_nb)) %>% ungroup %>% 
  pivot_wider(names_from = timestep, values_from = kp_nb) %>% 
  mutate(diff = ((`2100`-`2010`)/`2010`)*100) %>% 
  select(-c(`2010`:`2100`)) %>%  
  pivot_wider(names_from = code, values_from = diff) %>% 
  rename(group = species)

surf_radar_1 <- 
  ggradar(surf_radar_data_1, centre.y = -60, legend.position = "right",
          grid.min = -60, grid.max = 5, grid.mid = 0, 
          values.radar = c("-60 %", "0 %", ""), 
          group.colours = group.colours,
          gridline.label.offset	= 20, 
          legend.text.size = legend.text.size,
          group.line.width = group.line.width, 
          grid.label.size = grid.label.size,
          background.circle.colour = background.circle.colour, 
          group.point.size	= group.point.size, 
          axis.label.size = axis.label.size)
# surf_radar_1
ggsave("../thesis/figures/surf_radar_chap1.png", surf_radar_1,
       width = the_width, height = the_height, dpi = the_dpi)
```

![](../thesis/figures/surf_radar_chap1.png)

#### Chap 2

##### Linear Graph

Change in the number of features (in \% of the the 2010 flow) identified by the SURF analyssis between 2010 (observed) and 2100 (predicted), contrasting BAU scenario with other conservation scenarios - linear graph.

```{r surf_2, message=FALSE, warning=FALSE, out.width="100%"}
surf_2 <- surf %>% 
  mutate(scenario = as.factor(scenario)) %>% 
  filter(climate != "none", chapter %in% c("none", "both", "chap_2")) %>% 
  filter(climate != "none") %>% 
  group_by(scenario, species, iter) %>% 
  group_modify(~mutate(.x , baseline = subset(.x, timestep == 2010)$kp_nb, 
                       the_diff = ((kp_nb-baseline)/baseline)*100 )) %>% ungroup %>% 
  ggplot(aes(x = timestep, y = the_diff, color=climate)) +
  geom_smooth(aes(group = scenario, linetype=code_run), alpha=0.2, method="loess")+
  #geom_point(aes(group=scenario), alpha= 0.05, size = 0.5)+
  scale_linetype_manual(values=c("solid", "dashed", "longdash", "dotted")) +
  facet_wrap(~species, scales = "fixed")+
  # scale_color_manual(values = colors_sce$color,
  #                    labels = colors_sce$sce) +
  labs(y = "Change in nb of Keypoints detected (%)",
       x = "Year",
       col = "Climate",
       linetype = "Run") +
  guides(col = guide_legend(ncol = 2, nrow=2), 
         linetype = guide_legend(nrow=2, ncol=2, 
                                 override.aes = list(colour = 'black'))) +
  NULL
#surf_2
ggsave("../thesis/figures/surf_chap2.png", surf_2,
       width = the_width, height = the_height, dpi = the_dpi)
```

![](../thesis/figures/surf_chap2.png)

##### Radar Graph

Change in the number of features (in \% of the the 2010 flow) identified by the SURF analyssis between 2010 (observed) and 2100 (predicted), contrasting BAU scenario with other conservation scenarios - radar graph.

```{r surf_radar_2, message=FALSE, warning=FALSE, out.width="100%"}
surf_radar_data_2 <- surf %>% 
  filter(climate != "none", chapter %in% c("none", "both","chap_2")) %>% 
  filter(sce != "sce_0", name != "historic run") %>% 
  group_by(timestep, species, code) %>% 
  summarise(kp_nb = mean(kp_nb)) %>% ungroup %>% 
  pivot_wider(names_from = timestep, values_from = kp_nb) %>% 
  mutate(diff = ((`2100`-`2010`)/`2010`)*100) %>% 
  select(-c(`2010`:`2100`)) %>%  
  pivot_wider(names_from = code, values_from = diff) %>% 
  rename(group = species)

surf_radar_2 <- 
  ggradar(surf_radar_data_2, centre.y = -60, legend.position = "right",
          grid.min = -60, grid.max = 5, grid.mid = 0, 
          values.radar = c("-60 %", "0 %", ""), 
          group.colours = group.colours,
          gridline.label.offset	= 22, 
          legend.text.size = legend.text.size,
          group.line.width = group.line.width, 
          grid.label.size = grid.label.size,
          background.circle.colour = background.circle.colour, 
          group.point.size	= group.point.size, 
          axis.label.size = axis.label.size-2)
#surf_radar_2
ggsave("../thesis/figures/surf_radar_chap2.png", surf_radar_2, 
       width = the_width, height = the_height, dpi = the_dpi)
```

![](../thesis/figures/surf_radar_chap2.png)

### Prediction maps

```{r 2010_compare_prep, eval=!(REPRO_FIGS_ONLY), message=FALSE, warning=FALSE, out.width="100%"}
list_of_files <-  list.files("../libraries/stsim/monteregie-conncons-scripted.ssim.output/Scenario-37/stsim_OutputSpatialState", full.names = T)
list_of_files <- list_of_files[grep(x=list_of_files, "ts2")]
stratum <- raster("../data/stsim/aggregated/primary_stratum_mont_or_not_or_PA.tif")
```

Predicted probability of urbanization (left) vs observed urbanization (right) between 1990 and 2010.

```{r 2010_compare_urb, eval=!(REPRO_FIGS_ONLY), message=FALSE, warning=FALSE, out.width="100%"}
names <- paste0(c("historic_compare_agex", "historic_compare_urb"), ".png")

for (class in c(1:2)){

  the_stack <- stack(lapply(list_of_files, raster)) == class
  original <- raster("../data/land_use/aggregated/aggregated_lu_buffered_1990_patched.tif") == class
  final <- raster("../data/land_use/aggregated/aggregated_lu_buffered_2010_patched.tif") == class

  # Transform, give names
  the_stack[original == 1] <- 0
  final[original == 1] <- 0
  the_stack[stratum==0] <- NA
  final[stratum==0] <- NA
  the_stack <- trim(the_stack)
  final <- trim(final)

  the_mean <-  mean(the_stack)
  names(the_mean) <- "mean"
  names(final) <- "observation"

  # test <- the_mean
  # test[(which(values(!is.na(final)) & values(is.na(the_mean))))] <- 100
  # plot(test)

  # Fill the "other" values
  the_mean[(which(values(!is.na(final)) & values(is.na(the_mean))))] <- 0

  stack_to_plot <- stack(the_mean, final) 
  names(stack_to_plot) <- c("Model", "Observations")

  mxp <- 1e10000 # 1000

  myPal <- RColorBrewer::brewer.pal('OrRd', n=9)
  myTheme <- rasterTheme(region = myPal)
  probs <- levelplot(stack_to_plot$Model, scales=list(draw=FALSE),
                     maxpixels = mxp, 
                     par.settings = myTheme,
                     margin=FALSE, auto.key=FALSE, 
                     colorkey=list(space="bottom"), 
                     main = "A)")
  probs <- as.ggplot(probs)

  stack_to_plot$Observations <- ratify(stack_to_plot$Observations)
  rat <- levels(stack_to_plot$Observations)[[1]]
  rat$Urbanised <- c("Not Observed", "Observed")
  levels(stack_to_plot$Observations) <- rat
  bin <- levelplot(stack_to_plot$Observations, col.regions=c('grey', 'indianred4'), 
                   scales=list(draw=FALSE), 
                   maxpixels = mxp, 
                   colorkey=list(space="bottom", height=0.3), 
                   main = "B)")
  bin <- as.ggplot(bin)

  final_plot <- probs + bin

  ggsave(plot = final_plot, filename = paste0("../thesis/figures/",names[class]), width = 20, height = 10, dpi = the_dpi)
}
```

![](../thesis/figures/historic_compare_urb.png)

Predicted probability of agricultural expansion (left) vs observed agricultural expansion (right) between 1990 and 2010.

![](../thesis/figures/historic_compare_agex.png)

### Land use maps

Land use maps showing predictions of land use changes (Only BAU is displayed below).

```{r compare, eval=!(REPRO_FIGS_ONLY), message=FALSE, warning=FALSE, out.width="100%"}
source('../scripts/functions/draw_scenario.R')
mxp <- 1e10000 # 1000
the_dpi = 300
for (scenario in c(39, 42, 45, 48, 51)){
  the_plot <- draw_scenario(scenario, mxp = mxp, sce_folder = "/media/vlucet/backup/results_july/monteregie-conncons-scripted.ssim.output/")
  the_name <- paste0("scenario_", as.character(scenario),"_compare.png")
  ggsave(plot = the_plot,
  filename = paste0("../thesis/figures/",the_name), width = 15, height = 18, dpi = the_dpi)
}
```

Comparison of Monteregie at the beginning of the BAU scenario in 2010, and at the end in 2100 (under baseline climate scenario).

![](../thesis/figures/scenario_39_compare.png)
<!--
![](../thesis/figures/scenario_42_compare.png)
![](../thesis/figures/scenario_45_compare.png)
![](../thesis/figures/scenario_48_compare.png)
![](../thesis/figures/scenario_51_compare.png)
-->

### Bar charts

<!--
#### Scenario BAU

```{r BAU, message=FALSE, warning=FALSE, out.width="100%"}

final_df_39 <- bar_data %>% 
filter(scenario == 39)

bar_plot_39 <- final_df_39 %>% 
#filter(value > 10) %>% 
mutate(value = as.factor(value), timestep = as.factor(timestep)) %>% 
ggplot(aes(fill=timestep, y=area, x=new_class)) + 
geom_bar(position="dodge", stat="identity") +
#geom_errorbar(aes(ymin=mean_count-sd_count, ymax=mean_count+sd_count), width=.2,
#               position=position_dodge(.9)) +
scale_fill_manual(values = bi_colors) +
theme(legend.position = c(0.85, 0.7)) +
labs(x = "Land use Class", 
y = "Mean Area (hectare)", 
fill = "Year") +
NULL
#bar_plot_39
#ggsave("../thesis/figures/bar_BAU.png", bar_plot_39, width = 15, height = 10)
```

##### Scenario Reforestation

```{r reforestation,  message=FALSE, warning=FALSE, out.width="100%"}
# draw_scenario(42)
final_df_42 <- bar_data %>% 
filter(scenario == 42)

bar_plot_42 <- final_df_42  %>% 
#filter(value > 10) %>% 
mutate(value = as.factor(value), timestep = as.factor(timestep)) %>% 
ggplot(aes(fill=timestep, y=area, x=new_class)) + 
geom_bar(position="dodge", stat="identity") +
scale_fill_manual(values = bi_colors) +
#geom_errorbar(aes(ymin=mean_count-sd_count, ymax=mean_count+sd_count), width=.2,
#               position=position_dodge(.9)) +
theme(legend.position = c(0.85, 0.7)) +
labs(x = "Land use Class", 
y = "Mean Area (hectare)", 
fill = "Year") +
NULL
#bar_plot_42
#ggsave("../thesis/figures/bar_Ref.png", bar_plot_42, width = 15, height = 10)
```
-->

##### Both scenarios

Land use class frequencies in 2010 (observed) and 2100 (predicted) for the BAU and Reforestation scenarios (under baseline climate scenario).

```{r bar_boths}
final_df_39 <- final_df_39 %>% 
  mutate(sce = "BAU") %>% 
  mutate(type = ifelse(value <10, "non_forest", "forest"))
final_df_42 <- final_df_42 %>% 
  mutate(sce = "Reforestation")  %>% 
  mutate(type = ifelse(value <10, "non_forest", "forest"))

final_all <- bind_rows(final_df_39, final_df_42)

forest <- final_all %>% 
  filter(type == "forest") %>% 
  mutate(value = as.factor(value), timestep = as.factor(timestep)) %>% 
  ggplot(aes(fill=timestep, y=area, x=new_class)) + 
  geom_bar(position="dodge", stat="identity") +
  scale_fill_manual(values = bi_colors) +
  theme(legend.position = "none") +
  facet_grid(~sce) +
  labs(x = "Land use Class", 
       y = "Mean Area (hectare)", 
       fill = "Year") +
  NULL

non_forest <- final_all %>% 
  filter(type != "forest") %>% 
  filter(!(new_class %in% c("Water", "Wetlands", "Roads"))) %>% 
  mutate(value = as.factor(value), timestep = as.factor(timestep)) %>% 
  ggplot(aes(fill=timestep, y=area, x=new_class)) + 
  geom_bar(position="dodge", stat="identity") +
  scale_fill_manual(values = bi_colors) +
  theme(legend.position = "right") +
  facet_grid(~sce) +
  labs(x = "Land use Class", 
       y = "Mean Area (hectare)", 
       fill = "Year") +
  NULL

assembled <- wrap_plots(forest, non_forest, ncol=1, nrow=2)
#assembled
ggsave("../thesis/figures/bar_both.png", assembled,
       width = the_width, height = 20, dpi = the_dpi)
```

![](../thesis/figures/bar_both.png)

Land use class frequencies in 2010 (observed) and 2100 (predicted) for the BAU and conservation scenarios (under baseline climate scenario).

```{r eval=FALSE, include=FALSE, message=FALSE, warning=FALSE}
# draw_scenario(42) ...

all_chap2 <- bar_data %>% 
  filter(scenario %in% c(42, 45, 48, 51))

#forest <-

bar_chap2 <- all_chap2 %>% 
  mutate(value = ifelse(.$value>10, 10, .$value)) %>%
  mutate(new_class = ifelse(.$value == 10, "forest", .$new_class)) %>% 
  group_by(new_class, timestep, scenario, value) %>% 
  summarise(area = sum(area)) %>% ungroup() %>% 
  filter(value %in% c(1,2,10)) %>% 
  left_join(results_clean, by = c("scenario"="scenarioId")) %>% 
  mutate(value = as.factor(value), timestep = as.factor(timestep)) %>% 
  ggplot(aes(fill=timestep, y=area, x=new_class)) + 
  geom_bar(position="dodge", stat="identity") +
  scale_fill_manual(values = bi_colors) +
  facet_wrap(~name, labeller = label_wrap_gen(width = 30)) +
  theme(legend.position = c(0.9, 0.8)) +
  labs(x = "Land use Class", 
       y = "Mean Area (hectare)", 
       fill = "Year") +
  NULL
#bar_chap2
ggsave("../thesis/figures/bar_chap2.png", bar_chap2, 
       width = the_width, height = 20, dpi = the_dpi)
```

![](../thesis/figures/bar_chap2.png)

```{r eval=FALSE, include=FALSE}
##### ##### ##### ##### ##### ##### ##### 
#####  Other code for calculation   ##### 
##### ##### ##### ##### ##### ##### #####

# Compute auc / precision on predictions.

options(yardstick.event_first = T)
# crosstab(the_mean>0, final)

summary_vals <- data.frame()

for (class in c(1:2)){

  the_stack <- stack(lapply(list_of_files, raster)) == class
  original <- raster("../data/land_use/aggregated/aggregated_lu_buffered_1990_patched.tif") == class
  final <- raster("../data/land_use/aggregated/aggregated_lu_buffered_2010_patched.tif") == class

  # Transform, give names
  the_stack[original == 1] <- 0
  final[original == 1] <- 0
  the_stack[stratum==0] <- NA
  final[stratum==0] <- NA
  the_stack <- trim(the_stack)
  final <- trim(final)

  the_mean <-  mean(the_stack)
  names(the_mean) <- "mean"
  names(final) <- "observation"

  names(the_stack) <- rep("model", length(names(the_stack)))

  aucs <- c()
  precs <- c()
  for (x in 1:nlayers(the_stack)){
    the_df <- as.data.frame(stack(the_stack[[x]], final)) %>% 
      drop_na() %>% 
      rename(estimate = starts_with("model"), truth = "observation")
    the_df_2 <- the_df %>% 
      mutate(truth = factor(truth, ordered = F)) %>%
      mutate(truth = relevel(.$truth, "1"))
    auc <- the_df_2 %>% 
      roc_auc(truth = truth, estimate) %>% 
      pull(.estimate)
    prec <- the_df_2 %>% 
      average_precision(truth = truth, estimate) %>% 
      pull(.estimate)
    aucs <- c(aucs, auc)
    precs <- c(precs, prec)
  }

  summary_vals <- bind_rows(summary_vals, 
                            data.frame(class = class, 
                                       mean_auc = mean(aucs), sd_auc = sd(aucs), 
                                       mean_prec = mean(precs), sd_prec = sd(precs)))

}

write_csv(summary_vals, "../outputs/metrics/summary_vals.csv")
summary_vals
```

<!-- ### Map of change

Mapping takes a lot of time, not knitted by default

#### Chap

```{r figure_2_chap1, eval=FALSE, message=FALSE, warning=FALSE}
# change_1 <- df_final_fordiff_pivoted %>% 
#   filter(chapter %in% c("none", "both", "chap_1")) %>%
#   filter(sce %in% c('sce_39', 'sce_39')) %>% 
#   ggplot(aes(fill=change)) +
#   geom_sf(show.legend=T, lwd = 0)  + 
#   facet_wrap(~sce) + 
#   ggtitle("Connectivity change in %", 
#           subtitle = "2010-2100") +
#   scale_fill_binned(low='#d13e11', high='#fff7bc', 
#                     breaks=c(-50, -40, -30, -20, -10, 0, 10))
# change_1
#ggsave("../thesis/figures/connectivit_change_mun_chap1.png", change_1)
```

#### Chap 2

```{r figure_2_chap2, eval=FALSE, message=FALSE, warning=FALSE}
# change_2 <- df_final_fordiff_pivoted %>% 
#   filter(chapter %in% c("none", "both", "chap_2")) %>%
#   ggplot(aes(fill=change)) +
#   geom_sf(show.legend=T, lwd = 0)  + 
#   facet_wrap(~sce) + 
#   ggtitle("Connectivity change in %", 
#           subtitle = "2010-2100") +
#   scale_fill_binned(low='#d13e11', high='#fff7bc', 
#                     breaks=c(-50, -40, -30, -20, -10, 0, 10))
# change_2
#ggsave("../thesis/figures/connectivit_change_mun_chap2.png", change_2)
```

#### Both

```{r figure_2_all, eval=FALSE, message=FALSE, warning=FALSE}
# change_both <- df_final_fordiff_pivoted %>% 
#   ggplot(aes(fill=change)) +
#   geom_sf(show.legend=T, lwd = 0)  + 
#   facet_wrap(~sce) + 
#   ggtitle("Connectivity change in %", 
#           subtitle = "2010-2100") +
#   scale_fill_binned(low='#d13e11', high='#fff7bc', 
#                     breaks=c(-50, -40, -30, -20, -10, 0, 10))
# change_both
#ggsave("../thesis/figures/connectivit_change_mun_both.png", change_both)
```

-->

### Networks
```{r, message=FALSE, warning=FALSE, out.width="100%", fig.cap=""}
networks <- read_csv("../outputs/final/final_df_network_stats.csv") %>% 
  left_join(results_clean, by = c("scenario" = "scenarioId"))
```

##### CHAP 1
```{r, message=FALSE, warning=FALSE, out.width="100%", fig.cap=""}
networks <- read_csv("../outputs/final/final_df_network_stats.csv") %>% 
  left_join(results_clean, by = c("scenario" = "scenarioId"))

networks %>% 
  mutate(timestep = timestep*10 + 1990) %>% 
  filter(climate != "none", chapter %in% c("none", "both", "chap_1")) %>% 
  group_by(scenario, species, iteration) %>% 
  group_modify(~mutate(.x , baseline = subset(.x, timestep == 2010)$overallECgap, 
                       the_diff = ((overallECgap-baseline)/baseline)*100 )) %>%
  ggplot(aes(x=timestep, y=the_diff, col=climate)) +
  scale_color_manual(values = colors_sce$color, 
                     labels = colors_sce$sce) +
  geom_smooth(aes(group = sce, linetype=code_run), alpha=0.5, method="loess") +
  #geom_point(aes(group = sce)) +
  scale_linetype_manual(values=c("solid", "twodash")) +
  facet_wrap(~species, scales = "fixed") +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  #geom_vline(xintercept = 2020, linetype = "dashed", color = "darkGreen", alpha = 0.5) +
  labs(y = "overallECgap",
       x = "Year",
       col = "Climate",
       linetype = "Run") +
  guides(col = guide_legend(ncol = 2, nrow=2), 
         linetype = guide_legend(nrow=1, override.aes = list(colour = 'black'))) +
  theme(legend.position = c(0.85, 0.15))+
  NULL -> net_1

ggsave(net_1, 
       filename = "../thesis/figures/ECGAP.png", 
       width = the_width, height = the_height, dpi = the_dpi)

```

```{r, message=FALSE, warning=FALSE, out.width="100%", fig.cap=""}
networks %>% 
  mutate(timestep = timestep*10 + 1990) %>% 
  filter(climate != "none", chapter %in% c("none", "both", "chap_1")) %>% 
  group_by(scenario, species, iteration) %>% 
  group_modify(~mutate(.x , baseline = subset(.x, timestep == 2010)$overallECnatal, 
                       the_diff = ((overallECnatal-baseline)/baseline)*100 )) %>%
  ggplot(aes(x=timestep, y=the_diff, col=climate)) +
  scale_color_manual(values = colors_sce$color, 
                     labels = colors_sce$sce) +
  geom_smooth(aes(group = sce, linetype=code_run), alpha=0.5, method="loess") +
  #geom_point(aes(group = sce)) +
  scale_linetype_manual(values=c("solid", "twodash")) +
  facet_wrap(~species, scales = "fixed") +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  #geom_vline(xintercept = 2020, linetype = "dashed", color = "darkGreen", alpha = 0.5) +
  labs(y = "overallECnatal",
       x = "Year",
       col = "Climate",
       linetype = "Run") +
  guides(col = guide_legend(ncol = 2, nrow=2), 
         linetype = guide_legend(nrow=1, override.aes = list(colour = 'black'))) +
  theme(legend.position = c(0.85, 0.15))+
  NULL -> net_2

ggsave(net_2, 
       filename = "../thesis/figures/ECNATAL.png", 
       width = the_width, height = the_height, dpi = the_dpi)

```

```{r, message=FALSE, warning=FALSE, out.width="100%", fig.cap=""}
networks %>% 
  mutate(timestep = timestep*10 + 1990) %>% 
  filter(climate != "none", chapter %in% c("none", "both", "chap_1")) %>% 
  group_by(scenario, species, iteration) %>% 
  group_modify(~mutate(.x , baseline = subset(.x, timestep == 2010)$numNodes, 
                       the_diff = ((numNodes-baseline)/baseline)*100 )) %>%
  ggplot(aes(x=timestep, y=the_diff, col=climate)) +
  scale_color_manual(values = colors_sce$color, 
                     labels = colors_sce$sce) +
  geom_smooth(aes(group = sce, linetype=code_run), alpha=0.5, method="loess") +
  #geom_point(aes(group = sce)) +
  scale_linetype_manual(values=c("solid", "twodash")) +
  facet_wrap(~species, scales = "fixed") +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  #geom_vline(xintercept = 2020, linetype = "dashed", color = "darkGreen", alpha = 0.5) +
  labs(y = "numNodes",
       x = "Year",
       col = "Climate",
       linetype = "Run") +
  guides(col = guide_legend(ncol = 2, nrow=2), 
         linetype = guide_legend(nrow=1, override.aes = list(colour = 'black'))) +
  theme(legend.position = c(0.85, 0.15))+
  NULL -> net_3

ggsave(net_3, 
       filename = "../thesis/figures/NUMNODES.png", 
       width = the_width, height = the_height, dpi = the_dpi)

```

```{r, message=FALSE, warning=FALSE, out.width="100%", fig.cap=""}
networks %>% 
  mutate(timestep = timestep*10 + 1990) %>% 
  filter(climate != "none", chapter %in% c("none", "both", "chap_1")) %>% 
  group_by(scenario, species, iteration) %>% 
  group_modify(~mutate(.x , baseline = subset(.x, timestep == 2010)$numLinks, 
                       the_diff = ((numLinks-baseline)/baseline)*100 )) %>%
  ggplot(aes(x=timestep, y=the_diff, col=climate)) +
  scale_color_manual(values = colors_sce$color, 
                     labels = colors_sce$sce) +
  geom_smooth(aes(group = sce, linetype=code_run), alpha=0.5, method="loess") +
  #geom_point(aes(group = sce)) +
  scale_linetype_manual(values=c("solid", "twodash")) +
  facet_wrap(~species, scales = "fixed") +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  #geom_vline(xintercept = 2020, linetype = "dashed", color = "darkGreen", alpha = 0.5) +
  labs(y = "numLinks",
       x = "Year",
       col = "Climate",
       linetype = "Run") +
  guides(col = guide_legend(ncol = 2, nrow=2), 
         linetype = guide_legend(nrow=1, override.aes = list(colour = 'black'))) +
  theme(legend.position = c(0.85, 0.15))+
  NULL -> net_4

ggsave(net_4, 
       filename = "../thesis/figures/NUMLINKS.png", 
       width = the_width, height = the_height, dpi = the_dpi)

```

```{r, message=FALSE, warning=FALSE, out.width="100%", fig.cap=""}
networks %>% 
  mutate(timestep = timestep*10 + 1990) %>% 
  filter(climate != "none", chapter %in% c("none", "both", "chap_1")) %>% 
  group_by(scenario, species, iteration) %>% 
  group_modify(~mutate(.x , baseline = subset(.x, timestep == 2010)$meanAreaQuality, 
                       the_diff = ((meanAreaQuality-baseline)/baseline)*100 )) %>%
  ggplot(aes(x=timestep, y=the_diff, col=climate)) +
  scale_color_manual(values = colors_sce$color, 
                     labels = colors_sce$sce) +
  geom_smooth(aes(group = sce, linetype=code_run), alpha=0.5, method="loess") +
  #geom_point(aes(group = sce)) +
  scale_linetype_manual(values=c("solid", "twodash")) +
  facet_wrap(~species, scales = "fixed") +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  #geom_vline(xintercept = 2020, linetype = "dashed", color = "darkGreen", alpha = 0.5) +
  labs(y = "meanAreaQuality",
       x = "Year",
       col = "Climate",
       linetype = "Run") +
  guides(col = guide_legend(ncol = 2, nrow=2), 
         linetype = guide_legend(nrow=1, override.aes = list(colour = 'black'))) +
  theme(legend.position = c(0.85, 0.15))+
  NULL -> net_5

ggsave(net_5, 
       filename = "../thesis/figures/meanAreaQuality.png", 
       width = the_width, height = the_height, dpi = the_dpi)

```

```{r, message=FALSE, warning=FALSE, out.width="100%", fig.cap=""}
networks %>% 
  mutate(timestep = timestep*10 + 1990) %>% 
  filter(climate != "none", chapter %in% c("none", "both", "chap_1")) %>% 
  group_by(scenario, species, iteration) %>% 
  group_modify(~mutate(.x , baseline = subset(.x, timestep == 2010)$meanCost, 
                       the_diff = ((meanCost-baseline)/baseline)*100 )) %>%
  ggplot(aes(x=timestep, y=the_diff, col=climate)) +
  scale_color_manual(values = colors_sce$color, 
                     labels = colors_sce$sce) +
  geom_smooth(aes(group = sce, linetype=code_run), alpha=0.5, method="loess") +
  #geom_point(aes(group = sce)) +
  scale_linetype_manual(values=c("solid", "twodash")) +
  facet_wrap(~species, scales = "fixed") +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  #geom_vline(xintercept = 2020, linetype = "dashed", color = "darkGreen", alpha = 0.5) +
  labs(y = "meanCost",
       x = "Year",
       col = "Climate",
       linetype = "Run") +
  guides(col = guide_legend(ncol = 2, nrow=2), 
         linetype = guide_legend(nrow=1, override.aes = list(colour = 'black'))) +
  theme(legend.position = c(0.85, 0.15))+
  NULL -> net_6

ggsave(net_6, 
       filename = "../thesis/figures/meanCost.png", 
       width = the_width, height = the_height, dpi = the_dpi)

```

##### CHAP 2
```{r, message=FALSE, warning=FALSE, out.width="100%", fig.cap=""}
networks %>% 
  mutate(timestep = timestep*10 + 1990) %>% 
  filter(climate != "none", chapter %in% c("none", "both", "chap_2")) %>% 
  group_by(scenario, species, iteration) %>% 
  group_modify(~mutate(.x , baseline = subset(.x, timestep == 2010)$overallECgap, 
                       the_diff = ((overallECgap-baseline)/baseline)*100 )) %>% 
  ggplot(aes(x=timestep, y=the_diff, col=climate)) +
  scale_color_manual(values = colors_sce$color,
                     labels = colors_sce$sce) +
  geom_smooth(aes(group = sce, linetype=code_run), alpha=0.5, method="loess") +
  scale_linetype_manual(values=c("solid", "dashed", "longdash", "dotted")) +
  facet_wrap(~species, scales = "fixed") +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  geom_vline(xintercept = 2020, linetype = "dashed", color = "darkGreen", alpha = 0.5) +
  labs(y = "overallECgap",
       x = "Year",
       col = "Climate",
       linetype = "Run") +
  guides(col = guide_legend(ncol = 2, nrow=2), 
         linetype = guide_legend(nrow = 2, ncol=2, 
                                 override.aes = list(colour = 'black'))) +
  theme(legend.position = c(0.85, 0.15))+
  NULL -> net_1_2

ggsave(net_1_2, 
       filename = "../thesis/figures/ECGAP2.png", 
       width = the_width, height = the_height, dpi = the_dpi)
```

```{r, message=FALSE, warning=FALSE, out.width="100%", fig.cap=""}
networks %>% 
  mutate(timestep = timestep*10 + 1990) %>% 
  filter(climate != "none", chapter %in% c("none", "both", "chap_2")) %>% 
  group_by(scenario, species, iteration) %>% 
  group_modify(~mutate(.x , baseline = subset(.x, timestep == 2010)$overallECnatal, 
                       the_diff = ((overallECnatal-baseline)/baseline)*100 )) %>% 
  ggplot(aes(x=timestep, y=the_diff, col=climate)) +
  scale_color_manual(values = colors_sce$color,
                     labels = colors_sce$sce) +
  geom_smooth(aes(group = sce, linetype=code_run), alpha=0.5, method="loess") +
  scale_linetype_manual(values=c("solid", "dashed", "longdash", "dotted")) +
  facet_wrap(~species, scales = "fixed") +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  geom_vline(xintercept = 2020, linetype = "dashed", color = "darkGreen", alpha = 0.5) +
  labs(y = "overallECnatal",
       x = "Year",
       col = "Climate",
       linetype = "Run") +
  guides(col = guide_legend(ncol = 2, nrow=2), 
         linetype = guide_legend(nrow = 2, ncol=2, 
                                 override.aes = list(colour = 'black'))) +
  theme(legend.position = c(0.85, 0.15))+
  NULL -> net_2_2

ggsave(net_2_2, 
       filename = "../thesis/figures/ECNATAL2.png", 
       width = the_width, height = the_height, dpi = the_dpi)
```

```{r, message=FALSE, warning=FALSE, out.width="100%", fig.cap=""}
networks %>% 
  mutate(timestep = timestep*10 + 1990) %>% 
  filter(climate != "none", chapter %in% c("none", "both", "chap_2")) %>% 
  group_by(scenario, species, iteration) %>% 
  group_modify(~mutate(.x , baseline = subset(.x, timestep == 2010)$numNodes, 
                       the_diff = ((numNodes-baseline)/baseline)*100 )) %>% 
  ggplot(aes(x=timestep, y=the_diff, col=climate)) +
  scale_color_manual(values = colors_sce$color,
                     labels = colors_sce$sce) +
  geom_smooth(aes(group = sce, linetype=code_run), alpha=0.5, method="loess") +
  scale_linetype_manual(values=c("solid", "dashed", "longdash", "dotted")) +
  facet_wrap(~species, scales = "fixed") +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  geom_vline(xintercept = 2020, linetype = "dashed", color = "darkGreen", alpha = 0.5) +
  labs(y = "numNodes",
       x = "Year",
       col = "Climate",
       linetype = "Run") +
  guides(col = guide_legend(ncol = 2, nrow=2), 
         linetype = guide_legend(nrow = 2, ncol=2, 
                                 override.aes = list(colour = 'black'))) +
  theme(legend.position = c(0.85, 0.15))+
  NULL -> net_3_2

ggsave(net_3_2, 
       filename = "../thesis/figures/NUMNODES2.png", 
       width = the_width, height = the_height, dpi = the_dpi)
```

```{r, message=FALSE, warning=FALSE, out.width="100%", fig.cap=""}
networks %>% 
  mutate(timestep = timestep*10 + 1990) %>% 
  filter(climate != "none", chapter %in% c("none", "both", "chap_2")) %>% 
  group_by(scenario, species, iteration) %>% 
  group_modify(~mutate(.x , baseline = subset(.x, timestep == 2010)$numLinks, 
                       the_diff = ((numLinks-baseline)/baseline)*100 )) %>% 
  ggplot(aes(x=timestep, y=the_diff, col=climate)) +
  scale_color_manual(values = colors_sce$color,
                     labels = colors_sce$sce) +
  geom_smooth(aes(group = sce, linetype=code_run), alpha=0.5, method="loess") +
  scale_linetype_manual(values=c("solid", "dashed", "longdash", "dotted")) +
  facet_wrap(~species, scales = "fixed") +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  geom_vline(xintercept = 2020, linetype = "dashed", color = "darkGreen", alpha = 0.5) +
  labs(y = "numLinks",
       x = "Year",
       col = "Climate",
       linetype = "Run") +
  guides(col = guide_legend(ncol = 2, nrow=2), 
         linetype = guide_legend(nrow = 2, ncol=2, 
                                 override.aes = list(colour = 'black'))) +
  theme(legend.position = c(0.85, 0.15))+
  NULL -> net_4_2

ggsave(net_4_2, 
       filename = "../thesis/figures/NUMLINKS2.png", 
       width = the_width, height = the_height, dpi = the_dpi)
```

```{r, message=FALSE, warning=FALSE, out.width="100%", fig.cap=""}
networks %>% 
  mutate(timestep = timestep*10 + 1990) %>% 
  filter(climate != "none", chapter %in% c("none", "both", "chap_2")) %>% 
  group_by(scenario, species, iteration) %>% 
  group_modify(~mutate(.x , baseline = subset(.x, timestep == 2010)$meanAreaQuality, 
                       the_diff = ((meanAreaQuality-baseline)/baseline)*100 )) %>% 
  ggplot(aes(x=timestep, y=the_diff, col=climate)) +
  scale_color_manual(values = colors_sce$color,
                     labels = colors_sce$sce) +
  geom_smooth(aes(group = sce, linetype=code_run), alpha=0.5, method="loess") +
  scale_linetype_manual(values=c("solid", "dashed", "longdash", "dotted")) +
  facet_wrap(~species, scales = "fixed") +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  geom_vline(xintercept = 2020, linetype = "dashed", color = "darkGreen", alpha = 0.5) +
  labs(y = "meanAreaQuality",
       x = "Year",
       col = "Climate",
       linetype = "Run") +
  guides(col = guide_legend(ncol = 2, nrow=2), 
         linetype = guide_legend(nrow = 2, ncol=2, 
                                 override.aes = list(colour = 'black'))) +
  theme(legend.position = c(0.85, 0.15))+
  NULL -> net_5_2

ggsave(net_5_2, 
       filename = "../thesis/figures/meanAreaQuality_2.png", 
       width = the_width, height = the_height, dpi = the_dpi)
```

```{r, message=FALSE, warning=FALSE, out.width="100%", fig.cap=""}
networks %>% 
  mutate(timestep = timestep*10 + 1990) %>% 
  filter(climate != "none", chapter %in% c("none", "both", "chap_2")) %>% 
  group_by(scenario, species, iteration) %>% 
  group_modify(~mutate(.x , baseline = subset(.x, timestep == 2010)$meanCost, 
                       the_diff = ((meanCost-baseline)/baseline)*100 )) %>% 
  ggplot(aes(x=timestep, y=the_diff, col=climate)) +
  scale_color_manual(values = colors_sce$color,
                     labels = colors_sce$sce) +
  geom_smooth(aes(group = sce, linetype=code_run), alpha=0.5, method="loess") +
  scale_linetype_manual(values=c("solid", "dashed", "longdash", "dotted")) +
  facet_wrap(~species, scales = "fixed") +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  geom_vline(xintercept = 2020, linetype = "dashed", color = "darkGreen", alpha = 0.5) +
  labs(y = "meanCost",
       x = "Year",
       col = "Climate",
       linetype = "Run") +
  guides(col = guide_legend(ncol = 2, nrow=2), 
         linetype = guide_legend(nrow = 2, ncol=2, 
                                 override.aes = list(colour = 'black'))) +
  theme(legend.position = c(0.85, 0.15))+
  NULL -> net_6_2

ggsave(net_6_2, 
       filename = "../thesis/figures/meanCost_2.png", 
       width = the_width, height = the_height, dpi = the_dpi)
```

```{r}
networks %>% 
  filter(scenario != 37, timestep %in% c(2,10)) %>% 
  filter(species %in% c("MAAM", "URAM")) %>% 
  pivot_longer(numNodes:overallECnatal, names_to = "metric", values_to = "values") %>% 
  select (species, iteration, timestep, code_run, climate, metric, values) %>% 
  group_by(species, timestep, code_run, climate, metric) %>% 
  summarise(values = mean(values)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = timestep, names_glue = "ts_{timestep}_{.value}", values_from = values) %>% 
  mutate(percent_change = (ts_10_values - ts_2_values)/ts_2_values) %>% 
  select(-starts_with("ts")) %>% 
  pivot_wider(names_from = metric, values_from = percent_change) %>% 
  mutate(code_run = fct_relevel(code_run, c("BAU", "BAUR", "CP", "CPR", "CPTR"))) %>% 
  arrange(code_run) %>% 
  mutate(climate = fct_relevel(climate, c("baseline", "RCP 8.5")))  %>% 
  arrange(code_run, climate) %>% 
  mutate(scenario = paste(code_run, climate)) %>%
  arrange(code_run, climate, scenario) %>% 
  select(-c(code_run, climate)) %>%
  select(species, scenario, starts_with("overall"), numNodes, numLinks, meanAreaQuality, meanCost) %>%

  gt(rowname_col = "scenario", groupname_col = "species") %>% 
  fmt_percent(everything()) %>% 
  cols_label(
    overallECgap = "EC (gap)", 
    overallECnatal = "EC (natal)", 
    numNodes = "Nodes", 
    numLinks = "Links", 
    meanAreaQuality = "Mean Quality", 
    meanCost = "Mean Cost"
  ) %>% 
  data_color(
    columns = overallECgap:meanAreaQuality, 
    colors = scales::col_bin(palette = as.character(paletteer::paletteer_c("scico::cork", n = 100)), 
                             bins = seq(from =-1, to = 1, by = 0.1))) %>% 
  data_color(
    columns = meanCost, 
    colors = scales::col_bin(palette = as.character(paletteer::paletteer_c("scico::cork", n = 100)), 
                             bins = seq(from =-1, to = 1, by = 0.1), reverse = -1)) -> metrics_table 
metrics_table
gtsave(metrics_table, "../thesis/figures/metrics.png")

```

```{r}
networks_tbl <- networks
# networks_tbl$code_run[networks_tbl$code_run == "BAUR"] <- "\n    BAUR"
# networks_tbl$code_run[networks_tbl$code_run == "CP"] <- "    CP"
# networks_tbl$code_run[networks_tbl$climate == "RCP 8.5"] <- "RCP 8.5"
networks_tbl %>% 
  filter(scenario != 37, timestep %in% c(2,10)) %>% 
  # filter(species %in% c("MAAM", "URAM")) %>% 
  pivot_longer(numNodes:overallECnatal, names_to = "metric", values_to = "values") %>% 
  select (species, iteration, timestep, code_run, climate, metric, values) %>% 
  group_by(species, timestep, code_run, climate, metric) %>% 
  summarise(values = mean(values)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = timestep, names_glue = "ts_{timestep}_{.value}", values_from = values) %>% 
  mutate(percent_change = (ts_10_values - ts_2_values)/ts_2_values) %>% 
  select(-starts_with("ts")) %>%

  # pivot_wider(names_from = metric, values_from = percent_change) %>%

  mutate(code_run = fct_relevel(code_run, c("BAU", "BAUR", "CP", "CPR", "CPTR"))) %>% 
  arrange(code_run) %>% 
  mutate(climate = fct_relevel(climate, c("baseline", "RCP 8.5")))  %>% 
  arrange(code_run, climate) %>% 
  mutate(scenario = paste(code_run, climate)) %>%
  arrange(code_run, climate, scenario) %>%

  select(-c(code_run, climate)) %>%

  pivot_wider(names_from = "scenario", values_from = "percent_change") %>%

  # select(species, scenario, starts_with("overall"), numNodes, numLinks, meanAreaQuality, meanCost) %>% 
  filter(metric %in% c("overallECgap", "overallECnatal", "numNodes", 
                       "numLinks", "meanAreaQuality")) %>%

  # mutate(metric = fct_relevel(metric, c("overallECgap", "overallECnatal", "numNodes", 
  #                                       "numLinks", "meanAreaQuality"))) %>% 
  left_join(data.frame(metric = c("overallECgap", "overallECnatal", "numNodes", 
                                  "numLinks", "meanAreaQuality"), 
                       metric_name = c("EC (gap)", "EC (natal)", "Nodes", 
                                       "Links", "Quality"))) %>% 
  mutate(metric_name = fct_relevel(metric_name, c("EC (gap)", "EC (natal)", "Nodes", 
                                                  "Links", "Quality"))) %>%
  arrange(metric_name) %>%
  select(-metric) %>%

  gt(rowname_col = "metric_name", groupname_col = "species") %>% 
  fmt_percent(everything(), decimals=0) %>% 
  data_color(
    columns = `BAU baseline`:`CPTR RCP 8.5`, 
    colors = scales::col_bin(palette = as.character(paletteer::paletteer_c("scico::cork", n = 100)), 
                             bins = seq(from =-1, to = 1, by = 0.1))) %>%
  tab_options(data_row.padding = px(1)) %>% 
  cols_width(metric_name~px(89), everything()~px(68)) %>% 
  tab_style(locations = cells_body(), cell_text(size = 12)) -> metrics_table_2 # %>% # 89
# data_color(
#   columns = meanCost, 
#   colors = scales::col_bin(palette = as.character(paletteer::paletteer_c("scico::cork", n = 100)), 
#                            bins = seq(from =-1, to = 1, by = 0.1), reverse = -1))

metrics_table_2
gtsave(metrics_table_2, "../thesis/figures/metrics_2.png")

```

```{r}
URAM <- networks %>% 
  filter(scenario != 37, timestep %in% c(2,10)) %>% 
  pivot_longer(numNodes:overallECnatal, names_to = "metric", values_to = "values") %>% 
  select (species, iteration, timestep, code_run, climate, metric, values) %>% 
  group_by(species, timestep, code_run, climate, metric) %>% 
  summarise(values = mean(values)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = timestep, names_glue = "ts_{timestep}_{.value}", values_from = values) %>% 
  mutate(percent_change = (ts_10_values - ts_2_values)/ts_2_values*100) %>% 
  select(-starts_with("ts")) %>% 
  mutate(code_run = fct_relevel(code_run, c("BAU", "BAUR", "CP", "CPR", "CPTR"))) %>% 
  arrange(code_run) %>% 
  mutate(climate = fct_relevel(climate, c("baseline", "RCP 8.5")))  %>% 
  arrange(code_run, climate) %>% 
  mutate(scenario = paste(code_run, climate)) %>%
  arrange(code_run, climate, scenario) %>%

  select(-c(code_run, climate)) %>% 
  ungroup() %>% 
  filter(species == "URAM") %>% 
  filter(metric %in% c("overallECgap", "overallECnatal", "numNodes", 
                       "numLinks", "meanAreaQuality")) %>% 
  left_join(data.frame(metric = c("overallECgap", "overallECnatal", "numNodes", 
                                  "numLinks", "meanAreaQuality"), 
                       metric_name = c("EC (gap)", "EC (natal)", "Nodes", 
                                       "Links", "Quality"))) %>% 
  mutate(metric_name = fct_relevel(metric_name, c("EC (gap)", "EC (natal)", "Nodes", 
                                                  "Links", "Quality"))) %>%
  arrange(metric_name) %>%
  select(-metric) %>% 
  mutate(`% Change` = percent_change, Metric = metric_name, Scenario = scenario) %>% 
  tidyHeatmap::heatmap(Scenario, Metric, `% Change`, .scale="none", cluster_columns = FALSE)

BLBR <- networks %>% 
  filter(scenario != 37, timestep %in% c(2,10)) %>% 
  pivot_longer(numNodes:overallECnatal, names_to = "metric", values_to = "values") %>% 
  select (species, iteration, timestep, code_run, climate, metric, values) %>% 
  group_by(species, timestep, code_run, climate, metric) %>% 
  summarise(values = mean(values)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = timestep, names_glue = "ts_{timestep}_{.value}", values_from = values) %>% 
  mutate(percent_change = (ts_10_values - ts_2_values)/ts_2_values*100) %>% 
  select(-starts_with("ts")) %>% 
  mutate(code_run = fct_relevel(code_run, c("BAU", "BAUR", "CP", "CPR", "CPTR"))) %>% 
  arrange(code_run) %>% 
  mutate(climate = fct_relevel(climate, c("baseline", "RCP 8.5")))  %>% 
  arrange(code_run, climate) %>% 
  mutate(scenario = paste(code_run, climate)) %>%
  arrange(code_run, climate, scenario) %>%

  select(-c(code_run, climate)) %>% 
  ungroup() %>% 
  filter(species == "BLBR") %>% 
  filter(metric %in% c("overallECgap", "overallECnatal", "numNodes", 
                       "numLinks", "meanAreaQuality")) %>% 
  left_join(data.frame(metric = c("overallECgap", "overallECnatal", "numNodes", 
                                  "numLinks", "meanAreaQuality"), 
                       metric_name = c("EC (gap)", "EC (natal)", "Nodes", 
                                       "Links", "Quality"))) %>% 
  mutate(metric_name = fct_relevel(metric_name, c("EC (gap)", "EC (natal)", "Nodes", 
                                                  "Links", "Quality"))) %>%
  arrange(metric_name) %>%
  select(-metric) %>% 
  mutate(`% Change` = percent_change, Metric = metric_name, Scenario = scenario) %>% 
  tidyHeatmap::heatmap(Scenario, Metric, `% Change`, .scale="none", cluster_columns = FALSE)

MAAM <- networks %>% 
  filter(scenario != 37, timestep %in% c(2,10)) %>% 
  pivot_longer(numNodes:overallECnatal, names_to = "metric", values_to = "values") %>% 
  select (species, iteration, timestep, code_run, climate, metric, values) %>% 
  group_by(species, timestep, code_run, climate, metric) %>% 
  summarise(values = mean(values)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = timestep, names_glue = "ts_{timestep}_{.value}", values_from = values) %>% 
  mutate(percent_change = (ts_10_values - ts_2_values)/ts_2_values*100) %>% 
  select(-starts_with("ts")) %>% 
  mutate(code_run = fct_relevel(code_run, c("BAU", "BAUR", "CP", "CPR", "CPTR"))) %>% 
  arrange(code_run) %>% 
  mutate(climate = fct_relevel(climate, c("baseline", "RCP 8.5")))  %>% 
  arrange(code_run, climate) %>% 
  mutate(scenario = paste(code_run, climate)) %>%
  arrange(code_run, climate, scenario) %>%

  select(-c(code_run, climate)) %>% 
  ungroup() %>% 
  filter(species == "MAAM") %>% 
  filter(metric %in% c("overallECgap", "overallECnatal", "numNodes", 
                       "numLinks", "meanAreaQuality")) %>% 
  left_join(data.frame(metric = c("overallECgap", "overallECnatal", "numNodes", 
                                  "numLinks", "meanAreaQuality"), 
                       metric_name = c("EC (gap)", "EC (natal)", "Nodes", 
                                       "Links", "Quality"))) %>% 
  mutate(metric_name = fct_relevel(metric_name, c("EC (gap)", "EC (natal)", "Nodes", 
                                                  "Links", "Quality"))) %>%
  arrange(metric_name) %>%
  select(-metric) %>% 
  mutate(`% Change` = percent_change, Metric = metric_name, Scenario = scenario) %>% 
  tidyHeatmap::heatmap(Scenario, Metric, `% Change`, .scale="none", cluster_columns = FALSE)

PLCI <- networks %>% 
  filter(scenario != 37, timestep %in% c(2,10)) %>% 
  pivot_longer(numNodes:overallECnatal, names_to = "metric", values_to = "values") %>% 
  select (species, iteration, timestep, code_run, climate, metric, values) %>% 
  group_by(species, timestep, code_run, climate, metric) %>% 
  summarise(values = mean(values)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = timestep, names_glue = "ts_{timestep}_{.value}", values_from = values) %>% 
  mutate(percent_change = (ts_10_values - ts_2_values)/ts_2_values*100) %>% 
  select(-starts_with("ts")) %>% 
  mutate(code_run = fct_relevel(code_run, c("BAU", "BAUR", "CP", "CPR", "CPTR"))) %>% 
  arrange(code_run) %>% 
  mutate(climate = fct_relevel(climate, c("baseline", "RCP 8.5")))  %>% 
  arrange(code_run, climate) %>% 
  mutate(scenario = paste(code_run, climate)) %>%
  arrange(code_run, climate, scenario) %>%

  select(-c(code_run, climate)) %>% 
  ungroup() %>% 
  filter(species == "PLCI") %>% 
  filter(metric %in% c("overallECgap", "overallECnatal", "numNodes", 
                       "numLinks", "meanAreaQuality")) %>% 
  left_join(data.frame(metric = c("overallECgap", "overallECnatal", "numNodes", 
                                  "numLinks", "meanAreaQuality"), 
                       metric_name = c("EC (gap)", "EC (natal)", "Nodes", 
                                       "Links", "Quality"))) %>% 
  mutate(metric_name = fct_relevel(metric_name, c("EC (gap)", "EC (natal)", "Nodes", 
                                                  "Links", "Quality"))) %>%
  arrange(metric_name) %>%
  select(-metric) %>% 
  mutate(`% Change` = percent_change, Metric = metric_name, Scenario = scenario) %>% 
  tidyHeatmap::heatmap(Scenario, Metric, `% Change`, .scale="none", cluster_columns = FALSE)

RASY <- networks %>% 
  filter(scenario != 37, timestep %in% c(2,10)) %>% 
  pivot_longer(numNodes:overallECnatal, names_to = "metric", values_to = "values") %>% 
  select (species, iteration, timestep, code_run, climate, metric, values) %>% 
  group_by(species, timestep, code_run, climate, metric) %>% 
  summarise(values = mean(values)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = timestep, names_glue = "ts_{timestep}_{.value}", values_from = values) %>% 
  mutate(percent_change = (ts_10_values - ts_2_values)/ts_2_values*100) %>% 
  select(-starts_with("ts")) %>% 
  mutate(code_run = fct_relevel(code_run, c("BAU", "BAUR", "CP", "CPR", "CPTR"))) %>% 
  arrange(code_run) %>% 
  mutate(climate = fct_relevel(climate, c("baseline", "RCP 8.5")))  %>% 
  arrange(code_run, climate) %>% 
  mutate(scenario = paste(code_run, climate)) %>%
  arrange(code_run, climate, scenario) %>%

  select(-c(code_run, climate)) %>% 
  ungroup() %>% 
  filter(species == "RASY") %>% 
  filter(metric %in% c("overallECgap", "overallECnatal", "numNodes", 
                       "numLinks", "meanAreaQuality")) %>% 
  left_join(data.frame(metric = c("overallECgap", "overallECnatal", "numNodes", 
                                  "numLinks", "meanAreaQuality"), 
                       metric_name = c("EC (gap)", "EC (natal)", "Nodes", 
                                       "Links", "Quality"))) %>% 
  mutate(metric_name = fct_relevel(metric_name, c("EC (gap)", "EC (natal)", "Nodes", 
                                                  "Links", "Quality"))) %>%
  arrange(metric_name) %>%
  select(-metric) %>% 
  mutate(`% Change` = percent_change, Metric = metric_name, Scenario = scenario) %>% 
  tidyHeatmap::heatmap(Scenario, Metric, `% Change`, .scale="none", cluster_columns = FALSE)
```

```{r}
mtcars_tidy <- 
  mtcars %>% 
  as_tibble(rownames="Car name") %>% 
  mutate_at(vars(-`Car name`, -hp, -vs), scale) %>%
  pivot_longer(cols = -c(`Car name`, hp, vs), names_to = "Property", values_to = "Value")
mtcars_tidy_nested <- bind_rows(mutate(mtcars_tidy, Year=1990), mutate(mtcars_tidy, Year=2000)) %>% 
  mutate(Value = runif(n = n(), min = 0, max = 10))
mtcars_tidy_nested %>% 
  heatmap(`Car name`, Property, Value) 
```

#### HEATMAP

```{r}
HEATMAP_data <- networks %>% 
  filter(scenario != 37, timestep %in% c(2,10)) %>% 
  pivot_longer(numNodes:overallECnatal, names_to = "metric", values_to = "values") %>% 
  select (species, iteration, timestep, code_run, climate, metric, values) %>% 
  group_by(species, timestep, code_run, climate, metric) %>% 
  summarise(values = mean(values)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = timestep, names_glue = "ts_{timestep}_{.value}", values_from = values) %>% 
  mutate(percent_change = (ts_10_values - ts_2_values)/ts_2_values*100) %>% 
  select(-starts_with("ts")) %>% 
  mutate(code_run = fct_relevel(code_run, c("BAU", "BAUR", "CP", "CPR", "CPTR"))) %>% 
  arrange(code_run) %>% 
  mutate(climate = fct_relevel(climate, c("baseline", "RCP 8.5")))  %>% 
  arrange(code_run, climate) %>% 
  mutate(scenario = paste(code_run, climate)) %>%
  arrange(code_run, climate, scenario) %>%

  select(-c(code_run, climate)) %>% 
  ungroup() %>% 
  #filter(species == "URAM") %>% 
  filter(metric %in% c("overallECgap", "overallECnatal", "numNodes", 
                       "numLinks", "meanAreaQuality")) %>% 
  left_join(data.frame(metric = c("overallECgap", "overallECnatal", "numNodes", 
                                  "numLinks", "meanAreaQuality"), 
                       metric_name = c("EC (gap)", "EC (natal)", "Nodes", 
                                       "Links", "Quality"))) %>% 
  mutate(metric_name = fct_relevel(metric_name, c("EC (gap)", "EC (natal)", "Nodes",
                                                  "Links", "Quality"))) %>%
  arrange(metric_name) %>%
  select(-metric) %>% 
  mutate(`% Change` = percent_change, `Connectivity Metric` = metric_name, `Species / Scenario` = scenario) %>% 
  select(species, `% Change`, `Connectivity Metric`, `Species / Scenario`)

titles <- expand_grid(left = c("BAU", "BAUR", "CP", "CPR", "CPTR"), right = c("baseline", "RCP 8.5")) %>% mutate(final = paste(left, right)) %>% pull(final) %>% sort()

HEATMAP_data %>% 
  unite( "Species / Scenario", c(`Species / Scenario`, species), remove = F) %>%  group_by(species) %>% 
  tidyHeatmap::heatmap(`Species / Scenario`, `Connectivity Metric`, `% Change`, .scale="none", cluster_columns = FALSE, 
                       cluster_rows = FALSE, palette_value = c("red", "white", "darkGreen"), 
                       rect_gp = grid::gpar(col = "#161616", lwd = 0.1),
                       palette_grouping = list("grey"), row_labels = rep(titles, each = 5), 
                       heatmap_legend_param = list(at = c(-60, 0 , 60))) -> heatmap_networks

zissou_palette_custom <- c("#F21A00", "#E67D00", "#E4B80E",
                           "#FFFFFF",
                           "#9EBE91", "#4EA365", "#315849")

HEATMAP_data %>% 
  unite( "Species / Scenario", c(`Species / Scenario`, species), remove = F) %>%  group_by(species) %>% 
  tidyHeatmap::heatmap(`Species / Scenario`, `Connectivity Metric`, `% Change`, .scale="none", cluster_columns = FALSE, 
                       cluster_rows = FALSE, palette_value = 
                         circlize::colorRamp2(c(-60, -40, -20, 0, 20, 40, 60), 
                                              zissou_palette_custom), 
                       # rect_gp = grid::gpar(col = "#161616", lwd = 0.1),
                       palette_grouping = list("grey"), row_labels = rep(titles, each = 5), 
                       heatmap_legend_param = list(at = c(-60, -40, -20, 0, 20, 40, 60))
  ) -> heatmap_networks

heatmap_networks %>% 
  tidyHeatmap::save_pdf(filename = "../thesis/figures/heatmap_networks.pdf")
# width = the_width, height = the_height)
# heatmap(`Car name`, Property, Value) 
```

```{r}
joined %>% 
  filter(climate != "none", chapter %in% c("none", "both", "chap_1", "chap_2")) %>% 
  group_by(scenarioId, species, iteration) %>% 
  group_modify(~mutate(.x , baseline = subset(.x, timestep == 2010)$sum_cur, 
                       the_diff = ((sum_cur-baseline)/baseline)*100 )) %>% 
  ungroup() %>% 
  filter(timestep == 2100) %>%
  group_by(code, species) %>% 
  summarise(the_diff = mean(the_diff)) %>% 
  ungroup() %>% 
  mutate(`% Change` = the_diff, Species = species, Scenario = code) %>% 
  select(Species, `% Change`, Scenario) -> current_heatmap_df

current_heatmap_df %>% 
  # unite( "Scenario", c(Scenario, Species), remove = F) %>%  group_by(Species) %>% 
  tidyHeatmap::heatmap(Scenario, Species, `% Change`, .scale="none", cluster_columns = TRUE, 
                       cluster_rows = TRUE,
                       palette_value = circlize::colorRamp2(c(-15, -10, -5, 0), viridis::magma(4))) -> heatmap_current
```
#### Scenario map

```{r}
library(tmap)
library(stars)

classes <- read_csv("../config/rcl_tables/land_use/recode_table.csv")

corridors <- st_read("../data/raw/workshop/corridors/Key_links_smoothed_RegTables.shp", 
                     quiet = TRUE) %>% sf::st_make_valid() %>% st_buffer(1000) %>% 
  mutate(type = "Corridors")
areas <- st_read("../data/raw/workshop/ensembles/Key_ensembles_smoothed_RegTables.shp", 
                 quiet = TRUE) %>% sf::st_make_valid()%>% 
  mutate(type = "Cores")
background <- raster("../data/land_use/LandUse_mont_aafc_buffered_30by30_2010.tif")
background_cropped <- crop(background, st_transform(areas, st_crs(background)))

background_stars <- stars::st_as_stars(background_cropped) %>% 
  mutate(`Land Use Class` = 
           as.factor(case_when(LandUse_mont_aafc_buffered_30by30_2010 == 1 ~ "Agriculture", 
                               LandUse_mont_aafc_buffered_30by30_2010 %in% c(2,4) ~ "Urban / Roads",
                               LandUse_mont_aafc_buffered_30by30_2010 == 3 ~ "Forest",
                               LandUse_mont_aafc_buffered_30by30_2010 == 5 ~ "Water",
                               LandUse_mont_aafc_buffered_30by30_2010 == 6 ~ "Wetlands")))

tmap_mode("plot")
tmap_options(check.and.fix = TRUE)
test <- tm_shape(background_stars, raster.downsample = FALSE) + 
  tm_raster("Land.Use.Class", palette = c("#F9E79F", # Ag
                                          "#58D68D", # Fo
                                          "#ABB2B9", # Urb / Rd
                                          "#5DADE2", # Wat
                                          "#A569BD"),# Wet
            title = "Land Use Class") +
  tm_shape(areas) + 
  tm_borders(lwd=1, col ="#283747") + 
  tm_fill("type", col = "#196F3D", alpha = 0.3) +
  tm_shape(corridors) + 
  tm_borders(lwd=2, col = "#0B5345", lty = 3) +
  tm_layout(legend.bg.color = "white", legend.bg.alpha = .9, 
            legend.frame = "gray50")
tmap_save(test, "../thesis/figures/corridors_cores.png")
```